---
title: "Figures_Botocudo_manuscript"
author: "Diana Cruz"
date: "2023-02-14"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, warning = F, message = F, fig.align = "center")
```

# Preface

This RMarkdown was built to recreate the main and supplementary figures as well as the figures included in the supplementary datasets released along with the manuscript at https://doi.org/10.1101/2022.01.27.477466
(Indigenous peoples in eastern Brazil: insights from 19th century genomes and metagenomes; Cruz Dávalos et al., 2022, bioRxiv).


```{r set working paths}
require(stringr)

# Please change the prefix to indicate the directory that contains the
# "final_inputs" folder.
prefix_data <- "~/Projects/Botocudos/Files/backup"

# Functions written to help with the plotting
# Please change the prefix to indicate the directory that contains the
# "scripts" folder.
prefix_scripts <- "~/Projects/Botocudos/Files/backup/scripts"

source(str_glue("{prefix_scripts}/utilities_ngsadmix.R"))
source(str_glue("{prefix_scripts}/f3_utilities_plotting.R"))
source(str_glue("{prefix_scripts}/plot_Dstats.R"))

source(str_glue("{prefix_scripts}/plot_fsc2_functions.R"))
```



```{r install packages, eval = F}
# Run/adapt this part if any of the packages are not installed
for(
  package in c(
    "stringr", "scales", "ggplot2", "cowplot", "plyr", "mapdata", "colorspace",
    "data.table", "gridBase", "grid", "tidyr", "gg.gap"
  )
){
  if(!require(package)) install.packages(package)
}


if(!require("ComplexHeatmap")){
  # The developers of ComplexHeatMap suggest the install_github function from devtools
  # In my case, the install_github from the remotes package worked better
  # https://jokergoo.github.io/ComplexHeatmap-reference/book/
  if(!require("remotes")){
    install.packages("remotes")
    remotes::install_github("jokergoo/ComplexHeatmap")
  }
}

```


```{r load packages and functions}
require(stringr)
require(scales)
require(ggplot2)
require(cowplot)
require(plyr)
require(mapdata)
require(colorspace)
require(data.table)
require(gridBase)
require(grid)
suppressPackageStartupMessages(library(ComplexHeatmap))
require(tidyr)
require(gg.gap)
```


# Main figures


## Figure 1

Figure 1 contains a map (1A) and the calibrated radiocarbon dates (1B) for 22 (out of 24) individuals sequenced in the project, as well as two individuals (Polynesian-"Botocudos") published in 2014 (Malaspinas et al.).

### Fig. 1A Map

The part of the map was made in two parts. The main part corresponds to the map of Brazil, with the second part being a small inlet of the Americas.
These two maps were imported to Illustrator, and the labels, colors, and legend were adjusted there.


```{r map Americas, fig.width=10, fig.height=10}
# png("~/Projects/Botocudos/Plots/map_DBC.png",
#     width = 20, height = 15, res = 300, units = "in")

# pdf("~/Projects/Botocudos/Plots/map_DBC.pdf", width = 20, height = 15)

# layout(my_mat, widths = c(1,1), heights = c(2,2))

myColors <- c(sea = "#b8b8ff", land = "white", border = "gray90")
# pdf("~/Projects/Botocudos/Plots/Map/pre_Fig1_Americas.pdf",
#     width = 10, height = 10)
map(database = "world", interior = T, boundary = F,
    xlim = c(-180, -20),
    ylim = c(-90, 90),
    bg = as.character(myColors["sea"]),
    fill = T, 
    col = as.character(myColors["land"]),
    border = as.character(myColors["border"])
    )

coords_brazil <- c(x1 = -80, x2 = -30,
                   y1 = -40, y2 = 10)

rect(xleft = coords_brazil["x1"],
     xright = coords_brazil["x2"],
     ybottom = coords_brazil["y1"],
     ytop = coords_brazil["y2"],
     border = "white",
     col = "NA")
# dev.off()
```


```{r map Brazil, fig.width=10, fig.width=10}
# pdf("~/Projects/Botocudos/Plots/Map/pre_Fig1_Brazil.pdf",
#     width = 10, height = 10)
coordinates_pops <- read.table(str_glue("{prefix_data}/final_inputs/Fig_1/map_brazil_pops.coordinates"))
colnames(coordinates_pops) <- c("population", "y", "x", "type")

map(database = "world", interior = T, boundary = T,
    xlim = coords_brazil[c("x1", "x2")],
    ylim = coords_brazil[c("y1", "y2")],
    mar = c(0,0,0,0),
    col = as.character(myColors["land"]),
    bg = as.character(myColors["sea"]),
    fill = T,
    border = "gray50"
    )

points(
  x = coordinates_pops$x,
  y = coordinates_pops$y,
  pch = 1,
  cex = 2
)
text(
  x = coordinates_pops$x + 1 ,
  y = coordinates_pops$y + 1 ,
  labels = coordinates_pops$population
)
# dev.off()


```


### Fig. 1B Calibrated dates

To reproduce the calibration curves, you need to use OxCal , then 1
- open a new project
- import new plot
- copy paste "plot" code from input_oxcal_sorted.txt
- adjust settings as wished
- save

```{sh, eval = F}
 # content of final_inputs/Fig_1/input_oxcal_sorted.txt
 
 Plot()
 {
  Curve("ShCal13","ShCal13.14c");
  R_Date("MN00016", 327, 24);
  R_Date("MN00045", 249, 29);
  R_Date("MN00064", 199, 41);
  R_Date("MN0009", 185, 29);
  R_Date("MN00067", 170, 21);
  R_Date("MN00056", 169, 25);
  R_Date("MN0003", 160, 26);
  R_Date("MN00013", 150, 40);
  R_Date("MN00022", 145, 30);
  R_Date("MN0008", 149, 23);
  R_Date("MN00023", 137, 25);
  R_Date("MN00068", 121, 33);
  R_Date("MN00039", 106, 30);
  R_Date("MN00010", 84, 22);
  R_Date("MN00069", 79, 39);
  R_Date("MN00346", 76, 30);
  R_Date("MN00021", 66, 50);
  R_Date("MN00118", 48, 33);
  R_Date("MN00316", 40, 30);
  R_Date("MN00119", 23, 22);
  
  R_Date("MN1943", 1080, 25);
  R_Date("Bot15", 417, 25);

  Curve("Marine13","Marine13.14c");
  Delta_R("LocalMarine",33,24);
  Mix_Curve("Mixed","ShCal13","LocalMarine",50,0);
  R_Date("MN01701", 2035, 27);
  R_Date("Bot17", 487, 25);
 };


```


## Figure 2 MDS + fastNGSadmix

```{r  function to plot MDS}

mds_plot <- function(points, dim1, dim2, index, pop_color, myLegend, xlab, ylab, 
                     xlim, text_coords, 
                     do_plot = T, do_legend = T, main = "MDS with worldwide groups",
                     cex = 4,
                     eig = NULL, cex_text = 1){
  # to use alpha()
  
  if(do_plot){
    # modify x and y axes
    xlab <- paste0(
      "Dimension ", data_to_plot$dim1, ": ", 
      percent(eig[dim1]/sum(abs(eig)), accuracy = 0.01)
    )
    ylab <- paste0(
      "Dimension ", data_to_plot$dim2, ": ", 
      percent(eig[dim2]/sum(abs(eig)), accuracy = 0.01)
    )
    # plot axes 
    plot(points[,dim1], points[,dim2],
         xlab = xlab, ylab = ylab,
         bty = "n", main = main,
         type = "n",
         xlim = xlim,
         cex.axis = cex_text,
         cex.lab = cex_text*1.1)
    title()
    grid()
    # add other points
    
    points(points[ -index , dim1 ], points[ -index, dim2 ], 
           col = as.character(points$color[ -index ]),
           bg =  alpha(as.character(points$bg[ -index ]), 0.9),
           pch = points$pch[ -index ], 
           cex = cex)
    # points of interest
    points(points[index, dim1], points[index, dim2],
           col = as.character(points$color[index]),
           bg = alpha(as.character(points$bg[index]), 0.8),
           pch = points$pch[index],
           cex = cex)
  }
  
  # Add text
  text("Remote Oceania", 
       x = 0.15, y = -0.1, 
       col = pop_color["RemoteOceania"], font = 1, adj = 0)
  text("Americas", 
       x = 0.09, y = 0.1, 
       col = pop_color["Americas"], font = 1, adj = 0)
  text("New 'Botocudo' individuals", x = 0.15, y = 0, 
       col = pop_color["Botocudos"], font = 1, adj = 0)
  text("Africa", x = -0.25, y = -0.05, 
       col = pop_color["Africa"], font = 1, adj = 0)
  text("Europe and West Asia", x = 0.05, y = 0.17, 
       col = pop_color["Europe"], font = 1, adj = 0)
  text("East and Southeast Asia", x = 0.15, y = -0.05, 
       col = pop_color["EastAsia"], font = 1, adj = 0)
  text("Central and South Asia", x = -0.1, y = 0.05, 
       col = pop_color["CentralSouthAsia"], font = 1, adj = 0)
  text("Near Oceania", x = -0.05, y = -0.08, 
       col = pop_color["NearOceania"], font = 1, adj = 0)
  text("'Botocudo' individuals (Malaspinas et al., 2014)", x = -0.15, y = -0.11, 
       col = "black", font = 1, adj = 0)
  
  # Add legend
  if(do_legend){
    
    plot(1:10, 1:10, axes = F, ylab = NA, xlab = NA, type = "n" )
    legend(1,10, legend = myLegend$legend, 
           cex = 0.9,
           pt.cex = 1.5,
           col = as.character(myLegend$color),
           pt.bg = as.character(myLegend$bg),
           pch = myLegend$pch, bty = "n",
           ncol = 1)
  }
}

```

```{r make figure 2}
#pdf_path <- "~/Projects/Botocudos/Plots/dummy_figures/MDS_fastNGSadmix_Wollstein_rmTrans.pdf"
#png_path <- "~/Projects/Botocudos/Plots/ngsadmix/ISBA9/MDS_fastNGSadmix_Wollstein_rmTrans.png"
mds_save_path <- str_glue("{prefix_data}/final_inputs/Fig_2/MDS_rmTrans_data_to_plot.Rda")
ngsadmix_save_path <- str_glue("{prefix_data}/final_inputs/Fig_2/NGSadmix_rmTrans_data_to_plot.Rda")

load(mds_save_path)
load(ngsadmix_save_path)

#------------------------------------------------------------------------------#
  # MDS and fastNGSadmix results
  #pdf(pdf_path, width = 7, height = 3.5, useDingbats = F)
  #png(png_path, width = round(35 / 2.54), height = round(20 / 2.54), units = "in", res = 300)
  
  
  # Matrix to indicate the position of the eight plots 
  # (one MDS, and the admixture plot is done in several parts)
  mat <- matrix(
    c(
      1,2,
      1,3,
      1,4,
      1,5,
      1,6,
      1,7
    ),
    byrow = T, nrow = 6)
  #print(mat)
  layout(mat, widths = c(5,3), heights = c(0.8,0.3,0.3,.3*6,.4,5))
  

  #-----------------------------------------------------------
  #  MDS
  par(mar = c(bottom=4, left=4, top=2, right=0))
  do.call( mds_plot,  data_to_plot )

  # empty plot
  par(mar = c(0,0,0,0))
  plot(1,1,type="n", axes =F)
  #-----------------------------------------------------------
  # barplot/admixture
  mar_bars  = c(bottom=.5, left=0, top=.1,right=8)
  
  # This plot is done in several sections to help with readabilitymanage the space
  # One individual (Sambaqui, this study)
  object_to_plot$nInds_per_block <- 1
  object_to_plot$pop_breaks <- c("Sambaqui", "Sambaqui")
  object_to_plot$plot(do_layout = F, add_text = F, horiz = T, mar_bars = mar_bars, srt = 0, adj = 0)
  
  # One individual (Mummified individual from Minas Gerais, this study)
  object_to_plot$nInds_per_block <- 1
  object_to_plot$pop_breaks <- c("MinasGerais", "MinasGerais")
  object_to_plot$plot(do_layout = F, add_text = F, horiz = T, mar_bars = mar_bars, srt = 0, adj = 0)
  
  # Block with 24 individuals labeled as "Botocudos"
  object_to_plot$pop_breaks <- c("Botocudos", "Botocudos")
  object_to_plot$nInds_per_block <- 24
  object_to_plot$plot(do_layout = F, add_text = F, horiz = T,  mar_bars = mar_bars, srt = 0, adj = 0)
  
  # The two individuals from Malaspinas et al., 2014
  object_to_plot$nInds_per_block <- 2
  object_to_plot$pop_breaks <- c("Botocudos2014", "Botocudos2014")
  object_to_plot$plot(do_layout = F, add_text = F, horiz = T,  mar_bars = mar_bars, srt = 0, adj = 0)
  
  object_to_plot$qopt <- object_to_plot$qopt[
    -which(object_to_plot$new_panel$population == "Botocudos2014"),
  ]
  object_to_plot$new_panel <- 
    object_to_plot$new_panel[
      -which(object_to_plot$new_panel$population == "Botocudos2014"),
    ]
  object_to_plot$myAes <- object_to_plot$myAes[
    -which(object_to_plot$myAes$population == "Botocudos2014"),
  ]
  
  object_to_plot$pop_breaks <- c("YRI", "Totonac")
  
  object_to_plot$plot(do_layout = F, add_text = F, horiz = T,  mar_bars = mar_bars, srt = 0, adj = 0)
  
```

In this part of the code, there were two R objects that were important to load, one stored in `MDS_rmTrans_data_to_plot.Rda` and another one in `NGSadmix_rmTrans_data_to_plot.Rda`.

After loading `MDS_rmTrans_data_to_plot.Rda`, an object called `data_to_plot` will appear in the environment.

`data_to_plot` is a list, and one of the items in the list is a dataframe called `points`.
The rows in `points` correspond to the individuals, and the first 606 columns correspond to each dimension of the MDS.
The last columns have the metadata of the individuals, which is useful to make queries and to make the plot.

```{r}
dim(data_to_plot$points)
head(data_to_plot$points[,608:617])
```


Once `NGSadmix_rmTrans_data_to_plot.Rda` is loaded, an object called `object_to_plot` will be found in the environment.
This is an object I defined to help me plotting and modifying admixture plots as needed. However, it might not be easy to use.
If anyone needs to play with this object, I recommend using its `plot()` function right after loading it, as this will initialize many useful variables.
It will also print a default admixture plot with all the data.

The admixture proportions are stored in its `qopt` dataframe. This dataframe has an extra column (`width`) which helps to indicate the desired width for the genome of an individual in a plot.

The metadata (individuals' IDs, populations, etc.) is stored in the `new_panel` dataframe.
The color of each component can be adjusted by modifying the vector `colors`.


```{r}
load(ngsadmix_save_path)
# object_to_plot$plot()
head(object_to_plot$qopt)
head(object_to_plot$new_panel)
object_to_plot$colors

```


## Figure 3: F3 
This is a large figure, containing an F3 plot, and NGSadmix plot, and two qpGraph plots.

### Fig. 3A F3

```{r F3 main figure, fig.width=12, fig.height=3.2}

boto_f3 <- read.csv( str_glue("{prefix_data}/final_inputs/Fig_3/f3_metadata_plot.csv"))
boto_f3$Source_2 <- factor(boto_f3$Source_2, levels = boto_f3$Source_2, ordered = T)

boto_f3$disp_blocks <- cut_number(boto_f3$f_3, n = 5)
boto_f3$disp_blocks <- factor(
  boto_f3$disp_blocks, 
  levels = rev(levels(boto_f3$disp_blocks)), 
  ordered = T
  )



#pdf("~/Projects/Botocudos/Plots/F3/2021_06/ggplot_F3.pdf", width = 8.6, height = 3.2)
ggplot(boto_f3, aes(x = f_3, y = Source_2, color = Source_2, 
                    xmin = f_3 - 3*std.err,
                    xmax = f_3 + 3*std.err)) +
  geom_point(size = 0.8) +
  geom_errorbar(size = 0.34, width = 0) +
  scale_color_manual(values = boto_f3$color) +
  facet_wrap(~disp_blocks, nrow = 1, scales = "free_y", as.table = T) +
  theme_light() +
  theme(legend.position = "none",
        axis.text = element_text(size = 5),
        strip.text.x = element_blank()) +
  labs(x = "F3", y = NULL, title = NULL)

```

The dataframe `boto_f3` stores the output of admixtools. A few extra columns were added, such as the geographic coordinates for the plot, whether the population is ancient, and the color of the dots.


```{r}
head(boto_f3)
```


### Fig. 3B NGSAdmix (Americas' panels)


```{r NGSadmix Figure 3, fig.width=14, fig.height=11}

load(str_glue("{prefix_data}/final_inputs/Fig_3/object_to_plot_ngsadmix.Rda"))
object_to_plot$plot(do_layout = T, add_text = T)



#str(object_to_plot)
```


As noted for Figure 2, we are loading an object called `object_to_plot`. 
The dataframe `qopt` contains the estimated admixture proportions, and the data for each individual (ID, population, etc.) can be found in the `new_panel` dataframe.

With this object, you can also check which population has the highest average admixture proportions for each of the components.

```{r}
# load(str_glue("{prefix_data}/final_inputs/Fig_3/object_to_plot_ngsadmix.Rda"))

# object_to_plot$plot()

head(object_to_plot$qopt)
head(object_to_plot$new_panel)
head(object_to_plot$colors)

# populations with the highest average component
print(object_to_plot$max_anc_pop)

```


### Fig. 3C and 3D qpGraph





## Figure 4 Conditional heterozygosity, PSMC, ROH

```{r,  fig.width = 12, fig.height = 8}
#------------------
# Conditional heterozygosity
pop_color <- c(
  "#5c2c45",
  "#ffb852",
  "#ffa6d9",
  "#e81900",
  "#65A98F",
  "#29bdad",  
  "#b319ab",
  "#9C52F2",
  "black"
)

names(pop_color) <- c("Africa",
                      "Europe",
                      "CentralSouthAsia",
                      "EastAsia",
                      "NearOceania",
                      "RemoteOceania",
                      "Americas",
                      "Ancient_Americas",
                      "Botocudos")

main_size <- 12
legend_size <- 8

selected_pi <- read.csv(
  str_glue("{prefix_data}/final_inputs/Fig_4/conditional_het_plot.csv"), 
  stringsAsFactors = F
  )


selected_pi$Region <- factor(
  selected_pi$Region, 
  levels = c(
    "Africa", "WestEurasia", "EastAsia", "Oceania", "Americas_Ancient", 
    "Americas", "Botocudos"
    ), 
  ordered = T
  )

selected_pi <- selected_pi[order(selected_pi$Region, selected_pi$estimate), ]
selected_pi$Population.ID <- factor(selected_pi$Population.ID, levels = rev(selected_pi$Population.ID), ordered = T)

condHet_plot <- ggplot(selected_pi, aes(x = Population.ID, y = estimate, ymin = confint1, 
               ymax = confint2, color = Region, interaction = Region)) +
  geom_point(size = 1.2) +
  geom_errorbar(lwd = .6, width = 0.5) +
  theme_minimal() +
  guides(color=guide_legend(nrow=2)) +
  theme(axis.text.x = element_text(angle = 90, size = 9, hjust = 1, vjust = 0.5),
        legend.position = "none", 
        legend.text = element_text(size = legend_size),
        title = element_text(size = main_size),
        strip.text = element_text(size=7, margin = margin(.1, 0, .1, 0, "cm"))
        ) + 
  labs(title = "Conditional heterozygosity", 
       x = NULL, y = "Estimate") +
  scale_color_manual(
    values = as.character(
      pop_color[rev(c("Botocudos", "Ancient_Americas", 
                  "Americas",
                  "NearOceania", "EastAsia", "Europe",
                                          "Africa"))])
    ) +
  coord_trans(y = "log") +
  facet_grid(.~Region, scales = "free", space = "free")


#-----------------
# ROH
pop_color <- c(
  "#5c2c45",
  "#ffb852",
  "#e81900",
  "#65A98F",
  "#b319ab",
  "#9C52F2",
  "black"
)

names(pop_color) <- c("Africa",
                      "Europe",
                      "EastAsia",
                      "Oceania",
                      "Americas",
                      "Americas_Ancient",
                      "Botocudos")

lengths_all <- read.csv(str_glue("{prefix_data}/final_inputs/Fig_4/roh_plot_main_fig.csv"))

lengths_all <- lengths_all[lengths_all$length>0,]

lengths_all$sample <- factor(lengths_all$sample, levels = unique(lengths_all$sample), ordered = T)
region_levels <- unique(lengths_all$region)
lengths_all$region <- factor(lengths_all$region, levels = region_levels, ordered = T)


breaks_mb <- c(0, 0.5, seq(1, 32))
breaks_labels <- c("", 0.5,  seq(5, 32, 5))
xlim <-  c(0, 65)
ylim <- c(0, 850)

main <-  "ROH distribution"
xlab <-  "ROH length category (Mb)"
ylab <- "Sum (Mb)"

x <- "category" ; y <- "length"


roh_plot <- ggplot(lengths_all, aes( x = as.factor(lengths_all[, x]), y = lengths_all[, y] , 
                         fill = region)) +
  geom_bar(stat = "identity", position = position_dodge(), color = "white") +
  labs(x = xlab, y = ylab, title = main) +
  coord_cartesian(xlim = c(0, sum(breaks_mb <= xlim[2])),
                  ylim = c(1, 1000)) + 
  scale_x_discrete(breaks = breaks_mb[ breaks_mb <= xlim[2]], 
                   labels = sapply(breaks_mb, function(x) ifelse(x %in% breaks_labels, x, "")),)  +
  scale_y_log10() +
  facet_wrap(sample~., ncol = 2
             ) +
  theme_minimal() +
  theme(legend.position = "none", 
        axis.text.x = element_text(size = 8, angle = 90, hjust = 1, vjust = 0)
        ) +
  scale_fill_manual(values = (pop_color[region_levels]))

#roh_plot 
#-----------------
# PSMC

psmc <- read.csv(str_glue("{prefix_data}/final_inputs/Fig_4/psmc_plot.csv"))

psmc_plot <- ggplot(psmc, aes(x = time+1, y = popsize*10e4, color = sample)) +
  geom_step(stat="identity", size = 1.) +
  coord_cartesian(xlim = c(9000, 10e5)) +
  scale_x_log10(labels = scales::comma,
                minor_breaks = as.numeric(sapply(seq(3, 7), function(i) seq(1,10)*10**i))) +
  scale_y_log10(labels = scales::comma,
                minor_breaks = as.numeric(sapply(seq(3, 6), function(i) seq(1,10)*10**i))) +
  theme_minimal() +
  scale_color_manual(
    name = "Genome",
    values = alpha(c("black","#FF616B", "#A7374B", "#b319ab","#99B333", "#29BDAD"), 0.8),
    breaks = c( "MN0008_L3U_trim2",
               "SuruiA",
               "Karitiana_HGDP_TeamAB",
               "MayaG",
               "A_Dai",
               "A_Yoruba"),
    labels = c("MN0008 (Botocudo)",
               "A_Surui (Americas)",
               "A_Karitiana (Americas)",
               "A_Maya (Americas)",
               "A_Dai (East Asia)",
               "A_Yoruba (Africa)")) +
  guides(color=guide_legend(nrow=2)) +
  theme(legend.position = "top", #c(0.1, 0.7),
        legend.text = element_text(size = legend_size),
        title = element_text(size = main_size)) +
  labs(title = "Effective population size",
       x = "Years ago",
       y = expression(N[e]))

#psmc_plot

col1_x <- 0
col2_x <- 0.6
row1_y <- 0.6
row2_y <- 0

#pdf("~/Projects/Botocudos/Plots/dummy_figures/Fig4_condHet_ROH_PSMC.pdf",
#    width = 12, height = 8)
ggdraw() +
  draw_plot(condHet_plot, x = col1_x, y = row1_y, width = col2_x , height = 1 - row1_y) +
  draw_plot(roh_plot, col2_x, row2_y, 1 - col2_x, 1 - row2_y) +
  draw_plot(psmc_plot, col1_x, row2_y, col2_x, row1_y - 0.05) +
  draw_plot_label(c("A", "B", "C"), c(col1_x, col2_x, col1_x), c(1, 1, row1_y - 0.05), size = 15)

#dev.off()


```

```{r}
head(selected_pi)

head(lengths_all)

head(psmc)

```

## Figure 5 Demography

This figure was modified on Adobe Illustrator.

```{r, fig.width=14, fig.height=9}
plot_name <- "twelve_models"

#png(str_glue(png_path), width = 28, height = 18, res =300, units = "in")

par(mar = c(8, 6, 2, 6))

pop_names_colors <- c("yellow", "gray", "gray", "gray", "gray", "blue", "blue", "red")
names(pop_names_colors) <- c(
  "Yoruba",  "ANC",  "Out of Africa", "Americas",
  "South America", "Karitiana", "Surui", "Botocudos"  
)

load(
  str_glue("{prefix_data}/final_inputs/Fig_5/model_values.Rda")
)

make_demography_plot(
  pop_names_colors = pop_names_colors, do_log = T, max_width = 0.8,
  pop_new_order = c(
    "Yoruba",  "ANC",  "Out of Africa", "Americas",
    "South America", "Karitiana", "Surui", "Botocudos"  
  ),
  pop_old_order = c("Botocudos", "Karitiana", "Surui", "Americas", 
                    "Out of Africa", "ANC", "Yoruba",  "South America"),
  bottleneck_height = 0.01,
  alpha = 1, main = "",
  ... = model_values
)

#dev.off()
```


## Figure 6 Dstats

```{r, read files and set variables}

abba <- read.table(
  str_glue("{prefix_data}/final_inputs/Fig_6/July_09_trim5.ErrorCorr.TransRem.txt"),
  header = T, stringsAsFactors = F
  )

pops_in_title <- c(
  "Mixe",
  "Surui",
  "Karitiana",
  "MN0008",
  "MN0008_trim5",
  "MN0008_L3U",
  "MN0008_trim5",
  "MN0008_L3U_trim2",
  "MN0008_L3U_trim5",
  "22Botocudos_trim5",
  "LagoaSanta",
  "LagoaSanta_trim5",
  "Huichol",
  "Aymara"
)

pops_in_y <-  c(
  "Mixe",
  "Huichol",
  "Aymara",
  "Surui",
  "Karitiana",
  "MN0008_L3U_trim2",
  "22Botocudos_trim5",
  "LagoaSanta",
  "LagoaSanta_trim5"
                   )
pops_in_h3 <- c(
  "French", 
  "Han",
  "Andaman_trim5",
  "Papuan",
  "Australian"
  )

pop_colors <- c("deepskyblue1", "darkgoldenrod1", 
                "pink",  "brown3", "blueviolet",
                "black")
```

```{r try to get a main figure, fig.width = 9.5, fig.height = 7.5}

pops_in_title = c("Mixe")
pops_in_h3 <- c(
  "French", 
  "Han",
  "Andaman_trim5",
  "Papuan",
  "Australian"
  )
pops_in_h3_labels <-  c("French\n(n = 1)",
                                 "Han\n(n = 1)",
                                 "Andaman\n(n = 1)",
                                 "Papuan\n(n = 1)",
                                 "Australian\n(n = 7)")

pop_colors <- c("deepskyblue1", "darkgoldenrod1", 
                "pink",  "brown3", "blueviolet",
                "black")
pop_colors <- c("#0057BA", "#96BFE6", "#CC85D1", "#DE4500", "#9E194D", "black")
pop_colors <- c("gray60", "gray80", "#CC85D1", "#96BFE6", "#0057BA", "black")

pops_in_y <-  c(
  "Surui",
  "LagoaSanta",
  "LagoaSanta_trim5",
  "LagoaSanta",
  "MN0008_L3U_trim2",
  "Karitiana",
  "Aymara",
  "Huichol"
  )

new_labels = list(
  MN0008_L3U_trim2 = "MN0008\n(n = 1)",
  Surui = "Surui\n(n = 2)",
  LagoaSanta_trim5 = "LagoaSanta\n(n = 1)",
  Karitiana = "Karitiana\n(n = 4)",
  Aymara = "Aymara\n(n = 1)",
  Huichol = "Huichol\n(n = 1)"
)

pop <- "Mixe"
outgroup <- "Yoruba"
Dstat <- abba[(abba$H1 %in% pops_in_title |abba$H2 %in% pops_in_title) &
                (abba$H3 %in% pops_in_h3),]
df_y <- as.data.frame(pops_in_y)
df_y$y_axis <- rownames(df_y)
plot_list <- list()

  Dstat <- abba[(abba$H1 == pop |abba$H2 == pop) &
                  (abba$H3 %in% pops_in_h3),]
  Dstat <- switch_h2_h1(Dstat, pop)
  Dstat <- Dstat[!(Dstat$H1 %in% pops_in_h3),]
  
  
  Dstat <- Dstat[Dstat$H1 %in% pops_in_y,]
  
  for(label in names(new_labels)){
    Dstat[Dstat$H1 == label,]$H1 <- new_labels[[label]]
  }
  
  
  Dstat$H1 <- factor(Dstat$H1, 
                     levels = rev(unique(new_labels))
                     )
  Dstat$H3 <- factor(Dstat$H3, 
                     levels = c(pops_in_h3),
                     ordered = T)
  
  
  
  x_label <- expression(
    atop(
      "Z",
      paste("(", H[1], ", ", H[3], ") < --- > (Mixe, ", H[3], ")"),
      paste("(Mixe, Yoruba)",  " < --- > (", H[1], ", Yoruba)")
    )
    )
  
  
  
    
  p <- ggplot(Dstat, aes(x = Z, y = H1, 
                         fill = H3)) +
    
    coord_cartesian(xlim = c(-4,3.5)) +
    scale_fill_manual(values = alpha(pop_colors, .8),
                       breaks = pops_in_h3,
                      labels = pops_in_h3_labels,
                      name = expression(paste(H[3], ":"))) +
    geom_vline(xintercept = c(-3.3,3.3), lty = "dashed", col = "brown", lwd = 1)+
    geom_vline(xintercept = c(-4,-3,-2,-1,0,1,2,3), lty = "solid", col = "gray", lwd = 0.2) +
    geom_vline(xintercept = c(-3.5,-2.5,-1.5,-0.5,0.5,1.5,2.5,3.5), lty = "dashed", col = "gray", lwd = 0.2) +
    
    geom_point(shape = 23, size = 7, stroke= 1) +
    labs(
      x = x_label,
      y = expression(H[1]),
      title = expression(bold(paste("Z-scores for D(", H[1], ", ", H[2]," = Mixe; ", H[3], ", Yoruba)" )))) +
    theme_cowplot() +
    theme(legend.position = "top", 
          axis.text.x = element_text(size = 16),
          axis.text.y = element_text(size = 16)
          ) 
  
#pdf("~/Projects/Botocudos/Plots/dummy_figures/Dstat.pdf", width = 9.5, height = 7.5)
  p
#dev.off()


```

## Figure 7 Virome
Done

```{r paths and variables}


prot_names <- c(
  "11-kDa", "11 kDa protein",
  "7.5-kDa", "7.5 kDa protein",
  "hypothetical 11kDa protein", "hypothetical protein", 
  "ns", "NS",
  "ns1", "NS1",
  "vp1", "VP1", 
  "VP1/2", "vp2", 
  "VP2", "VP3",
  "protein X",
  "X", "X-9 kDa protein"
)

prot_colors <- c(
  rep("orange", 2),    # 11 kDa
  rep("plum" , 2),     # 7.5 kDa
  rep("#172713",  2),  # hypot 11 kDa
  rep("#172713", 4),   # NS1
  rep("#ff616b", 2),   # VP1
  "#de4500" ,        # Vp1/2
  rep("#328e13", 2),   # VP2
  "plum4",             # VP3
  rep("seagreen", 3)   # X
)

names(prot_colors) <- prot_names


```

```{r functions to plot}


plot_bamdamage <- function(five_prime, three_prime, titlecolor = "black", cex = 1, 
                           cexmain = 2, txt_reads = F,
                           y2 = 0.55, plot.title, plot.title2 = "", myColor = c(),
                           legend_only = F, ncol_legend = 3, legend_cex = 1){
  
  if (length(myColor) == 0) {
    myColor <- c("royalblue", "firebrick3", "lightblue3", "indianred1",
                 "orange","orange3", "plum", "plum4", "paleturquoise3", 
                 "paleturquoise4", "rosybrown2", "rosybrown3")
    
    names(myColor) <- c("C..T", "G..A", "T..C", "A..G", "C..A", "C..G",
                        "T..A", "T..G", "A..C", "A..T", "G..C", "G..T")
  }
  
  if (legend_only) {
    myIndex <- round(seq(1, 13, length.out = ncol_legend + 1))
    for (i in 1:(length(myIndex) - 1)) {
      plot(1:10, 1:10, type = "n", bty = "n", axes = F, xlab = NA, ylab = NA)
      
      if (i == 1) {
        legend(1, 10, legend = sub("\\.\\.", " to ",
                                   names(myColor)[myIndex[i]:(myIndex[i + 1] - 1)]),
               col = myColor[myIndex[i]:(myIndex[i + 1] - 1)],
               lty = 1, bty = "n", cex = legend_cex, lwd = 1.6,
               title = "Substitution type")
      }else{
        legend(1, 10, legend = sub("\\.\\.", " to ", names(myColor)[myIndex[i]:(myIndex[i + 1] - 1)]),
               col = myColor[myIndex[i]:(myIndex[i + 1] - 1)],
               lty = 1, bty = "n", cex = legend_cex, lwd = 1.6
        )
      }
    }
    return()
  }
  
  plot_lines <- function(damage, left = T){
    plot(damage$C..T[1:25], type = "n",
         bty = "n", 
        ylab = "Frequency", xlab = NA,
         cex = cex, axes = F, ylim = c(0, y2))
    
    if (left) {
      title(main = plot.title, xlab = "5'-position", cex.main = cexmain)
      axis(side = 1, at = seq(0, 25, 5))
      axis(side = 2)
    }else{
      axis(side = 1, at = seq(1, 25, 5), labels = seq(25, 1, -5))
      axis(side = 4)
      title(main = plot.title2, xlab = "3'-position", cex.main = cexmain, 
            col.main = titlecolor)
      nb_reads_legend <- paste0("Number of reads: ", txt_reads)
      text(x = 12, y = 0.51, labels = nb_reads_legend, cex = 1.15)
    }
    
    for (type in names(myColor)) {
      lines(damage[1:25, type, with = F],
            col = alpha(myColor[type], 0.8), 
            lwd = 1.8, 
            cex = cex)
    }
  }
  
  par(mar = c(4, 4, 3, 1))
  plot_lines(damage = five_prime)
  
  par(mar = c(4, 1, 3, 3))
  plot_lines(damage = three_prime[25:1, ], left = F)

}

blank_plot <- function(){
  par(mar = c(0,0,0,0))
  plot(i, i, type = "n", bty = "n", axes = F, xlab = "", ylab = "", main = "")
}


draw_coverage <- function(coverage, plot.title, myColor,
                          limy){
      barplot(height = coverage, 
            main = plot.title, 
            ylab = "Number of reads",
            col = myColor,
            border = NA,
            axes = T,
            ylim = c(0, limy), 
            space = 0,
            cex.main = 1.5)
    # x axis
    xtick <- round(seq(1, 5000, length.out = 6))
    axis(side = 1, at = xtick, labels = FALSE, tck = -0.01)
    text(x = xtick,  par("usr")[3], labels = paste0(xtick, "bp"), pos = 1, xpd = TRUE, 
         cex = 1)
}

draw_protein_boxes <- function(prots_coords, boxColors, limy, boxes_only = F){
  #browser()
  # Rectangles part
  max_coverage = limy
  
  if(boxes_only){
    limy = -max_coverage*5
    barplot(height = limy,
            axes = F,
            col = NA,
            border = NA, 
            ylim = c(0,max_coverage),
            xlim = c(1,5017)
    )
  }
  space_btwn_box <- -limy/20
  box_width <- -limy/20
  box_top <- box_width
  
  for (i in c(1:length(prots_coords$start))) {
    if (prots_coords[i, Name] %in% c("ns", "NS", "ns1", "NS1", "vp1", "VP1", "VP1/2")) {
      box_top <- box_width
    } else if (prots_coords[i, Name] %in% c("VP2", "vp2")) {
      box_top <- box_top + space_btwn_box 
    } else {
      box_top <- 2 * box_top + space_btwn_box
    }
    
    rect(xleft = prots_coords[i, start],
         xright = prots_coords[i, end],
         ytop = box_top,
         ybottom = box_top + box_width,
         col = boxColors[prots_coords[i, Name]],
         border = NA)
    
    box_top <- box_width
  }
}



plot_cov <- function(coverage, plot.title, myColor, boxColors, prots_coords, limy = limy,
                     print_coverage = T, print_boxes = T, boxes_only = F){
  par(xpd = T)
  
  if(print_coverage){
    draw_coverage(coverage, plot.title, myColor, limy)
  }
  
  if(print_boxes){
    draw_protein_boxes(prots_coords, boxColors, limy, boxes_only = boxes_only)
  }
  
  
}



```

```{r coverage_mn00346, fig.width=14, fig.height=10}


for(type in c("Cov", "Len", "Dam")){
  load(file = str_glue("{prefix_data}/final_inputs/Fig_7/my{type}.Rda"))
  
}

my_mat <- matrix(c(15,15,1,2,
                   15,15,1,3,
                   15,15,1,4,
                   15,15,1,5,
                   15,15,1,6,
                   11,12,7,8,
                   13,13,14,14,
                   9,10,14,14)
                  , 8, 4, byrow = T)
print(my_mat)


# pdf("~/Projects/Botocudos/Plots/Viruses/partial_fig_8.pdf", width = 5*4, height = 12)
nf <- layout(my_mat, widths = c(1.5,1.5,2,1.5), heights = c(rep(3, 5), 1.5, 9, 9))

# layout.show(nf)
#-----------------------------------------------------------------------------#
# Coverage plots
  
botocudos <-  c("MN00346", "MN00067", "MN00021", "MN00013", "MN00119", "MN00039")
for(boto in botocudos){
  if(boto != "MN00346"){
    par(mar = c(1.5,4,1.5,1), cex = 1.)
  }else{
    par(mar = c(1.5,4,4,2), cex = 1)
  }
  plot_cov(coverage = myCov[[boto]]$coverage$reads, boxColors = prot_colors, 
           prots_coords = myCov[[boto]]$prots, plot.title = boto, 
           myColor = "#9e194d",
           limy = max( myCov[[boto]]$coverage$reads), print_boxes = F)
}

par(mar = c(0,4,0,3))
for(i in c(1,2)){
  plot_cov(coverage = "", boxColors = prot_colors, 
           prots_coords = myCov[[boto]]$prots, plot.title = NA, 
           myColor = "#9e194d",
           print_coverage = F,
           print_boxes = T, boxes_only = T, limy = 50)
}


 #-----------------------------------------------------------------------------#
# Damage
boto <- "MN00346"
y2 <- 0.45
nreads <- scan(str_glue("{prefix_data}/final_inputs/Fig_7/MN00346.AY083234_1_final_reads.txt"))

plot.title <- boto
plot.title2 <- ""

five_prime <- myDam[[boto]][["five_prime"]]
three_prime <- myDam[[boto]][["three_prime"]]
plot_bamdamage(five_prime = five_prime, three_prime = three_prime,
               cex = 1, cexmain = 1.65, titlecolor = "black",
               txt_reads = nreads, y2 = y2, plot.title = plot.title, 
               plot.title2 = plot.title2, myColor = c())

#-----------------------------------------------------------------------------#
# 
blank_plot()
blank_plot()

#-----------------------------------------------------------------------------#
# read length
par(mar = c(5,4,1,0))
barplot(myLen$MN00346$counts,
        border = F,
        main = "Reads mapped to HPV B19 (MN00346)",
        col = "#ff616b",
        xlab = "Length (bp)",
        ylab = "Counts",
        xlim = c(10, 78))
axis(1, at = seq(10, 60, 10)*1.2+.5, labels = seq(30, 80, 10))

plot_bamdamage(legend_only = T, ncol_legend = 1)

# Heatmap

plot.new()             
vps <- baseViewports()
pushViewport(vps$figure) 

#------------------------------------------------------------------------------#
# heatmap

boc_wide <- read.csv(
  str_glue("{prefix_data}/final_inputs/Fig_7/breadth_coverage_for_plot.csv"), 
  row.names = 1
  )

samples_groups <- sapply(rownames(boc_wide), function(ind) ifelse(ind %in% c("MN1943", "MN01701"), "Pre-Contact", "Post-Contact") )
samples_groups <- factor(samples_groups, levels = c("Pre-Contact", "Post-Contact"), ordered = T)

index <- sort(unique(unlist(sapply(boc_wide, function(i) unique(round(i*100))))))

colors100  = sequential_hcl(max(index), palette = "SunsetDark")

colors = unlist(sapply(index, function(i)  colors100[i]))
colors = rev(colors)
colors[1] <- "white"

ht = Heatmap(t(as.matrix(boc_wide)), 
        name = "Fraction of the genome covered", #title of legend
        col = colors,
        show_row_dend = F,
        show_column_dend = F,
        column_title_side =  "bottom", 
        column_split = samples_groups,
        column_gap = unit(5, "mm"),
        rect_gp = gpar(col = "gray", lty = "dotted"),
        border = "black",
        row_names_gp = gpar(fontsize = 12), # Text size for row names
        row_names_side = c("right"),
        heatmap_legend_param = list(title = "Fraction of the genome covered",
                                    direction = "horizontal"),
        bottom_annotation = HeatmapAnnotation(
          df = data.frame("Period" = samples_groups),
          col = list("Period" = c("Pre-Contact" = "#0057ba",
                                "Post-Contact" = "#d60036"))
          ),
        width  = unit(10, "cm"),
        height = unit(6, "cm")
        )

vp1 <-plotViewport(c(1.8,1,0,1)) ## create new vp with margins, you play with this values 
par(mar = c(5,5,5,5))
p <- draw(ht, column_title = "Fraction of the genome covered at least once (viruses on the top 10%)",
     heatmap_legend_side = "bottom", annotation_legend_side = "bottom", 
     merge_legend = T, newpage = F
         )

#print(p, vp = vp1)


#dev.off()

```


In this part, several objects were loaded (`myCov`, `myLen`, `myDam`).

`myCov` is a list, with each element corresponding to an individual.
For each individual, you will find a dataframe indicating the number of reads observed at each position of the parvovirus genome.
There is also a dataframe with the coordinates of the proteins.

In `myLen`, each item of the list corresponds to an individual. 
For each individual, there is a vector corresponding to read lengths, and another vector counting the occurrences.

The `myDam` object includes, per individual, a dataframe indicating the average frequency of substitutions along the reads, counted from the 5' or 3' end.


```{r}

#str(myCov)

head(myCov[["MN00346"]]$coverage)
head(myCov[["MN00346"]]$prots)

#str(myLen)
head(myLen$MN00346$length)
head(myLen$MN00346$counts)

#str(myDam)
head(myDam$MN00346$five_prime)
```





# Supplement


## Fig. S1 Damage per sample

```{r plot_bamdamage function}
plot_bamdamage <- function(five_prime, three_prime, titlecolor = "black", cex = 1, 
                           cexmain = 2, txt_reads = F,
                           y2 = 0.55, plot.title, plot.title2 = "", myColor = c(),
                           legend_only = F, ncol_legend = 3, legend_cex = 1){
  
  if (length(myColor) == 0) {
    myColor <- c("royalblue", "firebrick3", "lightblue3", "indianred1",
                 "orange","orange3", "plum", "plum4", "paleturquoise3", 
                 "paleturquoise4", "rosybrown2", "rosybrown3")
    
    names(myColor) <- c("C..T", "G..A", "T..C", "A..G", "C..A", "C..G",
                        "T..A", "T..G", "A..C", "A..T", "G..C", "G..T")
  }
  
  if (legend_only) {
    myIndex <- round(seq(1, 13, length.out = ncol_legend + 1))
    for (i in 1:(length(myIndex) - 1)) {
      plot(1:10, 1:10, type = "n", bty = "n", axes = F, xlab = NA, ylab = NA)
      
      if (i == 1) {
        legend(1, 10, legend = sub("\\.\\.", " to ",
                                   names(myColor)[myIndex[i]:(myIndex[i + 1] - 1)]),
               col = myColor[myIndex[i]:(myIndex[i + 1] - 1)],
               lty = 1, bty = "n", cex = legend_cex, lwd = 1.6,
               title = "Substitution type")
      }else{
        legend(1, 10, legend = sub("\\.\\.", " to ", names(myColor)[myIndex[i]:(myIndex[i + 1] - 1)]),
               col = myColor[myIndex[i]:(myIndex[i + 1] - 1)],
               lty = 1, bty = "n", cex = legend_cex, lwd = 1.6
        )
      }
    }
    return()
  }
  
  plot_lines <- function(damage, left = T){
    plot(damage$C..T[1:25], type = "n",
         bty = "n", 
        ylab = "Frequency", xlab = NA,
         cex = cex, axes = F, ylim = c(0, y2))
    
    if (left) {
      title(main = plot.title, xlab = "5'-position", cex.main = cexmain)
      axis(side = 1, at = seq(0, 25, 5))
      axis(side = 2)
    }else{
      axis(side = 1, at = seq(1, 25, 5), labels = seq(25, 1, -5))
      axis(side = 4)
      title(main = plot.title2, xlab = "3'-position", cex.main = cexmain, 
            col.main = titlecolor)
      nb_reads_legend <- paste0("Number of reads: ", txt_reads)
      text(x = 12, y = 0.51, labels = nb_reads_legend, cex = 1.15)
    }
    
    for (type in names(myColor)) {
      lines(damage[1:25, type], # with = F],
            col = alpha(myColor[type], 0.8), 
            lwd = 1.8, 
            cex = cex)
    }
  }
  
  par(mar = c(4, 4, 3, 1))
  plot_lines(damage = five_prime)
  
  par(mar = c(4, 1, 3, 3))
  plot_lines(damage = three_prime[25:1, ], left = F)

}

```

```{r plot damage per sample, eval=T, fig.width=16, fig.height=16}
nCol <- 8
nRow <- 7
cex <- 1.2

blank_plot <- function(){
  plot(1:10,1:10, type = "n", bty = "n", axes = F, xlab = NA, ylab = NA)
}

# pdf(paste("~/Projects/Botocudos/Plots/Sequencing/Damage_", data_type, ".pdf", sep = ""),
#     width = 14, height = 12)

load(str_glue("{prefix_data}/final_inputs/Supplement/Damage/damage_samples.Rda"))

samples <- names(myDam)

layout(matrix(seq(nCol*nRow), ncol = nCol, byrow = T))
for(s in samples){
    y2 <- 0.35
    plot.title <- s
    five_prime <- myDam[[s]][["five_prime"]]
    three_prime <- myDam[[s]][["three_prime"]]
    plot_bamdamage(five_prime = five_prime, three_prime = three_prime, 
                  cex = 1, 
                  y2 = y2, plot.title = plot.title,
                  myColor = c())
}
plot_bamdamage(legend_only = T,
                  myColor = c(),
               legend_cex = 1, ncol_legend = 3)

# dev.off()
```


## Dataset S3 Damage per library

The plots below were included as a dataset (a PDF file with multiple pages).

```{r plot damage per library, eval=T, fig.width=12, fig.height=12}
load(str_glue("{prefix_data}/final_inputs/Supplement/Damage/damage_libraries.Rda"))

lib_stats <- read.csv(str_glue("{prefix_data}/final_inputs/Supplement/Length/library_stats.hg19.csv"))

samples <- names(myDam)
samples <- samples[! samples %in% c("MN0008_non_U", "MN0008_L3U")]

nCol <- max(sapply(myDam, length))
samples_per_page <- 5
cex <- 1.2

blank_plot <- function(){
  plot(1:10,1:10, type = "n", bty = "n", axes = F, xlab = NA, ylab = NA)
}

# pdf(paste("~/Projects/Botocudos/Plots/Sequencing/Damage_lib_", data_type, ".pdf", sep = ""),
#     width = 12, height = 12)
layout(matrix(seq(samples_per_page*nCol*2), nrow = samples_per_page, byrow = T))

for(s in samples){
  myLibs <- lib_stats$LB[lib_stats$SM == s]
  myRow <- which(samples == s)
  
  if(!myRow %% samples_per_page){
    # plot.new()
    layout(matrix(seq(samples_per_page*nCol*2), nrow = samples_per_page, byrow = T))  
  }
  
  
  for(l in myLibs){
    y2 <- 0.35
     plot.title <- paste(s, l)
    
    five_prime <- myDam[[s]][[l]][["five_prime"]]
    three_prime <- myDam[[s]][[l]][["three_prime"]]
    plot_bamdamage(five_prime = five_prime, three_prime = three_prime, 
                  cex = 1, 
                  y2 = y2, plot.title = plot.title,
                  myColor = c())
    
  
  }
  plots_left <- nCol*2 - length(myLibs)*2
  while(plots_left > 0){
    blank_plot()
    plots_left <- plots_left - 1
  }

}

plot_bamdamage(legend_only = T,
                  myColor = c(),
               legend_cex = 1, ncol_legend = 3)


# dev.off()
```


## Fig. S2 Read length per sample
```{r, fig.width=12, fig.height=12}


myColor <- "royalblue"
nCol <- 6
cex <- 1.1

blank_plot <- function(){
  plot(1,1, type = "n", bty = "n", axes = F, xlab = NA, ylab = NA)
}

# pdf(paste("~/Projects/Botocudos/Plots/Sequencing/Length_dist_",
#           data_type,".pdf", sep = ""),
    # width = 12, height = 12)
load(str_glue("{prefix_data}/final_inputs/Supplement/Length/samples_length.Rda"))
samples <- names(myLength)
samples_per_row <- ceiling(length(samples)/nCol)
layout(matrix(seq(1, nCol*samples_per_row), ncol = nCol, byrow = T))

for(s in samples){
    le <- myLength[[s]]
    plot.title <- s
    myMean <- round(sum(le$length*le$counts)/sum(le$counts), 1)
    barplot(le$counts/sum(le$counts), space = 0,
         col = myColor, bty = "n", 
         xlab = "Length", ylab = "Frequency", border = NA,
         main = plot.title, lwd = 0.2)
    axis(1, at= seq(0.5, nrow(le), 5), 
         labels = seq(min(le$length), max(le$length), 5))
    text(20, 0.9*max(le$counts/sum(le$counts)), 
         labels = paste("Mean: ", myMean, " bp", sep = ""),
        cex = 0.8)
    abline(v = myMean-20, lty = "dashed")
}
```

## Dataset S4 Read length per library
```{r plots per library, eval = T, fig.width=9, fig.height=12}
myColor <- "royalblue"

lib_stats <- read.csv(str_glue("{prefix_data}/final_inputs/Supplement/Length/library_stats.hg19.csv"))

load(str_glue("{prefix_data}/final_inputs/Supplement/Length/libraries_length.Rda"))

samples <- names(myLength)
samples <- samples[! samples %in% c("MN0008_L3U", "MN0008_non_U")]

nCol <- max(sapply(myLength, length))
samples_per_page <- 5
cex <- 1.2

blank_plot <- function(){
  plot(1,1, type = "n", bty = "n", axes = F, xlab = NA, ylab = NA)
}

# pdf(paste("~/Projects/Botocudos/Plots/Sequencing/Length_dist_libs_",data_type,".pdf",sep = ""),
#     width = 9, height = 12)

layout(matrix(seq(samples_per_page*nCol), nrow = samples_per_page, byrow = T))

for(s in samples){
  myLibs <- lib_stats$LB[lib_stats$SM == s]
  myRow <- which(samples == s)
  
  if(!myRow %% samples_per_page){
    # plot.new()
    layout(matrix(seq(samples_per_page*nCol), nrow = samples_per_page, byrow = T))  
  }
  
  for(l in myLibs){
    le <- myLength[[s]][[l]]
    # le <- le[le$counts >0,]
    
    plot.title <- paste(s, l)
    barplot(le$counts/sum(le$counts), space = 0,
         col = myColor, bty = "n", 
         xlab = "Length", ylab = "Frequency", #border = "white",
         main = plot.title, lwd = 0.2)
    axis(1, at= seq(0.5, nrow(le)-0.5, 5), 
         labels = seq(min(le$length), max(le$length), 5))
  }
  plots_left <- nCol - length(myLibs)
  while(plots_left > 0){
    blank_plot()
    plots_left <- plots_left - 1
  }
}

# dev.off()
# 


```

## Figs. S3 - S5 Error rates

```{r}
# Fig S3
load(str_glue("{prefix_data}/final_inputs/Supplement/Error_rates/avg_errors.Rda"))
# pdf("~/Projects/Botocudos/Plots/Sequencing/2020_02_25/avg_error_whole.pdf",
#     width = 6, height = 5)
avg_errors %>% subset(Type == "untrimmed") %>%
  ggplot(aes(x = Sample, y = error )) +
  geom_bar(stat = "identity", position = position_dodge(), fill = "salmon") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, size = 10, hjust = 1, vjust = 0.5)) +
  scale_fill_discrete() +
  labs(y = "Error rate", x = NULL, title = "Average error rates") +
  scale_y_continuous(labels = percent)
# dev.off()
```


```{r}
# Fig S4
# pdf("~/Projects/Botocudos/Plots/Sequencing/2020_02_25/error_type_whole.pdf",
#     width = 12, height = 10)
load(str_glue("{prefix_data}/final_inputs/Supplement/Error_rates/error_rates_types.Rda"))
dat %>% subset(Type == "untrimmed") %>%
  gather(name, value, -Sample, -Type) %>%
  mutate(name = sub("\\....", " to ", name)) %>%
  ggplot(aes(x = Sample, y = value)) +
  geom_bar(stat = "identity", position = position_dodge(), fill = "salmon") +
  facet_wrap(.~name, ncol = 3) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, size = 9, hjust = 1)) +
  scale_fill_discrete() +
  labs(y = "Error rate", x = NULL, title = "Type-specific error rates") +
  scale_y_continuous(labels = percent)

# dev.off()
```


```{r}
# Fig S5
# pdf("~/Projects/Botocudos/Plots/Sequencing/2020_02_25/avg_error_MN8_trim.pdf",
#     width = 6, height = 5)
load(str_glue("{prefix_data}/final_inputs/Supplement/Error_rates/trimmed_avg.Rda"))
trimmed_avg %>% 
  gather(name, value, -Sample, -Type, -percent, -error) %>%
  mutate(
         bp_trimmed = ifelse(Type == "trimmed", sub("\\.bam", "", sub(".*trim", "", Sample)), 0),
         Sample = sub("\\..*", "", Sample)
         ) %>%
  ggplot(aes(x = Sample, y = error, fill = bp_trimmed )) +
  geom_bar(stat = "identity", position = position_dodge()) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, size = 10, hjust = 1)) +
  scale_fill_discrete() +
  labs(y = "Error rate", x = NULL, title = "Average error rates") +
  scale_y_continuous(labels = percent)
#dev.off()
```




## Fig. S6 MDS - Dimensions 2 and 3
```{r supplementary dimensions, fig.width=5, fig.height=4.5}
#------------------------------------------------------------------------------#
# Set some paths and variables
mds_data_path <- str_glue("{prefix_data}/final_inputs/Fig_2/MDS_rmTrans_data_to_plot.Rda")
#pdf_path <- "~/Projects/Botocudos/Plots/dummy_figures/Supplement_MDS_two_three.pdf"

#------------------------------------------------------------------------------#
#pdf(pdf_path, width = 10, height = 4.5, useDingbats = F)
#------------------------------------------------------------------------------#
# MDS and fastNGSadmix results
# load data and do last modifications for the MDS
load(mds_data_path)

#-----------------------------------------------------------
# first mds
par(mar = c(bottom=4, left=4, top=2, right=0))
#-----------------------------------------------------------
# second mds
data_to_plot$dim1 <- 3
data_to_plot$dim2 <- 2

data_to_plot$xlab <- paste0(
  "Dim", data_to_plot$dim1, ": ",
  percent(data_to_plot$eig[data_to_plot$dim1]/sum(abs(data_to_plot$eig)))
)

par(mar = c(bottom=4, left=4, top=2, right=2))#, yaxt = "n")
do.call( mds_plot,  data_to_plot )

# dev.off()

```

## Datasets S5 and S6 MDS per individual

```{r, fig.width=16, fig.height=18}


path_mds <- str_glue("{prefix_data}/final_inputs/Fig_2/Multiple_MDS_data_to_plot.Rda")

rmTrans <- F

dim1 <- 1
dim2 <- 2
xlim <- c(-0.15, 0.3)
ylim <- c(-0.2, 0.2)

#pdf_path <- "~/Projects/Botocudos/Plots/MDS/2020_05/MDS_Wollstein_AllSamples_rmTrans.pdf"

label_x <- -.1
label_y <- -.1

dim1 <- 1
dim2 <- 2
for(rmTrans in c(F, T)){
  
  if(!rmTrans){
    cat("All SNPs")
    #pdf_path <- sub("_rmTrans", "", pdf_path)
    path_mds <- sub("_rmTrans", "", path_mds)
    xlim = c(-.25, .6)
    ylim = c(-.5, .5)
    label_x <- -.13
    label_y <- -.2
  }else{
    cat("Transitions removed")
  }
  
  load(path_mds)
  
  samples <- names(MDS)
  samples <- samples[-length(samples)]
  
  # pdf(pdf_path,
  #     width = 16, height = 18)
  
  layout(mat = matrix(seq(1,16), nrow = 4, byrow = T))
  
  for( sample in samples){
    
    points <- MDS[[sample]]$points
    xlab <- MDS[[sample]]$xlab
    ylab <- MDS[[sample]]$ylab
    subtitle <- MDS[[sample]]$subtitle
    
    index <- which(points$indID == paste(sample, ".hg19", sep = ""))
    
    plot(points[,dim1], points[,dim2],
         xlab = xlab, ylab = ylab,
         xlim = xlim,
         ylim = ylim,
         bty = "n", main = sample,
         sub = subtitle,
         type = "n")
    grid()
    points(points[ -index , dim1 ], points[ -index, dim2 ],
           col = alpha(as.character(points$color[ -index ]), 0.5),
           bg = as.character(points$bg[ -index ]),
           pch = points$pch[ -index ],
           cex = 2.2)
    points(points[index, dim1], points[index, dim2],
           col = as.character(points$color[index]),
           bg = as.character(points$bg[index]),
           pch = points$pch[index],
           cex = 2.2)
    # Add text
    text("Polynesia",
         x = -0.2, y = 0.3,
         col = "#29bdad", font = 2, adj = 0)
    text("Americas",
         x = 0, y = 0,
         col = "#b319ab", font = 2, adj = 0)
    
    text(
      label = sample, x =  label_x, y = label_y,
      font = 1, adj = c(0.5, 0.5)
    )
    
    arrows(
      x0 = label_x + 0.01, y0 = label_y + 0.01, 
      x1 = points[index, dim1], y1 = points[index, dim2] - 0.005,
      length = 0.1, angle = 15, col = "gray"
    )
  }
  
  myLegend <- MDS[["legend"]]
  
  plot(1:10, 1:10, axes = F, ylab = NA, xlab = NA, type = "n" )
  legend(1, 10, 
         legend = myLegend$legend, 
         cex = 1,
         pt.cex = 1.8, 
         pt.lwd = 1.2,
         col = as.character(myLegend$color),
         pt.bg = as.character(myLegend$bg),
         pch = myLegend$pch, bty = "n")
}

```



## Fig. S7 fastNGSadmix

```{r plot ancestry of 100 bootstraps, fig.width = 8, fig.height = 14}


myColors <- c(
  "#5c2c45",
  "#ffb852",
  "#e81900",
  "#b3d9a3",
  "#29bdad",  
  "#b319ab",
  "black"
)

ancestry_to_plot <- read.csv(str_glue("{prefix_data}/final_inputs/Supplement/fastNGSadmix/Ancestry_fastNGSadmix_Wollstein_plot.csv"))

change_labels <- function(old_labels){
  # browser()
  if("rmTrans" %in% colnames(old_labels)){
    new_labels <- lapply(old_labels, function(x) ifelse(x, "Transitions removed", "All SNPs"))
  }else if("k_pop" %in% colnames(old_labels)){
    lapply(old_labels, function(x) x)
  }

}

plot_title <- "Admixture proportions with fastNGSadmix and Wollstein_Xing panel (100 bootstraps)"


ancestry_to_plot$x_axis <- paste0(ancestry_to_plot$Sample, ", #SNPs: ",
                                  prettyNum(ancestry_to_plot$SNPs, big.mark = ","))

p1 <- ggplot(ancestry_to_plot,
       aes(x = Sample, y = avg_ancestry, color = k_pop,
           ymin = avg_ancestry - 1*sd_ancestry,
           ymax = avg_ancestry + 1*sd_ancestry)) +
  facet_grid(k_pop ~rmTrans, labeller = change_labels, drop = T) +
  geom_point(position = position_dodge(.9)
             ) +
  geom_errorbar(position = position_dodge(.9)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1), #element_blank(),
        legend.position = "none") +
  labs(x = NULL, y = "Average percentage of ancestry +- 1 SD",
       title = plot_title) +
  scale_x_discrete(drop = T) +
  scale_y_continuous(labels = percent) +
  scale_color_manual(values = myColors) 


empty_labels <- function(old_labels){
  
  if("rmTrans" %in% colnames(old_labels)){
    new_labels <- lapply(old_labels, function(x) ifelse(x, "", ""))
  }
}

my_annotation <- unique(ancestry_to_plot[,c("Sample", "DoC_endogenous", "SNPs", "rmTrans")])

p2 <- ggplot(my_annotation,
             aes(x = prettyNum(factor(SNPs), big.mark  =","), y = 1)) +
  facet_wrap(. ~ rmTrans, labeller = empty_labels, scale = "free_x") +
  coord_cartesian(ylim = c(0,1)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        axis.text.y = element_text(angle = 90),
        legend.position = "none",
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = margin(5,22,15,35)) +
  labs(x = NULL, y = NULL) +
  scale_y_continuous(breaks = c(0.,0.01), labels = c("", "SNPs")) 

p3 <- ggplot(my_annotation,
             aes(x = factor(signif(DoC_endogenous,3)), y = 1)) +
  facet_wrap(. ~ rmTrans, labeller = empty_labels, scale = "free_x") +
  coord_cartesian(ylim = c(0,1)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        axis.text.y = element_text(angle = 90),
        legend.position = "none",
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = margin(5,22,15,35)) +
  labs(x = NULL, y = NULL) +
  scale_y_continuous(breaks = c(0.,0.01), labels = c("", "DoC")) 

# pdf("~/Projects/Botocudos/Plots/dummy_figures/fastNGSadmix_100boot_Wollstein.pdf", width = 8, height = 14)
#pdf("~/Desktop/test.pdf", width = 10, height = 16)
plot_grid(p1, p2, p3, ncol = 1, rel_heights = c(6, 0.2, 0.2))
#dev.off()

```

## Figs. S8 - S12 F3


```{r}


colorspace::sequential_hcl(n = 50, h = c(286, -128), c = c(121, 61, 28), l = c(38, 90), power = c(0, 1.3), rev = F, register = "Custom-Palette")


```

```{r, fig.width=18, fig.height=9}
load(str_glue("{prefix_data}/final_inputs/Supplement/F3/f3_values.Rda"))

for(group in f3_values){
  
  boto_f3 <- group$boto_f3
  
  ancient <- which(boto_f3$type == "Ancient")
  boto_f3$Source_2[ancient] <- paste("+", boto_f3$Source_2[ancient])
  

  #------------------------------------------------------------------------------#
  # Map
  
  
  if(group[["map"]] == "Brazil"){
    cex_main_f3 <- 1.5
    # pdf(paste0("~/Projects/Botocudos/Plots/F3/2021_02/", group[["plot_basename"]], ".pdf"),
    #     width = group[["width"]] + 8, height = group[["height"]], useDingbats = F)
    
    par(mar = c(5,4,4,2), cex.main = 1.5)

    # layout(mat = matrix(1:2, nrow = 1), widths = c(1,6))
    layout(matrix(c(rep(1,21),2,3), nrow = 1, ncol = 23, byrow = TRUE),
           widths = c(rep(6/21, 21,), 1.5, 6))


    draw_map(
      xlim = c(-70, -30),
      ylim = c(-30, 10),
      land_color = "white",
      sea_color = "whitesmoke",
      border_color = "gray50",
      title_text =  paste("\n\n\n",  group[["main"]])
    )
    
    text(
      x = boto_f3$long + 2,
      y = boto_f3$lat + 2,
      labels = boto_f3$population,
      cex = 2
    )
  points(x = boto_f3$long,
         y = boto_f3$lat,
         pch = 16,
         col = ggplot2::alpha(boto_f3$color, 0.9),
         lwd = 2,
         cex = 5.5)
  }else{next}
  par(mar=c(5,3,6,6))
  draw_colorscale(boto_f3 = boto_f3, palette = "Custom-Palette",
                  palette_function = sequential_hcl, rev = T)

  #------------------------------------------------------------------------------#
  # F3
  par(bg="white", cex.main = cex_main_f3, cex.axis = 2)
  plot_F3(unique(boto_f3[,c("f_3", "Source_2","std.err","color")]),
        x = "f_3", y = "Source_2", 
        stdErr = "std.err", 
        lwd_1stdErr = 4, lwd_3stdErr = 1, 
        xlab = "", cex.axis = 1.5,
        y_mar = 20,
        cex_point = 2.2,
        main = group[["main"]],  
        x1 = group[["x1"]], x2 = group[["x2"]]
        )
  
  
  
  # dev.off()
  
}
```


```{r, fig.width=25, fig.height=25}
load(str_glue("{prefix_data}/final_inputs/Supplement/F3/f3_values.Rda"))

for(group in f3_values){
  
  boto_f3 <- group$boto_f3
  
  ancient <- which(boto_f3$type == "Ancient")
  boto_f3$Source_2[ancient] <- paste("+", boto_f3$Source_2[ancient])
  

  #------------------------------------------------------------------------------#
  # Map
  
  
  if(group[["map"]] == "Brazil"){
    next
  }else if(group[["map"]] == "Americas"){
    cex_main_f3 <- 2
    # pdf(paste0("~/Projects/Botocudos/Plots/F3/2021_02/", group[["plot_basename"]], ".pdf"),
    #     width = group[["width"]] + 14, height = group[["height"]], useDingbats = F)

    layout(matrix(c(rep(1,21),2,3), 
                  nrow = 1, ncol = 23, byrow = TRUE),
           widths = c(rep(12/21, 21), 1, 6))
    par(cex.main = 2)

    draw_map(xlim = c(-178, -13),
             ylim = c(-60, 90),
             land_color = "white",
             sea_color = "whitesmoke",
             border_color = "gray50",
             title_text = paste("\n\n\n", group[["main"]])
             )
      
    legend("bottomleft", pch = "+", col = "black", legend = "Ancient individual",
           bty = "n", cex = 2.5, bg = "white")
    
    points(x = boto_f3$long,
           y = boto_f3$lat,
           pch = 16,
           col = ggplot2::alpha(boto_f3$color, 0.9),
           lwd = 2,
           cex = 5.5)
    
    points(x = boto_f3$long[ancient],
           y = boto_f3$lat[ancient],
           pch = "+",
           col = "black",
           lwd = 2,
           cex = 5.5)
  }
  par(mar=c(5,3,6,6))
  draw_colorscale(boto_f3 = boto_f3, palette = "Custom-Palette",
                  palette_function = sequential_hcl, rev = T)

  #------------------------------------------------------------------------------#
  # F3
  par(bg="white", cex.main = cex_main_f3, cex.axis = 2)
  plot_F3(unique(boto_f3[,c("f_3", "Source_2","std.err","color")]),
        x = "f_3", y = "Source_2", 
        stdErr = "std.err", 
        lwd_1stdErr = 4, lwd_3stdErr = 1, 
        xlab = "", cex.axis = 1.5,
        y_mar = 20,
        cex_point = 2.2,
        main = group[["main"]],  
        x1 = group[["x1"]], x2 = group[["x2"]]
        )
  
  
  
  # dev.off()
  
}
```

## Datasets S8 - S14 NGSAdmix

```{r, fig.width=12, fig.height=10}

groups_to_plot <- c(
  "Raghavan278ind_Castro12NatAm_Skoglund", "Raghavan278ind_CastroTupGuaUnrel",
  "Raghavan278ind_CastroTupGuaUnrel_Skoglund_Castro12NatAm", "YF_Skoglund_Castro12NatAm",
  "YF_Skoglund_Castro12NatAm_CastroTupGuaUnrelated"
  )

for(group in groups_to_plot){
  load( str_glue("{prefix_data}/final_inputs/Supplement/NGSadmix/{group}.Rda") )
  for(plot_object in ngsadmix_object){
    plot_object$plot(do_layout = T, add_text = T)
  }
}



```


## Fig. S14 Conditional heterozygosity

```{r,  fig.width = 12, fig.height = 6}
boto_pi <- read.csv(
  str_glue("{prefix_data}/final_inputs/Supplement/Heterozygosity/heterozygosity_coverage_botocudos_plot.csv")
  )

all_pairs <- ggplot(boto_pi,
                    aes(
                      x = Population.ID, 
                      y = estimate, 
                      ymin = confint1, 
                      ymax = confint2, 
                      color = Region, 
                      interaction = Region
                    )
) +
  geom_violin() +
  geom_jitter() +
  labs(title = "All pairs", x = NULL) +
  theme(legend.position = "none", title = element_text(size = 8)) +
  coord_cartesian(ylim = c(0.05, 0.20))



below_above <- ggplot(boto_pi[boto_pi$ind1 %in% c("MN0008", "MN00056", "MN00013") 
                              | boto_pi$ind2 %in% c("MN0008", "MN00056", "MN00013"), ],
                      aes(
                        x = Population.ID, 
                        y = estimate, 
                        ymin = confint1, 
                        ymax = confint2, 
                        color = Region, 
                        interaction = Region
                      )
) +
  geom_violin() +
  geom_jitter() +
  labs(title = "One individual with DoC > 2x \nand one with DoC < 2x", x = NULL) +
  theme(legend.position = "none", title = element_text(size = 8)) +
  coord_cartesian(ylim = c(0.05, 0.20))

below <- ggplot(boto_pi[! (boto_pi$ind1 %in% c("MN0008", "MN00056", "MN00013") 
                           & boto_pi$ind2 %in% c("MN0008", "MN00056", "MN00013")), ],
                aes(
                  x = Population.ID, 
                  y = estimate, 
                  ymin = confint1, 
                  ymax = confint2, 
                  color = Region, 
                  interaction = Region
                )
) +
  geom_violin() +
  geom_jitter() +
  labs(title = "Both individuals with DoC < 2x", x = NULL) +
  theme(legend.position = "none", title = element_text(size = 8)) +
  coord_cartesian(ylim = c(0.05, 0.20))

above <- ggplot(boto_pi[boto_pi$ind1 %in% c("MN0008", "MN00056", "MN00013") 
                        & boto_pi$ind2 %in% c("MN0008", "MN00056", "MN00013"), ],
                aes(
                  x = Population.ID, 
                  y = estimate, 
                  ymin = confint1, 
                  ymax = confint2, 
                  color = Region, 
                  interaction = Region
                )
) +
  geom_violin() +
  geom_jitter() +
  labs(title = "Both individuals with DoC > 2x", x = NULL) +
  theme(legend.position = "none", title = element_text(size = 8)) +
  coord_cartesian(ylim = c(0.05, 0.20))

#pdf("~/Projects/Botocudos/Plots/Het/2021_01_13/condhet_by_coverage_Botocudos.pdf",
#    width = 12, height = 6)
plot_grid(all_pairs, below, below_above, above, nrow = 1)

#dev.off()
```

## Fig. S14 Marginal SFS
```{r Marginal SFS, fig.width=10, fig.height=3}
model <- "model2_estAfr_1moredeme"
#prefix <- str_glue("~/Projects/Botocudos/Files/fastsimcoal2/2022_12/summary_{model}/{model}")

# png(
#   "~/Projects/Botocudos/Plots/fastsimcoal2/sfs_fit_model2_estAfr_1moredeme.png",
#   width = 10, height = 3, res = 300, units = "in"
#      )
# pdf(
#   "~/Projects/Botocudos/Plots/fastsimcoal2/sfs_fit_model2_estAfr_1moredeme.pdf",
#   width = 10, height = 3
#      )

par(mfrow = c(1,4))

for(pop in c("Botocudos", "Karitiana", "Surui", "Yoruba")){
    get_marginal_SFS(
      miny = 0, maxy = 60000,
      bestrun_bestlhoods_path = str_glue("{prefix_data}/final_inputs/Fig_5/{model}.bestlhoods"),
      model = model,
      sfsFile.obs = str_glue("{prefix_data}/final_inputs/Fig_5/Botoc_Karit_Surui_Yoruba.multisfs"), #"~/Projects/Botocudos/Files/fastsimcoal2/2022_11/Botoc_Karit_Surui_Yoruba.multisfs",
      sfsFile.exp = str_glue("{prefix_data}/final_inputs/Fig_5/{model}_DSFS.txt"), #paste0(prefix, "_DSFS.txt"),
      new.pop.Labels = c('"Botocudos"', "Yjxa", "Paiter", "Yoruba"), 
      pop.Labels = c("Botocudos", "Karitiana", "Surui", "Yoruba"), 
      add_lines =  F,
      colors = c("red", "black"), population = pop,
      xlab = "Frequency (derived allele)",
      ylab = "Number of sites"
    )

}

legend(
  1, 55000,
  col = c("black", "red"),
  pch = 16,
  lwd = 1,
  bty = "n",
  cex = 1.2,
  legend = c("Observed", "Expected")
)

# dev.off()

```

## Fig. S15 Worse SFS entries
```{r worse SFS entries, fig.height=3, fig.width=9}
#par(mfrow = c(4, 6))
# pdf("~/Projects/Botocudos/Plots/fastsimcoal2/worse_entries_Jan2023.pdf",
#     width = 9, height = 3)

# png("~/Projects/Botocudos/Plots/fastsimcoal2/worse_entries_Jan2023.png",
#     width = 9, height = 3, units = "in", res = 300)
  par(mar =  c(4, 6, 2, 1), mfrow = c(1,2), cex = 0.7)

  plot_worst_entries(
    
    sfsFile.obs = str_glue("{prefix_data}/final_inputs/Fig_5/Botoc_Karit_Surui_Yoruba.multisfs"), 
    sfsFile.exp = str_glue("{prefix_data}/final_inputs/Fig_5/{model}_DSFS.txt"),
    color = "#c9303e", numBadEntriesToPLot = 20,
    ylab1 = "Difference in likelihood per entry",
    ylab2 = "Number of sites",
    main1 = "Likelihood difference", #str_glue("{model} \n diff in lhoods"),
    main2 = "Difference in number of sites" #str_glue("{model} \n diff in # of SFS entries"),
    
  )
  
# dev.off()

```

## Fig. S16 Bootstrapped parameters
```{r read parameters}
my_levels <- c(
  "Ne_Botocudos", "Ne_Karitiana", "Ne_Surui", "Ne_SAm", "Ne_Je", "Ne_Tupi",
  "Ne_Americas", "Ne_OutOfAfr", "Ne_Yoruba", "Ne_Anc",
  
  "FIS_Botocudos", "FIS_Karitiana", "FIS_Surui",
  
  "ADMBot", "ADMKar", "ADMSur", "ADMPR1", "ADMPR2",
  
  "TADM_Botocudos", "TADM_Karitiana", "TADM_Surui", "TADM_Yor1", "TADM_Yor2",
  
  "bn_Bot", "bn_Kar", "bn_Sur", "bn_Yor",
  
  "TBN_Botocudos", "TBN_Karitiana", "TBN_Surui", "TBN_Yoruba",
  
  "TDIV_Botocudos", "TDIV_Karitiana", "TDIV_Surui", "TDIV_SouthAm", "TDIV_Americas", 
  "TDIV_OutOfAfr", "TDIV_Yoruba"
)

params_long <- read.table(str_glue("{prefix_data}/final_inputs/Fig_5/params_long.txt"), header = T)
params_long$variable <- factor(params_long$variable, levels = my_levels, ordered = T)
```


```{r store plots}
#------------------------------------------------------------------------------
quantiles_95 <- function(x){
  r <- quantile(x, probs=c(0.025, 0.25, 0.5, 0.75, 0.975))
  names(r) <- c("ymin", "lower", "middle", "upper", "ymax")
  r
}

#------------------------------------------------------------------------------

options(scipen=10000)
gen_time <- 25
my_alpha <- 0.2
bg_color <- "gray90"
yoruba_color = "#ffff00"
font_size_x <- 9
font_size_y <- 8
size_y_label <- 10
strip_size <- 9
title_size  <- 10
title_face <- "bold"

pop_colors =  c(
  #"Botocudos" = 
    "#d50c42",   
  #"Karitiana" = 
    "#0057ba", 
  #"Surui" = 
    "#0057ba",
  #"South America" = 
    "gray", 
  #"Americas" = 
    "gray", 
  #"Out of Africa" = 
    "gray", 
  # Yoruba
    "yellow",
  #"Ancestral" = 
    "gray"
                )
dot_col <- "red" ; dot_fill <- "black"; dot_shape = 21

#------------------------------------------------------------------------------

# Effective population sizes
# NBOTO,NSURU,NKARI,NDINK,NAMER,NBRAZ,NTUPI,NJE,NANC,

plot_Ne <-  
ggplot(
  params_long[
    params_long$variable %in% c(
      "Ne_Anc",
      "Ne_Yoruba",
      "Ne_OutOfAfr", 
      "Ne_Americas", 
      "Ne_SAm",
      "Ne_Botocudos", "Ne_Karitiana", "Ne_Surui"
    ), 
  ], aes(
    x = variable, y = value / 2, color = variable
  )
) +
  geom_jitter(alpha = my_alpha) +
  stat_summary(fun.data = quantiles_95, geom = "boxplot", fill = NA, color = "black") +
  scale_y_log10() +
  theme_minimal_hgrid() +
  scale_x_discrete(labels = c(
    "Ne_Anc" = "Ancestral",
    "Ne_OutOfAfr" = "Out of Africa", 
    "Ne_Americas" = "Americas", 
    "Ne_SAm" = "South America",
    "Ne_Yoruba" = "Yoruba",
    "Ne_Botocudos" = '"Botocudos"', "Ne_Karitiana" = "Yjxa", "Ne_Surui" = "Paiter")) +
  scale_color_manual(values = pop_colors) +
  labs(title = "Effective population size", x = NULL, y = expression(paste(N[e]))) +
  theme(
    legend.position = "none", 
    axis.text.x = element_text(angle = 45, hjust = 1, size = font_size_x),
    axis.text.y = element_text(size = font_size_y),
    strip.background =element_rect(fill = bg_color),
    strip.text = element_text(size = strip_size),
    plot.title = element_text(face = title_face, size = title_size), 
    axis.title.y = element_text(size = size_y_label)
    ) +
  geom_point(
    data = data.frame(
      value = model_values$Nes / 2, 
      variable = c("Ne_Botocudos", "Ne_Karitiana", "Ne_Surui", "Ne_Americas", "Ne_OutOfAfr", "Ne_Anc", "Ne_Yoruba", "Ne_SAm")
    ),
    color = dot_col, fill = dot_fill, pch = dot_shape
  ) 

#------------------------------------------------------------------------------
# Split times
popnames <- list(
  model2_estAfr_1moredeme = c("Botocudos", "Karitiana",
                              "Surui", "Americas", "Out of Africa", "ANC", "Yoruba","South America")
)


is_split <- model_values$migrants == 1
sources <- popnames$model2_estAfr_1moredeme[ model_values$sources[is_split] + 1 ]
sinks <- popnames$model2_estAfr_1moredeme[ model_values$sinks[is_split] + 1 ]
splits_ml <- data.frame(
  variable = paste0("TDIV_", sub("Out of Africa", "OutOfAfr", sub("South America", "SouthAm", sources))), 
  value = model_values$times[is_split] 
  )

plot_tdiv <- 
ggplot(
  params_long[params_long$variable %in% c(
    "TDIV_OutOfAfr", 
    "TDIV_Yoruba",
    "TDIV_Americas", 
    "TDIV_SouthAm",
    "TDIV_Botocudos", "TDIV_Karitiana", "TDIV_Surui"
    
  ), ], aes(
    x = variable, y = value * gen_time, color = variable
  )
) +
  geom_jitter(alpha = my_alpha) +
  stat_summary(fun.data = quantiles_95, geom="boxplot", fill = NA, color = "black") +
  scale_y_log10() +
  theme_minimal_hgrid() +
  scale_x_discrete(labels = c(
    "TDIV_OutOfAfr" = "Out of Africa", 
    "TDIV_Americas" = "Americas", 
    "TDIV_Yoruba" = "Yoruba",
    "TDIV_SouthAm" = "South America",
    "TDIV_Botocudos" = '"Botocudos"', "TDIV_Karitiana" = "Yjxa", "TDIV_Surui" = "Paiter")
    ) +
  scale_color_manual(values = pop_colors) +
  theme(
    legend.position = "none", 
    axis.text.x = element_text(angle = 45, hjust = 1, size = font_size_x),
    axis.text.y = element_text(size = font_size_y),
        strip.background = element_rect(fill = bg_color),
    strip.text = element_text(size = strip_size),
    plot.title = element_text(face = title_face, size = title_size), 
    axis.title.y = element_text(size = size_y_label)
    ) +
  labs(title = "Split from ancestral population", x = NULL, y = "Years")  +
  geom_point(
    data = splits_ml,
    color = dot_col, fill = dot_fill, pch = dot_shape
  ) 

#------------------------------------------------------------------------------

# ggplot(params_wide, aes(x = Ne_Surui, y = TDIV_Surui, color = core_model, shape = type_model)) +
#   geom_point() +
#   facet_wrap(type_model ~ core_model)


#------------------------------------------------------------------------------

#------------------------------------------------------------------------------
# Inbreeding coefficients
# FIS_Botoc,FIS_Karit,FIS_Surui,FIS_Dinka,
# params_long$variable[params_long$variable == "FIS_Botoc"] <- "FIS_Botocudos"
# params_long$variable[params_long$variable == "FIS_Karit"] <- "FIS_Karitiana"

plot_fis <- 
  ggplot(
  params_long[params_long$variable %in% 
                c("FIS_Botocudos","FIS_Karitiana","FIS_Surui"), ], aes(
                  x = variable, y = value, color = variable
                )
) +
  geom_jitter(alpha = my_alpha) +
  stat_summary(fun.data = quantiles_95, geom="boxplot", fill = NA, color = "black") +
  theme_minimal_hgrid() +
  scale_x_discrete(labels = c(
    "FIS_Botocudos" = '"Botocudos"', "FIS_Karitiana" = "Yjxa", "FIS_Surui" = "Paiter"
  )
  ) +
  theme(legend.position = "none",  axis.text.x = element_text(angle = 45, hjust = 1, size = font_size_x),
        axis.text.y = element_text(size = font_size_y),
        strip.background =element_rect(fill=bg_color),
    strip.text = element_text(size = strip_size),
    plot.title = element_text(face = title_face, size = title_size), 
    axis.title.y = element_text(size = size_y_label)
    ) +
  scale_color_manual(values = pop_colors[1:3]) +
  labs(title = "Inbreeding coefficient", x = NULL, y = expression(F[IS])) +
  geom_point(
    data = data.frame(
      value = model_values$inbreeding_coef[1:3], 
      variable = c("FIS_Botocudos", "FIS_Karitiana", "FIS_Surui")
    ),
    color = dot_col, fill = dot_fill, pch = dot_shape
  ) 



#------------------------------------------------------------------------------
# admixture proportions
# adm_bot, adm_kar,adm_sur,adm_din,
is_gene_flow <- model_values$migrants < 1 & model_values$migrants > 0
sources <- popnames$model2_estAfr_1moredeme[ model_values$sources[is_gene_flow] + 1 ]
geneflow <- data.frame(
  value = model_values$migrants[is_gene_flow],
  variable = c("ADMBot", "ADMKar", "ADMSur", "ADMPR1", "ADMPR2")
)

plot_adm <-  
  ggplot(
  params_long[params_long$variable %in% 
                c("ADMBot", 
                  "ADMKar",
                  "ADMSur",
                  "ADMPR1", "ADMPR2"), ], aes(
                    x = variable, y = value, color = variable
                  )
) +
  geom_jitter(alpha = my_alpha) +
  stat_summary(fun.data = quantiles_95, geom="boxplot", fill = NA, color = "black") +
  theme_minimal_hgrid() +
  scale_x_discrete(labels = c(
    "ADMBot" = '"Botocudos"', "ADMKar" = "Yjxa", "ADMSur" = "Paiter", "ADMPR1" = "Yoruba (1)", "ADMPR2" = "Yoruba (2)")
  ) +
  scale_y_continuous(labels = percent) +
  scale_color_manual(values = c(pop_colors[1:3], yoruba_color, yoruba_color)) +
  theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust = 1, size = font_size_x),
        axis.text.y = element_text(size = font_size_y),
        strip.background =element_rect(fill=bg_color),
    strip.text = element_text(size = strip_size),
    plot.title = element_text(face = title_face, size = title_size), 
    axis.title.y = element_text(size = size_y_label)
    ) +
  labs(title = "Gene flow (%)", x = NULL, y = "Gene flow received")  +
  geom_point(
    data = geneflow,
    color = dot_col, fill = dot_fill, pch = dot_shape
  ) 


#------------------------------------------------------------------------------
# Bottleneck intensities
# bn_bot,bn_kar,bn_sur,bn_din,
bnecks <- params_long[params_long$variable %in% 
                        c("bn_Bot","bn_Kar","bn_Sur", "bn_Yor"),]
bnecks$population <- sub("bn_", "", bnecks$variable)
bnecks$intensity <- bnecks$value
bnecks$variable <- NULL
bnecks$value <- NULL
bnecks$Ne_bn <- 1 / bnecks$intensity  

sources <- popnames$model2_estAfr_1moredeme[ model_values$sources[model_values$instbot] +1  ]
bneck_ml <- data.frame(
  Ne_bn = 1 / model_values$new_sizes[model_values$instbot] / 2,
  population = c("Bot", "Kar", "Sur", "Yor")
)

plot_bn <- 
  ggplot(
  bnecks, aes(
    x = population, y = Ne_bn / 2, color = population
  )
) +
  geom_jitter(alpha = my_alpha) +
  stat_summary(fun.data = quantiles_95, geom = "boxplot", fill = NA, color = "black") +
  scale_color_manual(values = c(c(pop_colors[1:3]), yoruba_color)) +
  theme_minimal_hgrid() +
  scale_x_discrete(
    labels = c(
    "Bot" = '"Botocudos"', "Kar" = "Yjxa", "Sur" = "Paiter", "Yor" = "Yoruba"
    )
  ) +
  theme(
    legend.position = "none",  
    axis.text.x = element_text(angle = 45, hjust = 1, size = font_size_x),
    axis.text.y = element_text(size = font_size_y),
        strip.background = element_rect(fill=bg_color),
    strip.text = element_text(size = strip_size),
    plot.title = element_text(face = title_face, size = title_size), 
    axis.title.y = element_text(size = size_y_label)
    ) +
  labs(title = "Effective population size (at bottleneck)", x = NULL, y = expression(paste( N[e])))  +
  geom_point(
    data = bneck_ml,
    color = dot_col, fill = dot_fill, pch = dot_shape
  ) 



  
  # TBN_BOTOC,TBN_SURUI,TBN_KARIT,TBN_DINKA,
# TDIV_TUPI,TDIV_JE,TDIV_BRAZ,TDIV_AMER,
# TDIV_BOTOC,TDIV_KARIT,TDIV_SURUI,TDIV_DINKA
# TADM_BOTOC,TADM_SURUI,TADM_KARIT,TADM_DINKA,
#gen_time = 25
  
geneflow <- data.frame(
  value = model_values$times[is_gene_flow],
  variable = c("TADM_Botocudos", "TADM_Karitiana", "TADM_Surui", "TADM_Yor1", "TADM_Yor2")
)

plot_tadm <- 
  ggplot(
  params_long[params_long$variable %in% 
                c("TADM_Botocudos","TADM_Surui","TADM_Karitiana",
                  "TADM_Yor1", "TADM_Yor2"
                ), ], aes(
                  x = variable, y = value * gen_time, color = variable
                )
) +
  geom_jitter(alpha = my_alpha) +
  scale_y_log10() +
  stat_summary(fun.data = quantiles_95, geom="boxplot", fill = NA, color = "black") +
  theme_minimal_hgrid() +
  scale_x_discrete(labels = c(
    "TADM_Botocudos" = '"Botocudos"', "TADM_Karitiana" = "Yjxa", 
    "TADM_Surui" = "Paiter", "TADM_Yor1" = "Yoruba (1)", "TADM_Yor2" = "Yoruba (2)"
  )
  ) +
  theme(
    legend.position = "none", 
    axis.text.x = element_text(angle = 45, hjust = 1, size = font_size_x),
    axis.text.y = element_text(size = font_size_y),
    strip.background = element_rect(fill=bg_color),
    strip.text = element_text(size = strip_size),
    plot.title = element_text(face = title_face, size = title_size), 
    axis.title.y = element_text(size = size_y_label)
  ) +
  scale_color_manual(values = c(c(pop_colors[1:3]), yoruba_color, yoruba_color)) +
  labs(title = "Time to gene flow", y = "Years", x = NULL)  +
  geom_point(
    data = geneflow,
    color = dot_col, fill = dot_fill, pch = dot_shape
  ) 

```


```{r, fig.width=9, fig.height=5}

#png("~/Projects/Botocudos/Plots/fastsimcoal2/estimated_params_95CI.png", width = 9, height = 5, res = 600, units = "in")
plot_grid(
  plot_tdiv, plot_fis, plot_tadm,
  plot_Ne, plot_bn, plot_adm, 
  ncol = 3, 
  rel_widths = c(12,8,7)
  )
#dev.off()

```


## Fig. S17 Dstats

```{r supplementary figure,  fig.width = 9.5, fig.height = 7.5}
# 22 Botocudos

file <- "July_09_trim5"
abba <- read.table(
  str_glue("{prefix_data}/final_inputs/Supplement/Dstat/July_09_trim5.ErrorCorr.TransRem.txt"),
                   header = T, stringsAsFactors = F)



pops_in_title = c("Mixe")
pops_in_h3 <- c(
  "French", 
  "Han",
  "Andaman_trim5",
  "Papuan",
  "Australian"
  )
pops_in_h3_labels <-  c("French\n(n = 1)",
                                 "Han\n(n = 1)",
                                 "Andaman\n(n = 1)",
                                 "Papuan\n(n = 1)",
                                 "Australian\n(n = 7)")


pop_colors <- c("gray60", "gray80", "#CC85D1", "#96BFE6", "#0057BA", "black")

pops_in_y <-  c(
  "Surui",
  "LagoaSanta",
  "LagoaSanta_trim5",
  "LagoaSanta",
  "MN0008",
  # "MN0008_L3U_trim5",
  "MN0008_L3U_trim2",
  "Karitiana",
  "Aymara",
  "22Botocudos_trim5",
  "Huichol"
                   )

new_labels = list(
  "MN0008_L3U_trim2" = "MN0008_L3U_trim2\n(n = 1)",
  "22Botocudos_trim5" = "Botocudos_trim5\n(n = 22)",
  "Surui" = "Surui\n(n = 2)",
  "LagoaSanta_trim5" = "Lagoa Santa\n(n = 1)",
  # "MN0008_L3U_trim5" = "MN0008_L3U_trim5\n(n = 1)",
  # "MN0008" = "MN0008\n(n = 1)",
  "Karitiana" = "Karitiana\n(n = 4)",
  "Aymara" = "Aymara\n(n = 1)",
  "Huichol" = "Huichol\n(n = 1)"
)

pop <- "Mixe"
outgroup <- "Yoruba"

Dstat <- abba[(abba$H1 %in% pops_in_title |abba$H2 %in% pops_in_title) &
                (abba$H3 %in% pops_in_h3),]
df_y <- as.data.frame(pops_in_y)
df_y$y_axis <- rownames(df_y)
plot_list <- list()

Dstat <- abba[(abba$H1 == pop |abba$H2 == pop) &
                (abba$H3 %in% pops_in_h3),]
Dstat <- switch_h2_h1(Dstat, pop)
Dstat <- Dstat[!(Dstat$H1 %in% pops_in_h3),]


Dstat <- Dstat[Dstat$H1 %in% pops_in_y,]

for(label in names(new_labels)){
  print(label)
  Dstat[Dstat$H1 == label,]$H1 <- new_labels[[label]]
}


Dstat$H1 <- factor(Dstat$H1, 
                   levels = rev(unique(new_labels))
)
Dstat$H3 <- factor(Dstat$H3, 
                   levels = c(pops_in_h3),
                   ordered = T)

x_label <- expression(
  atop(
    "Z",
    paste("(", H[1], ", ", H[3], ") < --- > (Mixe, ", H[3], ")"),
    paste("(Mixe, Yoruba)",  " < --- > (", H[1], ", Yoruba)")
  )
)

p <- ggplot(Dstat, aes(x = Z, y = H1, 
                       fill = H3)) +
  
  coord_cartesian(xlim = c(-4,3.5)) +
  scale_fill_manual(values = alpha(pop_colors, .8),
                    breaks = pops_in_h3,
                    labels = pops_in_h3_labels,
                    name = expression(paste(H[3], ":"))) +
  geom_vline(xintercept = c(-3.3,3.3), lty = "dashed", col = "brown", lwd = 1)+
  geom_vline(xintercept = c(-4,-3,-2,-1,0,1,2,3), lty = "solid", col = "gray", lwd = 0.2) +
  geom_vline(xintercept = c(-3.5,-2.5,-1.5,-0.5,0.5,1.5,2.5,3.5), lty = "dashed", col = "gray", lwd = 0.2) +
  
  geom_point(shape = 23, size = 7, stroke= 1) +
  labs(
    x = x_label,
    y = expression(H[1]),
    title = expression(bold(paste("Z-scores for D(", H[1], ", ", H[2]," = Mixe; ", H[3], ", Yoruba)" )))) +
  theme_cowplot() +
  theme(legend.position = "top", 
        axis.text.x = element_text(size = 16),
        axis.text.y = element_text(size = 16)
  ) 


#pdf("~/Projects/Botocudos/Plots/dummy_figures/Dstat_SI.pdf", width = 9.5, height = 7.5)
p
#dev.off()

```

## Fig. S18 Diet comparison (Isotopes)
Done

```{r}

isotopes <- read.csv(
  str_glue("{prefix_data}/final_inputs/Supplement/Isotopes/comparisons_August2021.csv"),
                    stringsAsFactors = F)

boto <- subset(isotopes, Name == '"Botocudos"')
isotopes <- subset(isotopes, Include_in_analysis == "Yes")
#isotopes <- subset(isotopes, Include_in_analysis == "Yes"|Name == 'Polynesian-"Botocudos"')
# prepare and order factors for the plot, according to what is written in the manuscript

names_groups <- list(
  '"Botocudos"'                              = '"Botocudos" (this study)',
  "Sambaqui"                                 = "Sambaqui (this study)",
  #'Polynesian-"Botocudos"'                   = 'Polynesian-"Botocudos" (2014)',
  "Lagoa Santa"                              = "(i) Early Holocene Hunter-gatherers",
  "Riverine sambaqui"                        = "(ii) Mid-Holocene riverin sambaquis",
  "Coastal sambaqui"                         = "(iii) 3K-1K yo Coastal sambaquis",
  "Present day, Rio de Janeiro and Brasilia" = "(iv) Present day, Rio de Janeiro and Brasilia"
)

for(label in names(names_groups)){
  index <- which(isotopes$Name == label)
  isotopes$Name[index] = names_groups[[label]]
}

isotopes$Name <- factor(
  isotopes$Name,
  levels = sapply(1:length(names_groups), function(i) unlist(names_groups)[[i]]),
  ordered = T
)



ggplot(isotopes, aes(d13C, d15N, fill = Name)) +
  geom_point(size = 5, alpha = 0.8, stroke = 1, shape = 21, color = "black") +
  theme_minimal() +
  labs(
    x = expression(
      paste(delta^13, "C")
    ),
    y = expression(
      paste(delta^15, "N")
    ),
    title = "Isotopic values, bones and permanent second and third molars"
  ) +
  scale_fill_discrete() +
  theme(legend.position = c(0.3,0.8))


# ggsave("~/Projects/Botocudos/Plots/Isotopes/isotopes_SI_2023.pdf", width = 7, height = 7)

```

```{r}

require(car)
#------------------------------------------------------------------------------#
# checking equal variances
# p-values > 0.05, then equal variances can be assumed
Boto_Rio <- subset(isotopes, Name %in% c('"Botocudos" (this study)', "(iv) Present day, Rio de Janeiro and Brasilia"))
Boto_Lagoa <- subset(isotopes, Name %in% c('"Botocudos" (this study)', "(i) Early Holocene Hunter-gatherers"))

leveneTest(d15N ~ Name, Boto_Rio)
leveneTest(d13C ~ Name, Boto_Rio)

leveneTest(d15N ~ Name, Boto_Lagoa)
leveneTest(d13C ~ Name, Boto_Lagoa)

# (we can assume equal variances in the four comparisons)
#------------------------------------------------------------------------------#
# 
TukeyHSD(aov(d15N ~ Name, Boto_Rio))
summary(aov(d15N ~ Name, Boto_Rio))

TukeyHSD(aov(d15N ~ Name, Boto_Lagoa))
summary(aov(d15N ~ Name, Boto_Lagoa))


```




# Session info

```{r}
sessionInfo()
```

