---
title: "Non NatAm ancestry"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, mmessage = F, warning = F)
```

```{r}
require(ggplot2)
require(cowplot)
require(scales)
```

# Results
## Addressing a trans-Pacific contact

The discovery of two individuals of Polynesian ancestry in the Botocudo
collection has led to several speculations about the origins and fate of those two individuals.
Their remains were presumably collected around the Doce River, 
a river spanning 853 kilometers in the state of Minas Gerais. 
It remains a mystery how they could have reached the Eastern part of Brazil, and 
whether they were the only Polynesians in the area at the time they were alive (that is, around or prior to 1500 AD).
Furthermore, it is unknown whether they could have mixed with local groups in Minas Gerais.

One of the questions we seek to address in this project is whether any of the 
newly sequenced individuals are admixed with groups out of the Americas.
Of particular interest is to know if we can detect recent admixture from Polynesian
populations into the Botocudo groups.

We performed admixture analyses for the 24 newly sequenced individuals and a panel of
$n$ worldwide populations (described in Wollstein et al., 2010 and Malaspinas et al., 2014),
including the two Polynesians found in the Botocudo collection in 2014.

As the depth of coverage is highly variable among these samples, the number of SNPs
with data per individual varied too.
Thus, we merged each of the target individuals separately to the reference panel, 
generating 24 different genotype likelihood matrices.
The number of SNPs varied from matrix to matrix, as we kept only sites were
there were data for the target individual.

NGSadmix was then run assuming K = 7 ancestral components, and 10 replicates per input matrix.
Based on these seven components, and to help with the discussion, 
we suggest a manual grouping of the 20 populations into 7 regions: 

* Africa
* Europe and West Asia
* East and Southeast Asia
* Central and South Asia
* Near Oceania
* Remote Oceania
* Americas

For all 24 individuals, at least 79% main ancestry is derived from a source population 
related to the people in the Americas included in this analysis. 
However, it is noteworthy to mention that **eleven** of the samples had a depth of coverage 
smaller than **0.05**$\times$ (and less than 10,000 SNPs with data), and **seven** of those
samples had a coverage smaller than **0.01**$\times$ (and less than 3,000 SNPs with data).


Remove outliers (DoC < threshold)

```{r}

make_model_plot <- function( x, xlab, title, color, model){
  
  summary_model <- summary( model )
  
  confidence_model <- predict( model, interval = "confidence")
  
  model$model$low <- confidence_model[,2]
  model$model$up <- confidence_model[,3]  
  
  r_sq <- round( summary_model$adj.r.squared, digits = 3 )
  p_val <- round(summary_model$coefficients[2, "Pr(>|t|)"], digits = 3 )
  slope <- round(summary_model$coefficients[2, "Estimate"], digits = 3 )
  
  model_plot  <- ggplot(model$model, aes( x = x , y = model$model[,1])) +
                        geom_point( col = color) +
                        theme_minimal() +
                        labs( title = title,
                              x = xlab, 
                              y = "Percentage of non-Native-American ancestry") +
                        geom_abline( intercept = summary_model$coefficients[1,1],
                                     slope = summary_model$coefficients[2,1],
                                     linetype = "dashed") +
                        geom_line(aes(y = low), linetype = "dashed", color = "gray") +
                        geom_line(aes(y = up), linetype = "dashed", color = "gray") +
                        scale_y_continuous(labels = percent) +
                        annotate("text", x = min(x), y = max(model$model[,1]),
                                      label =  paste("paste(italic(R) ^ 2, \" = \" ,", r_sq, 
                                                     ", \"; slope = \", ", slope,
                                                     ", \"; p-value = \", ", p_val,
                                                     ")"), 
                                 parse = TRUE,
                                 vjust = 0, hjust = 0
                                  )

  return( model_plot )
}

query_model_details <- function( model, parameter ){
  summary_model <- summary( model )
  if( parameter == "R^2"){
    my_value <- round( summary_model$adj.r.squared, digits = 3 )
  }else if( parameter == "p_value"){
    my_value <- round(summary_model$coefficients[2, "Pr(>|t|)"], digits = 3 )
  }else if( parameter == "slope"){
    my_value <- round(summary_model$coefficients[2, "Estimate"], digits = 3 )
  }
  return(my_value)
}

make_hist_plot <- function(data, x, title, xlab) {
  hist_plot <- ggplot( data, aes(x = x) ) +
    geom_histogram(bins = 10) +
    theme_minimal() + 
    labs(title = title, x  = xlab)
  return(hist_plot)
}
```

```{r residuals plot}
make_residuals_plot <- function(d, model, x, title, xlab){
  
  d$predicted <- predict( model )
  d$residuals <- residuals( model )
  
  res_plot <- ggplot(d, aes(x = x, y = non_NatAm)) +
                  geom_smooth(method = "lm", se = FALSE, color = "lightgrey") +
                  geom_segment(aes(xend = x, yend = predicted), alpha = .2) +
                  # > Color AND size adjustments made here...
                  geom_point(aes(color = abs(residuals), size = abs(residuals))) + # size also mapped
                  scale_color_continuous(low = "black", high = "red") +
                  guides(color = FALSE, size = FALSE) +  # Size legend also removed
                  # <
                  geom_point(aes(y = predicted), shape = 1) +
                  theme_bw() +
    labs(x = xlab, title = title)
  return(res_plot)
}


```

```{r simple model, fig.width=12, fig.height=12}
anc_err <- read.csv("~/Projects/Botocudos/Files/ADMIX/2020_06_04/non_NatAm_anc.txt")

index <- anc_err$non_NatAm  >= 0
error_model <- lm(non_NatAm  ~ error, anc_err[ index, ] )
cont_model <-  lm(non_NatAm ~ contamMix_est_all , anc_err[ index, ] ) 
doc_model <- lm(non_NatAm ~  DoC_endogenous , anc_err[ index, ] )  

simple_model <- data.frame(model = "Simplest model",
                           explanatory_variable = c( "Error rate", "DoC", "Contamination"),
                           r_sq = sapply(list(error_model, doc_model, cont_model), 
                                         function(model) query_model_details(model, parameter = "R^2")),
                           p_val = sapply(list(error_model, doc_model, cont_model), 
                                          function(model) query_model_details(model, parameter = "p_value")),
                           slope = sapply(list(error_model, doc_model, cont_model), 
                                          function(model) query_model_details(model, parameter = "slope")))

err_plot <- make_model_plot( x = error_model$model$error, 
                 xlab = "Error rate (%)", 
                 title = "Error rate",
                 color = "gold2",
                 model = error_model) + scale_x_continuous(labels = percent) 

doc_plot <- make_model_plot( x = doc_model$model$DoC_endogenous, 
                 xlab = "Depth of coverage", 
                 title = "Depth of coverage",
                 color = "deepskyblue1",
                 model = doc_model) 

cont_plot <- make_model_plot( x = cont_model$model$contamMix_est_all, 
                 xlab = "Contamination estimate", 
                 title = "Contamination",
                 color = "salmon",
                 model = cont_model) 

err_hist <- make_hist_plot(data = anc_err[index, ], 
                           x = anc_err$error[index], 
                           title = "Error rate", xlab = "Error rate (%)")
doc_hist <- make_hist_plot(data = anc_err[index, ], 
                           x = anc_err$DoC_endogenous[index], 
                           title = "Depth of coverage", xlab = "Depth of coverage")
cont_hist <- make_hist_plot(data = anc_err[index, ], 
                            x = anc_err$contamMix_est_all[index], 
                            title = "Contamination", xlab = "Contamination estimate")
error_res <- make_residuals_plot(d = anc_err[index,], 
                               model = error_model, 
                               x = log(anc_err[index,]$error),
                               title = "Error rate", 
                               xlab = "Error rate")

doc_res <- make_residuals_plot(d = anc_err[index,], 
                               model = doc_model, 
                               x = log(anc_err[index,]$DoC_endogenous),
                               title = "Depth of coverage", 
                               xlab = "Depth of coverage")

cont_res <- make_residuals_plot(d = anc_err[index,], 
                               model = cont_model, 
                               x = log(anc_err[index,]$contamMix_est_all),
                               title = "Contamination", 
                               xlab = "Contamination estimate")

plot_grid( err_hist, doc_hist, cont_hist, 
 err_plot, doc_plot, cont_plot, 
 error_res, doc_res, cont_res,
 nrow = 3)
```

```{r log model, fig.width=12, fig.height=12}
anc_err <- read.csv("~/Projects/Botocudos/Files/ADMIX/2020_06_04/non_NatAm_anc.txt")

index <- anc_err$non_NatAm  >= 0
error_model <- lm(non_NatAm  ~ error, anc_err[ index, ] )
cont_model <-  lm(non_NatAm ~ log( contamMix_est_all ), anc_err[ index, ] ) 
doc_model <- lm(non_NatAm ~ log( DoC_endogenous ), anc_err[ index, ] )  


log_model <- data.frame(model = "Log model",
                           explanatory_variable = c( "Error rate", "DoC (log)", "Contamination (log)"),
                           r_sq = sapply(list(error_model, doc_model, cont_model), 
                                         function(model) query_model_details(model, parameter = "R^2")),
                           p_val = sapply(list(error_model, doc_model, cont_model), 
                                          function(model) query_model_details(model, parameter = "p_value")),
                           slope = sapply(list(error_model, doc_model, cont_model), 
                                          function(model) query_model_details(model, parameter = "slope")))

err_plot <- make_model_plot( x = error_model$model$error, 
                 xlab = "Error rate (%)", 
                 title = "Error rate",
                 color = "gold2",
                 model = error_model) + scale_x_continuous(labels = percent) 

doc_plot <- make_model_plot( x = doc_model$model$`log(DoC_endogenous)`, 
                 xlab = "Depth of coverage (log)", 
                 title = "Depth of coverage",
                 color = "deepskyblue1",
                 model = doc_model) 

cont_plot <- make_model_plot( x = cont_model$model$`log(contamMix_est_all)`, 
                 xlab = "Contamination estimate (log)", 
                 title = "Contamination",
                 color = "salmon",
                 model = cont_model) 

err_hist <- make_hist_plot(data = anc_err[index, ], 
                           x = anc_err$error[index], 
                           title = "Error rate", xlab = "Error rate (%)")
doc_hist <- make_hist_plot(data = anc_err[index, ],
                           x = log(anc_err[index,]$DoC_endogenous),
                           title = "Depth of coverage",
                           xlab = "Depth of coverage")
cont_hist <- make_hist_plot(data = anc_err[index, ], 
                            x = log(anc_err$contamMix_est_all[index]),
                            title = "Contamination",
                            xlab = "Contamination estimate")

error_res <- make_residuals_plot(d = anc_err[index,], 
                               model = error_model, 
                               x = log(anc_err[index,]$error),
                               title = "Error rate", 
                               xlab = "Error rate")

doc_res <- make_residuals_plot(d = anc_err[index,], 
                               model = doc_model, 
                               x = log(anc_err[index,]$DoC_endogenous),
                               title = "Depth of coverage", 
                               xlab = "Depth of coverage")

cont_res <- make_residuals_plot(d = anc_err[index,], 
                               model = cont_model, 
                               x = log(anc_err[index,]$contamMix_est_all),
                               title = "Contamination", 
                               xlab = "Contamination estimate")

plot_grid( err_hist, doc_hist, cont_hist, 
 err_plot, doc_plot, cont_plot, 
 error_res, doc_res, cont_res,
 nrow = 3)


```

```{r log, removing outliers with DoC less than 0.005, fig.width=12, fig.height=12 }
anc_err <- read.csv("~/Projects/Botocudos/Files/ADMIX/2020_06_04/non_NatAm_anc.txt")

index <- anc_err$DoC_endogenous  >= 0.005 
error_model <- lm(non_NatAm  ~ error, anc_err[ index, ] )
cont_model <-  lm(non_NatAm ~ log( contamMix_est_all ), anc_err[ index, ] ) 
doc_model <- lm(non_NatAm ~ log( DoC_endogenous ), anc_err[ index, ] )  


log_005_model <- data.frame(model = "Log model, DoC >= 0.005",
                           explanatory_variable = c( "Error rate", "DoC (log)", "Contamination (log)"),
                           r_sq = sapply(list(error_model, doc_model, cont_model), 
                                         function(model) query_model_details(model, parameter = "R^2")),
                           p_val = sapply(list(error_model, doc_model, cont_model), 
                                          function(model) query_model_details(model, parameter = "p_value")),
                           slope = sapply(list(error_model, doc_model, cont_model), 
                                          function(model) query_model_details(model, parameter = "slope")))

err_plot <- make_model_plot( x = error_model$model$error, 
                 xlab = "Error rate (%)", 
                 title = "Error rate",
                 color = "gold2",
                 model = error_model) + scale_x_continuous(labels = percent) 

doc_plot <- make_model_plot( x = doc_model$model$`log(DoC_endogenous)`, 
                 xlab = "Depth of coverage (log)", 
                 title = "Depth of coverage",
                 color = "deepskyblue1",
                 model = doc_model) 

cont_plot <- make_model_plot( x = cont_model$model$`log(contamMix_est_all)`, 
                 xlab = "Contamination estimate (log)", 
                 title = "Contamination",
                 color = "salmon",
                 model = cont_model) 

err_hist <- make_hist_plot(data = anc_err[index, ], 
                           x = anc_err$error[index], 
                           title = "Error rate", xlab = "Error rate (%)")
doc_hist <- make_hist_plot(data = anc_err[index, ],
                           x = log(anc_err[index,]$DoC_endogenous),
                           title = "Depth of coverage",
                           xlab = "Depth of coverage")
cont_hist <- make_hist_plot(data = anc_err[index, ], 
                            x = log(anc_err$contamMix_est_all[index]),
                            title = "Contamination",
                            xlab = "Contamination estimate")

error_res <- make_residuals_plot(d = anc_err[index,], 
                               model = error_model, 
                               x = log(anc_err[index,]$error),
                               title = "Error rate", 
                               xlab = "Error rate")

doc_res <- make_residuals_plot(d = anc_err[index,], 
                               model = doc_model, 
                               x = log(anc_err[index,]$DoC_endogenous),
                               title = "Depth of coverage", 
                               xlab = "Depth of coverage")

cont_res <- make_residuals_plot(d = anc_err[index,], 
                               model = cont_model, 
                               x = log(anc_err[index,]$contamMix_est_all),
                               title = "Contamination", 
                               xlab = "Contamination estimate")

plot_grid( err_hist, doc_hist, cont_hist, 
 err_plot, doc_plot, cont_plot, 
 error_res, doc_res, cont_res,
 nrow = 3)

```

```{r log, removing outliers with DoC less than 0.01, fig.width=12, fig.height=12}
anc_err <- read.csv("~/Projects/Botocudos/Files/ADMIX/2020_06_04/non_NatAm_anc.txt")

index <- anc_err$DoC_endogenous  >= 0.01 
error_model <- lm(non_NatAm  ~ error, anc_err[ index, ] )
cont_model <-  lm(non_NatAm ~ log( contamMix_est_all ), anc_err[ index, ] ) 
doc_model <- lm(non_NatAm ~ log( DoC_endogenous ), anc_err[ index, ] )  


log_01_model <- data.frame(model = "Log model, DoC >= 0.01",
                           explanatory_variable = c( "Error rate", "DoC (log)", "Contamination (log)"),
                           r_sq = sapply(list(error_model, doc_model, cont_model), 
                                         function(model) query_model_details(model, parameter = "R^2")),
                           p_val = sapply(list(error_model, doc_model, cont_model), 
                                          function(model) query_model_details(model, parameter = "p_value")),
                           slope = sapply(list(error_model, doc_model, cont_model), 
                                          function(model) query_model_details(model, parameter = "slope")))

err_plot <- make_model_plot( x = error_model$model$error, 
                 xlab = "Error rate (%)", 
                 title = "Error rate",
                 color = "gold2",
                 model = error_model) + scale_x_continuous(labels = percent) 

doc_plot <- make_model_plot( x = doc_model$model$`log(DoC_endogenous)`, 
                 xlab = "Depth of coverage (log)", 
                 title = "Depth of coverage",
                 color = "deepskyblue1",
                 model = doc_model) 

cont_plot <- make_model_plot( x = cont_model$model$`log(contamMix_est_all)`, 
                 xlab = "Contamination estimate (log)", 
                 title = "Contamination",
                 color = "salmon",
                 model = cont_model) 

err_hist <- make_hist_plot(data = anc_err[index, ], 
                           x = anc_err$error[index], 
                           title = "Error rate", xlab = "Error rate (%)")
doc_hist <- make_hist_plot(data = anc_err[index, ],
                           x = log(anc_err[index,]$DoC_endogenous),
                           title = "Depth of coverage",
                           xlab = "Depth of coverage")
cont_hist <- make_hist_plot(data = anc_err[index, ], 
                            x = log(anc_err$contamMix_est_all[index]),
                            title = "Contamination",
                            xlab = "Contamination estimate")

error_res <- make_residuals_plot(d = anc_err[index,], 
                               model = error_model, 
                               x = log(anc_err[index,]$error),
                               title = "Error rate", 
                               xlab = "Error rate")

doc_res <- make_residuals_plot(d = anc_err[index,], 
                               model = doc_model, 
                               x = log(anc_err[index,]$DoC_endogenous),
                               title = "Depth of coverage", 
                               xlab = "Depth of coverage")

cont_res <- make_residuals_plot(d = anc_err[index,], 
                               model = cont_model, 
                               x = log(anc_err[index,]$contamMix_est_all),
                               title = "Contamination", 
                               xlab = "Contamination estimate")

plot_grid( err_hist, doc_hist, cont_hist, 
 err_plot, doc_plot, cont_plot, 
 error_res, doc_res, cont_res,
 nrow = 3)

```



```{r table with details of the models}
all_models <- do.call(rbind, list( simple_model, log_model, log_01_model, log_005_model))

all_models$model <- factor(all_models$model, levels = c("Simplest model",
                                                        "Log model",
                                                        "Log model, DoC >= 0.005",
                                                        "Log model, DoC >= 0.01"))

all_models$explanatory_variable <- factor(all_models$explanatory_variable,
                                          levels = c("Error rate", "DoC", "DoC (log)", "Contamination", "Contamination (log)"))

knitr::kable(all_models)
```

