---
title: "Supplementary tables for mapping statistics and contamination"
author: "DI Cruz Dávalos"
date: "1/23/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = F, warning = F)
```

```{r}
require(plyr)
require(ggplot2)
require(cowplot)
require(tidyr)
require(scales)
require(knitr)
```

```{r}
#------------------------------------------------------------------------------#
# Load isotopic data if available
backbone <- read.csv("~/Projects/Botocudos/Files/Tables/Backbone_Table1_2020.csv")
colnames(backbone) <- c("Sample", "CalibratedDate", "Area", "State") 

```

```{r read in file}
myPath <- "~/Projects/Botocudos/Files/Summaries/2019_11_21/"
dp <- "sample_depth.hg19.csv"
st <- "sample_stats.hg19.csv"

whole_dp <- read.csv(paste(myPath, "Whole/", dp, sep = ""))
whole_dp$SM <- sub(".hg19", "", whole_dp$Sample)
whole_st <- read.csv(paste(myPath, "Whole/", st, sep = ""))

trim_dp <- read.csv(paste(myPath, "Trimmed/", dp, sep = ""))
trim_dp$SM <- sub(".hg19", "", trim_dp$Sample)
trim_st <- read.csv(paste(myPath, "Trimmed/", st, sep = ""))

whole <- join(whole_dp, whole_st, by = "SM")
whole$type <- "whole"
trim <- join(trim_dp, trim_st, by = "SM")
trim$type <- "trimmed"


myStats <- rbind(trim, whole)
myStats$type <- factor(myStats$type, levels = c("whole", "trimmed"), ordered = T)
myStats <- myStats[order(myStats$endo_final_prop),]
myStats$SM <- factor(myStats$SM, levels = unique(myStats$SM), ordered = T)
```

## Supplementary Table 1:  Mapping stats per sample

```{r define column names}

myColNames <- c("Sample",            "Sample",                          "National Museum of Brazil (Museu Nacional, MN) identifier",
                
                "SM",            "Sample",                          "National Museum of Brazil (Museu Nacional, MN) identifier",
                "library",                        "Library", "DNA library identifier",
                "LB",                        "Library", "DNA library identifier",
                
                "Area",              "Region",                          "Sample origin (region) according to MN's archives",
                "State",             "State",                           "Sample origin (state) according to MN's archives", 
                "CalibratedDate",    "Calibrated date (AD)",            "OxCal estimated mean and standard deviation for calibrated dates",
                
                "reads_raw",         "sequenced_reads",                 "Number of sequenced reads", 
                "reads_trim",        "retained_reads_trimming",         "Number of reads retained after trimming",
                
                "duplicates",        "duplicated_reads",                "Number of reads flagged as PCR duplicates by picardtools MarkDuplicates",
                "duplicates_prop",   "percent_duplicated_reads",        "Percentage of reads flagged as PCR duplicates by picardtools MarkDuplicates",
                 
                "mapping_final",     "unique_reads_endogenous",         "Number of unique reads mapped to the human genome (build 37.1)",
                "endo_prop",         "percent_unique_reads_endogenous", "Percentage of retained reads that were uniquely mapped to the human genome",
                
                "AvgReadDepth_tot",  "DoC_endogenous",                  "Depth of coverage on human genome (number of unique bases divided by the length of the human genome reference build 37.1)",
                "AvgReadDepth_MT",   "DoC_MT",                          "Depth of coverage on mitochondrial genome (number of unique bases divided by the length of the mitochondrial genome, build 37.1)",
                
                "AvgReadLength",     "length_reads_endogenous",         "Average length of uniquely mapped reads",
                
                "R_y",               "Ry_95CI",                          "Ratio of the reads mapping to the Y-chromosome and the reads mapping to sexual chromosomes (X and Y) and 95% confidence interval",
                "Sex",               "Sex",                              "Molecular sex determined by the Ry ratio"
                )

change_name <- function(name){
  new_name <- myColNames[which(myColNames == name)[1] + 1]
  ifelse(is.na(new_name), return(name), return(new_name))
}

```

```{r prepare caption}
myCaption <- paste(paste(myColNames[seq(2, length(myColNames), 3)], 
                   myColNames[seq(3, length(myColNames), 3)], sep = ": "),
                   collapse = "\n\n")
```

```{r prepare numbers format}
# colname, accuracy, percent
myNumFormat <- matrix(c("sequenced_reads", 1, 0, 
                 "retained_reads_trimming", 1, 0,
                 "duplicated_reads", 1, 0,
                 "percent_duplicated_reads", 3, 1,
                 "unique_reads_endogenous", 1, 0,
                 "percent_unique_reads_endogenous", 4, 1,
                 "DoC_endogenous", 3, 0,
                 "DoC_MT", 1, 0,
                 "length_reads_endogenous", 1, 0), ncol = 3, byrow = T)

format_column_num <- function(col_name, percent, accuracy, myT1){
  if(percent){
    myT1[,col_name]  <- percent(myT1[,col_name], accuracy = accuracy)
  }else{
    myT1[,col_name] <- round(myT1[,col_name], accuracy)
  }
  myT1[,col_name] <- prettyNum(myT1[,col_name], big.mark = ",")
  return(myT1)
}

```

```{r prepare table}
myT1 <- myStats[myStats$type == "whole",]

myT1$R_y <-  paste(myT1$R_y, " ( ", myT1$X95.CI, ")", sep = "")

myT1 <- myT1[,c("Sample", "reads_raw", "reads_trim", "trim_prop", 
                "duplicates", "duplicates_prop", "mapping_final",
                "endo_prop", "AvgReadDepth_tot", "AvgReadDepth_MT", "AvgReadLength",
                "Sex", "R_y")]

myT1 <- join(backbone, myT1, by = "Sample")

colnames(myT1) <- sapply(colnames(myT1), function(name) change_name(name))

myT1$percent_duplicated_reads <- myT1$percent_duplicated_reads/100
myT1$percent_unique_reads_endogenous <- myT1$percent_unique_reads_endogenous/100

samples_order <- myT1$Sample[order(myT1$DoC_endogenous)]

myT1$Sample <- factor(myT1$Sample, levels = samples_order, ordered = T)
myT1 <- myT1[order(myT1$Sample, decreasing = T),]
```

```{r format prettyNum}
for(i in seq(1, nrow(myNumFormat))){
  myT1 <- format_column_num(myNumFormat[i, 1], 
                            as.numeric(myNumFormat[i,3]), 
                            as.numeric(myNumFormat[i,2]),
                            myT1)
}
```

```{r}
kable(myT1, caption = myCaption, digits = 1, row.names = F)
```

```{r}
# add plot for merged libraries
```


## Supplementary Table 2: Mapping stats per library

```{r prepare number format}
myNumFormat <- matrix(c("sequenced_reads", 1, 0, 
                 "retained_reads_trimming", 1, 0,
                 "duplicated_reads", 1, 0,
                 "percent_duplicated_reads", 3, 1,
                 "unique_reads_endogenous", 1, 0,
                 "percent_unique_reads_endogenous", 4, 1
                 ), ncol = 3, byrow = T)
```

```{r}
lib_st <- read.csv("~/Projects/Botocudos/Files/Summaries/2019_11_21/Whole/library_stats.hg19.csv")
whole_st$LB <- "All"

myT2 <- rbind(whole_st, lib_st)
myT2 <- myT2[order(myT2$SM, myT2$LB),]
myT2$LB <- factor(myT2$LB, levels = c("mtCapture", "L3U", "L2", "L1", "All"),
                  ordered = T)
myT2$SM <- factor(myT2$SM, levels = samples_order, ordered = T)
myT2 <- myT2[order(myT2$SM, myT2$LB, decreasing = T),]



myT2 <- myT2[,c("SM", "LB","reads_raw", "reads_trim", "trim_prop", 
                "duplicates", "duplicates_prop", "mapping","mapping_final",
                "endo_prop")]

myT2_final <- myT2
myT2_final <- myT2_final[, c("SM", "LB", "reads_raw", "reads_trim", "trim_prop",
                             "duplicates", "duplicates_prop", "mapping_final", "endo_prop")]
colnames(myT2_final) <- sapply(colnames(myT2_final), function(name) change_name(name))
myT2_final$percent_duplicated_reads <- myT2_final$percent_duplicated_reads/100
myT2_final$percent_unique_reads_endogenous <- myT2_final$percent_unique_reads_endogenous/100

for(i in seq(1, nrow(myNumFormat))){
  myT2_final <- format_column_num(myNumFormat[i, 1], 
                            as.numeric(myNumFormat[i,3]), 
                            as.numeric(myNumFormat[i,2]),
                            myT2_final)
}
```


```{r}
kable(myT2_final, row.names = F, digits = 1)

```


Figure for internal discussion; will we add any figure for SI?

```{r, fig.width=14, fig.height=14}
stats_to_plot <- myT2 %>% 
  mutate(log_reads_raw = log10(reads_raw),
         log_reads_trim = log10(reads_trim),
         log_mapping = log10(mapping),
         log_duplicates = log10(duplicates),
         log_mapping_final = log10(mapping_final),
         log_endo_prop = log10(endo_prop)) %>% 
  gather( name, value,  -SM, -LB) 


stats_to_plot$name <- factor(stats_to_plot$name, levels = c("reads_raw",
                                                            "log_reads_raw",
                                                            "reads_trim",
                                                            "log_reads_trim",
                                                            "trim_prop",
                                                            "mapping",
                                                            "log_mapping",
                                                            "duplicates",
                                                            "log_duplicates",
                                                            "duplicates_prop",
                                                            "mapping_final",
                                                            "log_mapping_final",
                                                            "Nseqs",
                                                            "log_Nseqs",
                                                            "NchrY.NchrX",
                                                            "log_NchrY.NchrX",
                                                            "NchrY",
                                                            "log_NchrY",
                                                            "R_y",
                                                            "endo_prop",
                                                            "log_endo_prop",
                                                            "endo_final_prop",
                                                            "log_endo_final_prop",
                                                            "AvgReadDepth_tot",
                                                            "log_AvgReadDepth_tot",
                                                            "AvgReadDepth_MT",
                                                            "log_AvgReadDepth_MT",
                                                            "AvgReadLength"),
                             ordered = T)

####
# y-axis to meaningful numbers (i.e., 10^10, 10^4, etc...)

# one plot in linear scale, another in log

stats_to_plot %>%
ggplot(  aes(x = SM, y = value, fill = LB)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 90, size = 10),
        strip.text.x = element_text(size = 12)) +
  facet_wrap(facets = vars(name), scales = "free_y", ncol = 3) +
   # scale_fill_discrete() +
  scale_fill_manual(values =  c("#c4bf33","#0d75ff",
                                "#e0b81f","#b08699",
                                 "gray40"
                                )) +
  labs(x = NULL, y = NULL)

```


### Length

Lots of plots per library, merged in one pdf

```{r load files per sample}
myPath <- "~/Projects/Botocudos/Files/bamdamage/2020_01_23/Whole/"

samples <- as.character(unique(lib_st$SM))
if(!"MN0008_L3U" %in% samples){
  samples <- c(samples, "MN0008_L3U", "MN0008_non_U")  
}
myLength <- vector("list", length = length(samples))
names(myLength) <- samples

for(s in samples){
  path_length <- paste(myPath, s, sep = "/")
  path_length <- paste(path_length, "/", s, ".hg19.length.csv", sep = "")
  tmp_length <- read.csv(path_length)
  myLength[[s]] <- tmp_length
  }   


```

```{r plots per sample, eval = T, fig.width=16, fig.height=16}
myColor <- "royalblue"
nCol <- 6 
# samples_per_row <- 5
cex <- 1.2

blank_plot <- function(){
  plot(1,1, type = "n", bty = "n", axes = F, xlab = NA, ylab = NA)
}

# pdf("~/Projects/Botocudos/Plots/Sequencing/Length_dist.pdf",
    # width = 9, height = 12)
layout(matrix(seq(1, nCol*nCol), ncol = nCol, byrow = T))

for(s in samples){
    le <- myLength[[s]]
    le <- le[le$counts >0,]
    plot.title <- s
    plot(x = le$length, y = le$counts/sum(le$counts), type = "b",
         col = myColor, bty = "n", 
         xlab = "Length", ylab = "Frequency", lwd = 1.5, 
         main = plot.title, cex = cex)
}

# dev.off()
```

```{r load files per library}
myPath <- "~/Projects/Botocudos/Files/bamdamage/2020_01_23/Whole/"

myLength <- vector("list", length = length(samples))
names(myLength) <- samples

for(s in samples){
  myLibs <- lib_st$LB[lib_st$SM == s]
  some_length <- vector("list", length = length(myLibs))
  names(some_length) <- myLibs
  
  for(l in myLibs){
    path_length <- paste(myPath, s, l, "library_bamdamage", sep = "/")
    path_length <- paste(path_length, "/", l, ".hg19.length.csv", sep = "")
    tmp_length <- read.csv(path_length)
    some_length[[l]] <- tmp_length
  }
  myLength[[s]] <- some_length
  }


```

```{r plots per library, eval = F}
myColor <- "royalblue"
nCol <- max(sapply(myLength, length))
samples_per_page <- 5
cex <- 1.2

blank_plot <- function(){
  plot(1,1, type = "n", bty = "n", axes = F, xlab = NA, ylab = NA)
}

pdf("~/Projects/Botocudos/Plots/Sequencing/Length_dist.pdf",
    width = 9, height = 12)
layout(matrix(seq(samples_per_page*nCol), nrow = samples_per_page, byrow = T))

for(s in samples){
  myLibs <- lib_st$LB[lib_st$SM == s]
  myRow <- which(samples == s)
  
  if(!myRow %% samples_per_page){
    # plot.new()
    layout(matrix(seq(samples_per_page*nCol), nrow = samples_per_page, byrow = T))  
  }
  
  for(l in myLibs){
    le <- myLength[[s]][[l]]
    le <- le[le$counts >0,]
    plot.title <- paste(s, l)
    plot(x = le$length, y = le$counts/sum(le$counts), type = "b",
         col = myColor, bty = "n", 
         xlab = "Length", ylab = "Frequency", lwd = 1.5, 
         main = plot.title, cex = cex)
  }
  plots_left <- nCol - length(myLibs)
  while(plots_left > 0){
    blank_plot()
    plots_left <- plots_left - 1
  }
}

dev.off()
```

### Damage

Lots of plots per library, merged in one pdf


```{r load damage per sample}
myPath <- "~/Projects/Botocudos/Files/bamdamage/2020_01_23/Whole"

if(!"MN0008_L3U" %in% samples){
  samples <- c(samples, "MN0008_L3U", "MN0008_non_U")  
}

myDam <- vector("list", length = length(samples))
names(myDam) <- samples

for(s in samples){
  path_dam <- paste(myPath, s, sep = "/")
  path_dam <- paste(path_dam, "/", s, ".hg19.dam_5prime.csv", sep = "")
  tmp_dam5 <- read.csv(path_dam)
  path_dam <- sub("5prime", "3prime", path_dam)
  tmp_dam3 <- read.csv(path_dam)
  
  some_dam <- list(five_prime = tmp_dam5, three_prime = tmp_dam3)
  myDam[[s]] <- some_dam
}


```

```{r plot damage per sample, eval=T, fig.width=16, fig.height=16}
myColor_CT <- "royalblue"
myColor_GA <- "firebrick3"
nCol <- 8
samples_per_page <- 5
samples <- sort(samples)
cex <- 1.2

blank_plot <- function(){
  plot(1:10,1:10, type = "n", bty = "n", axes = F, xlab = NA, ylab = NA)
}

# pdf("~/Projects/Botocudos/Plots/Sequencing/Length_dist.pdf",
    # width = 9, height = 12)
layout(matrix(seq(nCol*nCol), ncol = nCol, byrow = T))

for(s in samples){
  
    y2 <- max(c( myDam[[s]][["five_prime"]]$C..T,  
                 myDam[[s]][["three_prime"]]$G..A))
    le <- myDam[[s]][["five_prime"]]
    
    plot.title <- s
    par(mar = c(4, 4, 3, 1))
    plot(le$C..T[1:25], type = "b",
         col = myColor_CT, bty = "n", 
         xlab = "5' position", ylab = "Frequency", lwd = 1.5, 
         main = plot.title, cex = cex, axes = F, ylim = c(0, y2))
    axis(side = 1, at = seq(0,25,5))
    axis(side=2)
    
    le <- myDam[[s]][["three_prime"]]
    par(mar = c(4,1,3,3))
    plot(rev(le$G..A[1:25]), type = "b",
         col = myColor_GA, bty = "n", 
          lwd = 1.5, 
         cex = cex, axes = F, xlab = "3' position", ylab = NA, ylim = c(0, y2))
     axis(side = 1, at = seq(1,25,5), labels = seq(25, 1,-5))
     axis(side = 4)
     

}

blank_plot()
legend(1,10, legend = c("C -> T", "G -> A"), col = c(myColor_CT, myColor_GA),
       lty = 1, pch = 21, bty = "n", cex = 2)

# dev.off()
```


```{r load damage per library}
myPath <- "~/Projects/Botocudos/Files/bamdamage/2020_01_23/Whole/"
samples <- unique(lib_st$SM)
myDam <- vector("list", length = length(samples))
names(myDam) <- samples

for(s in samples){
  myLibs <- lib_st$LB[lib_st$SM == s]
  some_dam <- vector("list", length = length(myLibs))
  names(some_dam) <- myLibs
  
  for(l in myLibs){
    path_dam <- paste(myPath, s, l, "library_bamdamage", sep = "/")
    path_dam <- paste(path_dam, "/", l, ".hg19.dam_5prime.csv", sep = "")
    tmp_dam5 <- read.csv(path_dam)
    path_dam <- sub("5prime", "3prime", path_dam)
    tmp_dam3 <- read.csv(path_dam)
    
    some_dam[[l]] <- list(five_prime = tmp_dam5, three_prime = tmp_dam3)
  }
  myDam[[s]] <- some_dam
  }


```

```{r plot damage per library, eval=F}
myColor_CT <- "royalblue"
myColor_GA <- "firebrick3"
nCol <- max(sapply(myDam, length))
samples_per_page <- 5
cex <- 1.2

blank_plot <- function(){
  plot(1:10,1:10, type = "n", bty = "n", axes = F, xlab = NA, ylab = NA)
}

# pdf("~/Projects/Botocudos/Plots/Sequencing/Length_dist.pdf",
    # width = 9, height = 12)
layout(matrix(seq(samples_per_page*nCol), nrow = samples_per_page, byrow = T))

for(s in samples){
  myLibs <- lib_st$LB[lib_st$SM == s]
  myRow <- which(samples == s)
  
  if(!myRow %% samples_per_page){
    # plot.new()
    layout(matrix(seq(samples_per_page*nCol*2), nrow = samples_per_page, byrow = T))  
  }
  
  for(l in myLibs){
    y2 <- max(c( myDam[[s]][[l]][["five_prime"]]$C..T,  
                 myDam[[s]][[l]][["three_prime"]]$G..A))
    le <- myDam[[s]][[l]][["five_prime"]]
    
    plot.title <- paste(s, l)
    par(mar = c(4, 3, 3, 1))
    plot(le$C..T[1:25], type = "b",
         col = myColor_CT, bty = "n", 
         xlab = "5' position", ylab = "Frequency", lwd = 1.5, 
         main = plot.title, cex = cex, axes = F, ylim = c(0, y2))
    axis(side = 1, at = seq(0,25,5))
    axis(side=2)
    
    le <- myDam[[s]][[l]][["three_prime"]]
    par(mar = c(4,1,3,3))
    plot(rev(le$G..A[1:25]), type = "b",
         col = myColor_GA, bty = "n", 
          lwd = 1.5, 
         cex = cex, axes = F, xlab = "3' position", ylab = NA, ylim = c(0, y2))
     axis(side = 1, at = seq(1,25,5), labels = seq(25, 1,-5))
     axis(side = 4)
  }
  plots_left <- nCol*2 - length(myLibs)*2
  while(plots_left > 0){
    blank_plot()
    plots_left <- plots_left - 1
  }
  legend(1,10, legend = c("C -> T", "G -> A"), col = c(myColor_CT, myColor_GA),
       lty = 1, pch = 21, bty = "n", cex = 2)
}

# blank_plot()


# dev.off()
```


## Supplementary Table 3: Contamination estimates

!!!
cleaning and gathering output from schmutzi, contaminationX and contamMix
!!!

```{r add results from contaminationX}
#------------------------------------------------------------------------------#
## X
minDepth <- seq(2,3)
path <- "~/Projects/Botocudos/Files/Contamination/2020_01_24/contaminationX/"
mySamples <- unique(myT1$Sample)
libs <- c("All")

s <- mySamples[1]
l <- libs[1]
md <- minDepth[2]

empty_df <- function(sample, lib){
  myDF <- data.frame(method = c("One-cns", "Two-cns"),
                     estimate = c(NA, NA), lb = c(NA, NA), ub = c(NA, NA),
                     err = c(NA, NA), nSites = c(NA, NA))
  return(myDF)
}

x_cont <- data.frame()
for(s in mySamples){
  for(l in libs){
    for(md in minDepth){
      
      tmp_res <- tryCatch(tmp_res <- read.table(paste(path, 
                                                      s,".hg19_", l, 
                                                      "_md", md, ".result", sep = "")),
                          error =  function(e) empty_df(s, l),
                          finally = function() return(tmp_res))
      
      colnames(tmp_res) <- c("method", "estimate", "lb", "ub", "err", "nSites")
      tmp_res$Sample <- s
      tmp_res$library <- l
      tmp_res$depth <- md
      x_cont <- rbind(x_cont, tmp_res)
    }
  }
}

x_cont <- join(x_cont, myT1[,c("Sex", "Sample")], by = "Sample")
x_cont <- x_cont[!is.na(x_cont$estimate) & x_cont$Sex %in% c("XY", "Not Assigned", "consistent with XY but not XX"),]

males <- x_cont[x_cont$method == "Two-cns",]

# write.table(males, "~/Projects/Botocudos/Files/Contamination/2020_01_24/contaminationX/estimates_X_males.txt", 
# sep = "\t", quote = F, col.names = T, row.names = F)

males_fmt <- males

males_fmt$estimate <- percent(males_fmt$estimate)
males_fmt$lb <- percent(males_fmt$lb)
males_fmt$ub <- percent(males_fmt$ub)

colnames(males_fmt) <- c("X_method", "X_estimate", "X_low_est", "X_upper_est",
                         "X_error", "X_num_sites", "sample", "library", "X_min_depth",
                         "sex")


```

```{r, eval = F}
kable(males_fmt, row.names = F)
```

```{r add results from contammix}
prefix <- "~/Projects/Botocudos/Files/Contamination/2020_01_24/contamMix/"
samples <- samples_order
rmTrans <- c("all", "rmTrans")

# We ran the chains for 100k iterations, 
# we will subsample a fraction of these samples to plot the posterior distribution

subsample_contammmix <- function(path, sample, rmTrans, size = 100, lib = "All",
                                 burnIn = 0.1){

    myRda <- paste(path, sample, "/", s, ".hg19_", lib, "_", rmTrans, ".Rdata", sep = "")

  print(myRda)
  load(myRda)
  notBurnIn <- as.integer(burnIn*size)
  tiny <- # sample(
    #remove burnin
    c(res$chains[[1]][size - ((size- notBurnIn):0),2])#,
                 #res$chains[[2]][,2], res$chains[[3]][,2])#, size)
  result <- data.frame(estimate = tiny, damage = rmTrans, lib = lib,
                       sample = sample)
  return(result)
}

# contammix <- data.frame()
# for(s in samples){
#   libs <- (myT2$LB[myT2$SM == s])
#   for(l in libs){
#     for(r in rmTrans){
#       tmp_contam <- subsample_contammmix(prefix, s, r, 100000, l)
#       contammix <- rbind(contammix, tmp_contam)
#     }
#   }
# }


#------------------------------------------------------------------------------#
# Order libraries and save (or load) results
# contammix$lib <- factor(contammix$lib,
#                         levels = c("L1", "L2", "L3U", "mtCapture","All"), 
#                         ordered = T)
# contammix$coverage <- 0
# for(s in unique(contammix$sample)){
#   for(l in unique(contammix$lib[contammix$sample == s])){
#     contammix$coverage <- boto$hits_coverage_mitochondrial[boto$sample == s & boto$library == l]
#   }
# }

# save(contammix, file="~/Projects/Botocudos/Files/Contamination/2020_01_24/contamMix//contamination.Rdata")
load("~/Projects/Botocudos/Files/Contamination/2020_01_24/contamMix/contamination.Rdata")


path <- "~/Projects/Botocudos/Files/Contamination/2020_01_24/contamMix/"
rmTrans <- c("all", "rmTrans")
annot_mito <- data.frame()

for(s in samples){
  libs <- (myT2$LB[myT2$SM == s])
  for(lib in libs){
    for(r in rmTrans){
      myRda <- paste(path, s, "/", s, ".hg19_", lib, "_", r, ".Rdata", sep = "")
      #print(myRda)
      load(myRda)
      nReads <- dim(res$mnMatrix)[1]
      # coverage <- boto$hits_coverage_mitochondrial[boto$sample == s & boto$library == lib]
      annot_tmp <- data.frame(nReads = nReads, #coverage = coverage, 
                              damage = r, lib = lib,
                           sample = s)
      
      annot_mito <- rbind(annot_mito, annot_tmp)
    }
  }
}



MAP <- read.csv("~/Projects/Botocudos/Files/Contamination/2020_01_24/contamMix/estimates.txt", header = F)

colnames(MAP) <- c("sample", "lib", "damage","map", "low", "high")
MAP$lib <- sub("S1", "L1", MAP$lib)
annot_mito <- join(annot_mito, MAP, by = c("sample", "lib", "damage"))

```

```{r print estimates from contamMix, eval = F}

kable(mito, row.names = F)
```

```{r add results from schmutzi}
path <- "~/Projects/Botocudos/Files/Contamination/2020_01_24/schmutzi/"

schmutzi_files <- list.files(path, pattern =".est")

schmutzi <- data.frame()

for(f in schmutzi_files){
  tmp_schmutzi <- read.table(paste(path, f, sep =""))
  colnames(tmp_schmutzi) <- c("estimate", "low", "up")
  tmp_schmutzi$sample <- strsplit(f, "_")[[1]][1]
  schmutzi <- rbind(schmutzi, tmp_schmutzi)
  
}
```

```{r print estimates from schmutzi, eval = F}
kable(schmutzi, row.names = F)
```

```{r}
myAccuracy <- 1

# contamMix
mito <- annot_mito[annot_mito$lib == "All" & annot_mito$damage == "all",]
mito$damage <- NULL
mito$lib <- NULL
colnames(mito) <- c("contamMix_num_reads", "Sample", 
                    "contamMix_estimate", "contamMix_low", "contamMix_up")

mito$contamMix_estimate <- percent(1 - mito$contamMix_estimate, accuracy = myAccuracy)
mito$contamMix_low <- percent(1 - mito$contamMix_low, accuracy = myAccuracy)
mito$contamMix_up <- percent(1 - mito$contamMix_up, accuracy = myAccuracy)
mito$contamMix_estimate <- paste(mito$contamMix_estimate, 
                                 " (", mito$contamMix_up, 
                                 ", ", mito$contamMix_low, ")", sep = "")
mito <- mito[, c("Sample", "contamMix_num_reads", "contamMix_estimate")]


# schmutzi
mito_sch <- schmutzi
colnames(mito_sch ) <- c("schmutzi_estimate", "schmutzi_low", "schmutzi_up", "Sample")
mito_sch$schmutzi_estimate <- percent(mito_sch$schmutzi_estimate, accuracy = myAccuracy)
mito_sch$schmutzi_low <- percent(mito_sch$schmutzi_low, accuracy = myAccuracy)
mito_sch$schmutzi_up <- percent(mito_sch$schmutzi_up, accuracy = myAccuracy)
mito_sch$schmutzi_estimate <- paste(mito_sch$schmutzi_estimate, 
                                    " (", mito_sch$schmutzi_low, 
                                    ", ",mito_sch$schmutzi_up, ")", sep = "")
mito_sch <- mito_sch[,c("Sample", "schmutzi_estimate")]

# contaminationX
x <- males_fmt[males_fmt$X_min_depth == 3,]
x$library <- NULL
x$X_method <- NULL
x$sex <- NULL
x$X_error <- NULL
x$X_min_depth <- NULL
colnames(x) <- c("contaminationX_estimate", "contaminationX_low", 
                 "contaminationX_up", "contaminationX_num_sites", "Sample")
x$contaminationX_estimate <-  paste(x$contaminationX_estimate, 
                                    " (", x$contaminationX_low, 
                                    ", ",x$contaminationX_up, ")", sep = "")
x <- x[,c("Sample", "contaminationX_estimate")]

# Merge
cont <- join(myT1[,c("Sample", "Sex")], mito, by = "Sample")
cont <- join(cont, mito_sch, by = "Sample")
cont <- join(cont, x, by = "Sample")

kable(cont, row.names = F)
```

## Supplementary Table 4: Haplogroups?

!!!
Haplogroups called (whenever it was possible)
cleaning and gathering output from haplogrep and David's script
!!!

```{r parse mt-haplogroups}
myPath <- "~/Projects/Botocudos/Files/Haplogroups/2020_01_23/"
hap_files <- list.files(myPath, pattern = ".haplo$")
mt_hap <- data.frame()

for(f in hap_files){
  tmp_hap <- read.table(paste(myPath, f, sep = ""), header = T)
  s <- sub(".haplo", "", f)
  tmp_hap$Sample <- s
  mt_hap <- rbind(mt_hap, tmp_hap)
  }

mt_hap <- mt_hap[,c("Sample", "Haplogroup")]


y_hap <- data.frame(Sample = c("MN0003",
                               "MN00064",
                               "MN00067",
                               "MN0008",
                               "MN0009",
                               "MN00346"),
                    Haplogroup = c("Q1a2a1-L54",
                                   "Q1a2a1a1-M3",
                                   "Q1a2a1a1-M3",
                                   "Q1a2a1a1-M3",
                                   "Q1a2a1a1-M3",
                                   "Q1a2a1a1-M3"))

myHaplos <- join(myT1[,c("Sample", "Sex")], mt_hap, by = "Sample")
myHaplos <- join(myHaplos, y_hap, by = "Sample")

kable(myHaplos, row.names = F)
```
