---
title: "Figures_MidThesis"
author: "Cruz-Dávalos, Diana I."
date: "5/20/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
library(plyr)
library(RColorBrewer)
library(colorspace)
library(maps)
```

```{r}
# Read in some files I will use for most of the figures
panel <- read.table("~/Projects/Botocudos/Files/Panels/Magic.panel", header = T)
panel$cd
coordinates <- read.table("~/Projects/Botocudos/Files/Maps/Magic.coordinates", header = F)
colnames(coordinates) <- c("population", "lat", "long", "type")
head(panel) ; head(coordinates)
languages <- read.table("~/Projects/Botocudos/Files/Panels/Maanasa/Maanasa.americas.Moreno_names.txt", header = F)
colnames(languages) <- c("population", "language")

```
```{r cmyk and colors}
#' assumes integer input for CMYK
cmyk <- function(C,M,Y,K) {

  C <- C / 100.0
  M <- M / 100.0
  Y <- Y / 100.0
  K <- K / 100.0

  n.c <- (C * (1-K) + K)
  n.m <- (M * (1-K) + K)  
  n.y <- (Y * (1-K) + K)

  r.col <- ceiling(255 * (1-n.c))
  g.col <- ceiling(255 * (1-n.m))
  b.col <- ceiling(255 * (1-n.y))

  x <- col2rgb(sprintf("#%02s%02s%02s",
                     as.hexmode(r.col), 
                     as.hexmode(g.col), 
                     as.hexmode(b.col)))

return(rgb(t(x)/255))
}

# Choose 11 colors
# 3 colors for the Asias
cobaltGreen <- cmyk(42,0,42,0) 
rainetteGreen<-cmyk(42,20,62,28) ; salviaBlue<-cmyk(41,25,10,0)
# Europe, neareast and caucasus
grayishLavender<-cmyk(28,28,0,0)
helvetiaBlue <- cmyk(100,62,19,10) ; aconiteViolet <-cmyk(39,68,5,0)
# 
blue <-cmyk(95,54, 0,0) ; yellowOrange <-cmyk(0,33,100,0)
benzolGreen <- cmyk(100,15,55,0); seashellPink<-cmyk(0,19,23,0)

```

```{r cmyk function}
x <- 1
y <- 2

plotwclus <- function(x,y){
  
  smallClus <- find2clus(points = points[,c(x,y)])
  
  categories <- c("indID", "population", "language", "region")
  i <- 0
  nGroups <- 20
  while(nGroups > 10){
    i <- i+1
    nGroups <- length(unique(points[smallClus, categories[i]]))
  }
  
  points$color <- "gray"
  j <- 1
  myColors <- rainbow(nGroups)
  for(g in unique(points[smallClus, categories[i]])){
    points$color[points[,categories[i]] == g] <- myColors[j]
    j <- j+1
  }
  points$color[points$population == "Botocudos"] <- "black"
  plot(x = points[,x]/2e5, y = points[,y]/2e5, col = alpha(points$color, 0.5), 
       pch = 16, cex = 2, bty = "n", xlab = x, ylab = y)
  legend(bty = "n", x = 0, y = 0, pch = 16, cex = 1, col = myColors,
         legend = unique(points[smallClus, categories[i]]))
}
```

```{r define colors}
# Give color and shape to 30!!! groups

# Saqqaq, NaDene, EskimoAleut
apricotYellow <- cmyk(0,10,100,0)

# SpiritCave, Clovis, TrailCreek, USR1, Lovelock, Anzick1
oliveYellow <- cmyk(40,30,80,0)

# SouthWesternOntario, NorthernAmerind
benzolGreen <- cmyk(100,15,55,0)

# Belize, CentralAmerind, Mexican, ChibchanPaezan
oldRose <- cmyk(15, 70, 40, 0)

# Andean, EarlyAndes, LateCentralAndes, Southern_Cone, HistoricalAndean
ivoryBuff <- cmyk(8,15,40,0)

# Brazil, LagoaSanta, Taino, EcuatorialTucanoan, Ge-PanoCaribean
spectrumRed <-cmyk(5,100,100,0)

# Aconcagua, PuntaSantaAna, AncientYamana, Ayayema, AncientKaweskar
rainetteGreen<-cmyk(42,20,62,28)

# Botocudos
dullVioletBlack <- cmyk(100,90,38,50)

# Other groups 
lightBrownDrab <- cmyk(8, 30, 20, 25)
nileBlue <- cmyk(25, 0, 10, 0)
rawSienna <- cmyk(18, 58, 100, 12)

paleRawUmber <- cmyk(46, 63, 87, 32)
sulphurYellow <- cmyk(4, 4, 28, 0)
ochraceousSalmon <- cmyk(15, 38, 55, 0)
slateColor <- cmyk(85, 70, 62, 30)
```

# Heterozygosity

```{r}
simons <- read.table("~/Projects/Botocudos/Files/Panels/SGDP/Simons_sample_pop_region_country.txt",
                    header = T, sep = "\t")

```

```{r}
path_to_pi <- "~/Projects/Botocudos/Files/Heterozygosity/2019_05_22/"
files <- list.files(path_to_pi, pattern = ".*.pi.*")

pops <- unique(sub("\\.pi.*", "", files))

pi <- data.frame()

for(p in pops){
  current_pi <- read.table(paste(path_to_pi, p, ".pi.stats.txt", sep = ""))
  colnames(current_pi) <- c('estimate','bias','stderror','confint1','confint2')
  current_pi$Population.ID <- p
  if(dim(pi)[1] == 0){
    pi <- current_pi
  }else{
    pi <- rbind(pi, current_pi)
  }
}

pi <- join(pi, unique(simons[,c("Population.ID", "Region")]),
           by = "Population.ID")
levels(pi$Region) <- c(levels(pi$Region), "Botocudos")
pi$Region[!(complete.cases(pi))] <- "Botocudos"

popsize <- data.frame(table(simons$Population.ID)) 
colnames(popsize) <- c("Population.ID", "popsize")
pi <- join(pi, popsize, by = "Population.ID")
pi$label <- paste(pi$Population.ID, "(n=", pi$popsize, ")", sep="")

```


```{r}
pi_B <- pi

regionOrder <- sapply(levels(pi_B$Region), 
                      function(x) c(x, mean(pi_B$estimate[pi_B$Region == x])) )
regionOrder <- data.frame(Region = regionOrder[1,], mean = regionOrder[2,])

pi_B <- join(pi_B, regionOrder, by = "Region")
pi_B <- pi_B[order(pi_B$mean, pi_B$estimate),]
pi_B$Population.ID <- factor(pi_B$Population.ID, 
                             levels = unique(pi_B$Population.ID), ordered = T)

pi_B <- pi_B[-which(pi_B$Population.ID == "Yoruba"),]
pdf("~/Projects/Botocudos/Plots/Het/condHet.pdf",
    width = 19, height = 12)
ggplot(pi_B, aes(x = Population.ID, y = estimate, ymin = confint1, 
               ymax = confint2, color = Region, interaction = Region)) +
  geom_point() +
  geom_errorbar() +
  #scale_x_discrete( breaks = pi_B$Population.ID, labels= pi_B$label) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, size = 12),
        legend.position = c(0.1,0.7), legend.text = element_text(size = 14),
        title = element_text(size = 20)
        ) + 
  labs(title = "Conditional heterozygosity", 
       x = NULL, y = "Estimate") +
  scale_color_discrete(name = NULL)
dev.off()
```

Heterozygosity per chromosomes

```{r}
boto <- read.table("~/Projects/Botocudos/Files/Heterozygosity/2019_05_22/Botocudos.pi.txt")
surui <- read.table("~/Projects/Botocudos/Files/Heterozygosity/2019_05_22/Surui.pi.txt")
refalt <- read.table("~/Projects/Botocudos/Files/Heterozygosity/2019_05_22/Boto/Nigeria_B_Yoruba-3.refalt")
colnames(refalt) <- c("chr", "start")

colnames(boto) <- c("chr", "start", "end", "pi")
colnames(surui) <- colnames(boto)

chr<-1

layout(matrix(seq(1, 24), byrow = T, nrow = 4, ncol = 6),
       widths = rep(2,6), heights = rep(1,4))

for(chr in 1:22){
  refChr <- refalt[refalt$chr == chr,]
  myChr <- boto[boto$chr == chr,]
  myChr$start <- refChr$start[myChr$start+1]
  myChr$end <- refChr$start[myChr$end+1]
  
  su <- surui[surui$chr == chr,]
  su$start <- refChr$start[su$start+1]
  su$end <- refChr$start[su$end+1]
  plot(bty = "n",# xlim = c(min(myChr$start), max(myChr$end)),
       ylim = c(0,max(c(su$pi, myChr$pi))), xlim = c(0, 2.5e8),
       type = "n", x = 0, y = 0, xlab = NA, ylab = NA,
       main = chr)
  #segments(x0 = su$start, x1 = su$end, y0 = su$pi, lwd = 1.2)
  #segments(x0 = myChr$start, x1 = myChr$end, y0 = myChr$pi, lwd = 1.2, col = "red")
  points(x = su$start, y = su$pi, lwd = 1.2, type = "b")
  points(x = myChr$start, y = myChr$pi, lwd = 1.2, col = "red", type = "b")
  
}
```




# MDS
## "Maanasa" panel

```{r}
ind <- read.table("~/Projects/Botocudos/Files/MDS/2019_05_15/Maanasa_mask1_flip.B.P.ASM.VMAnc.mind0.95.haploid.dist.id")
#Dist <- as.matrix(read.table("~/Projects/Botocudos/Files/MDS/2019_05_15/Maanasa_mask1_flip.B.P.ASM.VMAnc.mind0.95.haploid.dist.gz"))

colnames(ind) <- c("indID", "x")
myPanel <- join(ind, panel, by = "indID")
#res <- cmdscale(dist(Dist), k = dim(Dist)[1]-1, eig = T)
#save(res, file="~/Projects/Botocudos/Files/MDS/2019_05_15/mds_maansa.Rda")
load("~/Projects/Botocudos/Files/MDS/2019_05_15/mds_maansa.Rda")
points <- as.data.frame(res$points)
points <- cbind(points, myPanel)
```

### Make a 'simple' plot


```{r assign colors}
myColors <- c(blue,  yellowOrange, benzolGreen,
              seashellPink, cobaltGreen, rainetteGreen, salviaBlue,
              grayishLavender,
              helvetiaBlue, aconiteViolet)
points$color <- "black"
regions <- c("Africa", "Europe", "Caucasus", "NearEast", 
             "EastAsia", "SouthAsia", "SoutheastAsia",
             "Siberia", "Oceania", "Americas")

for(i in 1:length(regions)){
  points$color[points$region == regions[i]] <- myColors[i]
}

myColors <- c(myColors, "black")
regions <- c(regions, "Botocudos")

wholeMaanasa <- points
```

```{r make the plot}
layout(matrix(seq(1, 2), byrow = T, nrow = 1, ncol = 2),
       widths = c(2,1), heights = c(1))

plot(bty = "n", x = points[,1]/2e5,
     y = points[,2]/2e5, col = alpha(points$color,0.7),
     pch = 16,  xlab = NA, ylab = NA, lwd = 2, cex = 2,
     main = "Multidimensional scaling")
points(x = points[points$region == "Botocudos",1]/2e5,
     y = points[points$region == "Botocudos",2]/2e5,
     col = alpha(points$color[points$region == "Botocudos"],0.7),
     cex = 2, pch = 16)
points(x = points[points$population == "Botocudos2014",1]/2e5,
     y = points[points$population == "Botocudos2014",2]/2e5,
     bg = alpha(points$color[points$population == "Botocudos2014"],0.7),
     cex = 2, pch = 23, col ="black")

plot(1:10, 1:10, type = "n", bty = "n", axes = F, xlab=NA, ylab=NA)
legend(legend = c(regions[1:9], "Botocudos 2014", regions[10:11]), 
       col = c(myColors[1:9], "black", myColors[10:11]),
       x = 1, y = 9, bty = "n", pch = c(rep(16,9), 23, 16),
       pt.bg = c(rep(NA, 9), myColors[9], NA, NA),
       cex = 1.5)

```

#### play with clusters

```{r, eval = F}
ind <- read.table("~/Projects/Botocudos/Files/MDS/2019_05_15/Maanasa_mask1_flip.B.P.ASM.VMAnc.mind0.95.haploid.dist.id")

Dist <- as.matrix(read.table("~/Projects/Botocudos/Files/MDS/2019_05_15/Maanasa_mask1_flip.B.P.ASM.VMAnc.mind0.95.haploid.dist.gz"))
colnames(Dist) <- ind$indID
fit <- hclust(as.dist(Dist), method="ward") 
plot(fit) # display dendogram
groups <- cutree(fit, k=5) # cut tree into 5 clusters
# draw dendogram with red borders around the 5 clusters 
rect.hclust(fit, k=5, border="red")
```

```{r}
wss <- (nrow(Dist)-1)*sum(sapply(Dist[,1],var))
for (i in 2:15) wss[i] <- sum(kmeans(Dist[,1], 
   centers=i)$withinss)
plot(1:15, wss, type="b", xlab="Number of Clusters",
  ylab="Within groups sum of squares")

# K-Means Cluster Analysis
fit <- kmeans(Dist, 5) # 5 cluster solution
# get cluster means 
x <- aggregate(Dist,by=list(fit$cluster),FUN=mean)
# append cluster assignment
mydata <- data.frame(Dist, fit$cluster)
```

```{r}
find2clus <- function(points){
  # find smallest cluster on two dimensions
  wss <- (nrow(points)-1)*sum(apply(points,2,var))
  for (i in 2:15) wss[i] <- sum(kmeans(points, 
                                       centers=i)$withinss)
  # plot(1:15, wss, type="b", xlab="Number of Clusters",
  #      ylab="Within groups sum of squares")

  # K-Means Cluster Analysis
  fit <- kmeans(points, 3) # 5 cluster solution
  # get cluster means 
  aggregate(points,by=list(fit$cluster),FUN=mean)
  smallClus <- names(table(fit$centers))[2:3]
  # append cluster assignment
  return(which(fit$cluster %in% smallClus[1:2]))
}
```

```{r}
# Get some data
ind <- read.table("~/Projects/Botocudos/Files/MDS/2019_05_15/Maanasa_mask1_flip.B.P.ASM.VMAnc.mind0.95.haploid.dist.id", stringsAsFactors = F)

Dist <- as.matrix(read.table("~/Projects/Botocudos/Files/MDS/2019_05_15/Maanasa_mask1_flip.B.P.ASM.VMAnc.mind0.95.haploid.dist.gz"))
colnames(Dist) <- ind$indID

americas <- read.table("~/Projects/Botocudos/Files/Panels/Maanasa/Maanasa.americas.Moreno_names.txt",
                       header = F, stringsAsFactors = F)

colnames(americas) <- c("population", "language")
colnames(ind) <- c("indID", "x")
```

```{r}
png("~/Projects/Botocudos/Plots/MDS/50dim_Maanasa_americas_clus.png",
    width = 20, height = 20, res =250, units = "in")
par(mfrow = c(7,7))
for(i in 1:48){
  plotwclus(i, i+1)
}
dev.off()
```


### Maanasa Americas

Try to make a nice plot within the Americas

```{r load coordinate files}
load("~/Projects/Botocudos/Files/MDS/2019_05_15/mds_maanasa_americas.Rda")
points <- res$points ; dim(points)
ind <- read.table("~/Projects/Botocudos/Files/MDS/2019_05_15/Maanasa_mask1_flip.B.P.ASM.VMAnc.haploid.fred.mind0.95.dist.id")
colnames(ind) <- c("indID", "x")

# Add "language" label
myPanel <- join(ind, panel, by = "indID")


myPanel <- join(myPanel, languages, by = "population")
myPanel$language <- as.character(myPanel$language)
myPanel <- join(myPanel, coordinates, by = "population")

# If not "language" label given, take the population name
index <- is.na(myPanel$language)
myPanel[index,]$language <- as.character(myPanel[index,]$label)
print(unique(myPanel$language[myPanel$region %in% c("Americas", "Botocudos") & myPanel$type == "Ancient"]))

# Select populations in the Americas
index <- which(myPanel$region %in% c("Botocudos", "Americas"))
myPanel <- myPanel[index,] ; dim(myPanel)

```

```{r function to assign color, shape}
giveMeAesthetics <- function(pops, color, shapes, myPanel){
  myPanel$color[myPanel$language %in% pops] <- color
  for(i in 1:length(shapes)){
    myPanel$shape[myPanel$language %in% pops[i]] <- shapes[i]
  }
  return(myPanel)
}

```


```{r define colors per group}
# Give color and shape to 30!!! groups
myPanel$color <- "gray"
myPanel$shape <- 16
myPanel$bg <- "black"
popLevels <- c()

# Saqqaq, NaDene, EskimoAleut
apricotYellow <- cmyk(0,10,100,0)
shapes <- c(21,16,17)
pops <- c("Saqqaq", "NaDene", "EskimoAleut")
myPanel <- giveMeAesthetics(pops, apricotYellow, shapes, myPanel)
popLevels <- c(popLevels, pops)

# SpiritCave, Clovis, TrailCreek, USR1, Lovelock, Anzick1
oliveYellow <- cmyk(40,30,80,0)
pops <- c("SpiritCave", "Clovis", "TrailCreek", "USR1", "Lovelock", "Anzick1")
shapes <- c(21,22,23,22,24,25)
myPanel <- giveMeAesthetics(pops, oliveYellow, shapes, myPanel)
popLevels <- c(popLevels, pops)

# SouthWesternOntario, NorthernAmerind
benzolGreen <- cmyk(100,15,55,0)
pops <- c("SouthWesternOntario", "NorthernAmerind")
shapes <- c(21, 16)
myPanel <- giveMeAesthetics(pops, benzolGreen, shapes, myPanel)
popLevels <- c(popLevels, pops)

# Belize, CentralAmerind, Mexican, ChibchanPaezan
oldRose <- cmyk(15, 70, 40, 0)
shapes <- c(21, 16, 17, 18)
pops <- c("Belize", "CentralAmerind", "Mexican", "ChibchanPaezan")
myPanel <- giveMeAesthetics(pops, oldRose, shapes, myPanel)
popLevels <- c(popLevels, pops)

# Andean, EarlyAndes, LateCentralAndes, Southern_Cone, HistoricalAndean
ivoryBuff <- cmyk(8,15,40,0)
pops <-  c("Andean", "EarlyAndes", "LateCentralAndes", 
           "Southern_Cone", "HistoricalAndean")
shapes <- c(16, 22, 21, 23, 24)
myPanel <- giveMeAesthetics(pops, ivoryBuff, shapes, myPanel)
popLevels <- c(popLevels, pops)

# Brazil, LagoaSanta, Taino, EcuatorialTucanoan, Ge-PanoCaribean
spectrumRed <-cmyk(5,100,100,0)
pops <- c("Brazil", "LagoaSanta", "Taino", "EcuatorialTucanoan", "Ge-PanoCaribean")
shapes <- c(22,21,23,16,17)
myPanel <- giveMeAesthetics(pops, spectrumRed, shapes, myPanel)
popLevels <- c(popLevels, pops)

# Aconcagua, PuntaSantaAna, AncientYamana, Ayayema, AncientKaweskar
rainetteGreen<-cmyk(42,20,62,28)
pops <- c("Aconcagua", "PuntaSantaAna", "AncientYamana", "Ayayema", "AncientKaweskar")
shapes <- c(21:25)
myPanel <- giveMeAesthetics(pops, rainetteGreen, shapes, myPanel)
popLevels <- c(popLevels, pops)

# Botocudos
dullVioletBlack <- cmyk(100,90,38,50)
pops <- c("Botocudos")
shapes <- c(16)
myPanel <- giveMeAesthetics(pops, dullVioletBlack, shapes, myPanel)
myPanel[myPanel$language == "USR1", c("bg", "color")] <- "turquoise"
popLevels <- c(popLevels, pops)

```

```{r join colors, shapes to coordinates}
pointsWColor <- cbind(points, myPanel)

ancient <- which(pointsWColor$type == "Ancient")
modern <- which(pointsWColor$type == "PresentDay")
pointsWColor$bg <- NA
pointsWColor$bg[ancient] <- pointsWColor$color[ancient]
pointsWColor$color[ancient] <- "black"
pointsWColor <- pointsWColor[-which(pointsWColor$population == "Anzick1"),]
pointsWColor <- pointsWColor[-which(pointsWColor$population == "Mexican"),]

pointsWColor$language <- factor(pointsWColor$language,
                                levels = popLevels, ordered = T)
pointsWColor <- pointsWColor[order(pointsWColor$language),]

myLegend <- unique(pointsWColor[,c("language", "color", "shape", "bg")])
myLegend <- myLegend[order(myLegend$language),]

  
americasMaanasa <- pointsWColor
```


```{r plot 50 dimensions}

png("~/Projects/Botocudos/Plots/MDS/50dim_Maanasa_americas.png",
    width = 45, height = 25, res = 200, units = "in")
layout(matrix(seq(1, 45), byrow = T, nrow = 5, ncol = 9))

for(i in seq(1,44)){
  name <- paste("~/Projects/Botocudos/Plots/MDS/mini/", i, ".png", sep = "")
  png(name, width = 4, height = 4, res = 200, units = "in", bg = NA)
  par(mar = rep(4,4))
  if(i %%2){
      x <- i ; y <- i+1
  }else{
    x <- i + 1; y <- i
  }

  plot(bty = "n", xlab = x, ylab = y,
       x = pointsWColor[, x]/2e5, y = pointsWColor[, y]/2e5,
       pch = pointsWColor$shape, 
       col = alpha(pointsWColor$color, 0.9), 
       bg = pointsWColor$bg, cex = 2, lwd = 1.2)
  
  points(x = pointsWColor[pointsWColor$population == "Botocudos", x]/2e5, 
         y = pointsWColor[pointsWColor$population == "Botocudos", y]/2e5,
         pch = pointsWColor$shape[pointsWColor$population == "Botocudos"], 
         bg = alpha(pointsWColor$color[pointsWColor$population == "Botocudos"], 0.9), 
         cex = 2, lwd = 2)
  # points(x = pointsWColor[pointsWColor$population == "Karitiana", x]/2e5, 
  #        y = pointsWColor[pointsWColor$population == "Karitiana", y]/2e5,
  #        pch = 4, col = "black",
  #        cex = 1)
    # points(x = pointsWColor[pointsWColor$population == "Surui", x]/2e5, 
    #      y = pointsWColor[pointsWColor$population == "Surui", y]/2e5,
    #      pch = 1, col = "black",
    #      cex = 1)
    dev.off()
}

plot(1:10, 1:10, axes = F, xlab=NA, ylab=NA, type = "n", bty = "n")
legend(x = 2, y = 9, bty = "n", ncol = 2, legend = myLegend$language,
       pt.bg = myLegend$bg,
       pch = myLegend$shape,
       col = myLegend$color, pt.lwd = 1.2, cex = 1.2)

dev.off()
```


Plot desired coordinates/populations
```{r put.fig.letter}
put.fig.letter <- function(label, location="topleft", x=NULL, y=NULL, 
                           offset=c(0, 0), ...) {
  if(length(label) > 1) {
    warning("length(label) > 1, using label[1]")
  }
  if(is.null(x) | is.null(y)) {
    coords <- switch(location,
                     topleft = c(0.015,0.98),
                     topcenter = c(0.5525,0.98),
                     topright = c(0.985, 0.98),
                     bottomleft = c(0.015, 0.02), 
                     bottomcenter = c(0.5525, 0.02), 
                     bottomright = c(0.985, 0.02),
                     c(0.015, 0.98) )
  } else {
    coords <- c(x,y)
  }
  this.x <- grconvertX(coords[1] + offset[1], from="nfc", to="user")
  this.y <- grconvertY(coords[2] + offset[2], from="nfc", to="user")
  text(labels=label[1], x=this.x, y=this.y, xpd=T, ...)
}
```



```{r}
nCol <- 3
nRow <- 3
legendCol <- 2

# pdf("~/Projects/Botocudos/Plots/MDS/MDS_midThesis.pdf",
#    width = 9, height = 9)
png("~/Projects/Botocudos/Plots/MDS/MDS_midThesis.png",
   width = 9, height = 9, res = 300, units = "in", bg = "NA")
layout(matrix(seq(1, nRow*nCol), byrow = T, nrow = nRow, ncol = nCol),
       widths = c(rep(2, nCol-1 ), 2.5), heights = rep(2, nRow))

#------------------------------------------------------------------------------#
# All the panel

for(i in 1:2){
par(xpd = NA)
plot(bty = "n", x = wholeMaanasa[,i]/2e5,
     y = wholeMaanasa[,i+1]/2e5, col = alpha(wholeMaanasa$color,0.7),
     pch = 16,  xlab = i, ylab = i+1, lwd = 2, cex = 2,
     main = "MDS, all populations")
points(x = wholeMaanasa[wholeMaanasa$region == "Botocudos",i]/2e5,
     y = wholeMaanasa[wholeMaanasa$region == "Botocudos",i+1]/2e5,
     col = alpha(wholeMaanasa$color[wholeMaanasa$region == "Botocudos"],0.7),
     cex = 2, pch = 16)
points(x = wholeMaanasa[wholeMaanasa$population == "Botocudos2014",1]/2e5,
     y = wholeMaanasa[wholeMaanasa$population == "Botocudos2014",2]/2e5,
     bg = alpha(wholeMaanasa$color[wholeMaanasa$population == "Botocudos2014"],0.7),
     cex = 2, pch = 23, col ="black")
put.fig.letter(LETTERS[i], location = "topleft", font = 2)
}
par(xpd = NA)
plot(1:10, 1:10, type = "n", bty = "n", axes = F, xlab=NA, ylab=NA, xpd = T)
legend(legend = c(regions[1:9], "Botocudos 2014", regions[10:11]), 
       col = c(myColors[1:9], "black", myColors[10:11]),
       x = 0, y = 10, bty = "n", pch = c(rep(16,9), 23, 16),
       pt.bg = c(rep(NA, 9), myColors[9], NA, NA),
       cex = 1.2, ncol = 2)

#------------------------------------------------------------------------------#
# Only the Americas

j <- 2
i <- 3
for(x in c(1,5,6,7)){
  y <- x + 1
  plot(bty = "n", xlab = x, ylab = y,
       x = pointsWColor[, x]/2e5, y = pointsWColor[, y]/2e5,
       pch = pointsWColor$shape, 
       col = alpha(pointsWColor$color, 0.9), 
       bg = pointsWColor$bg, cex = 2, lwd = 1.2,
       main = "MDS, Native Americans")
  
  points(x = pointsWColor[pointsWColor$population == "Botocudos", x]/2e5, 
         y = pointsWColor[pointsWColor$population == "Botocudos", y]/2e5,
         pch = pointsWColor$shape[pointsWColor$population == "Botocudos"], 
         bg = alpha(pointsWColor$color[pointsWColor$population == "Botocudos"],
                    0.9), 
         cex = 2, lwd = 2)
  put.fig.letter(LETTERS[i], "topleft", font = 2) ; i <- i+1
  if(j %% nCol == 0){
    plot(1:10, 1:10, axes = F, xlab=NA, ylab=NA, type = "n", bty = "n")
    par(xpd=TRUE)
    legend(x = 0, y = 12, bty = "n", ncol = legendCol, legend = myLegend$language,
           pt.bg = myLegend$bg,
           pch = myLegend$shape,
           col = myLegend$color, pt.lwd = 1, cex = 1.2)
    #plot(1:10, 1:10, type = "n", bty = "n", axes = F, xlab=NA, ylab=NA)
  }
    j <- j + 1
}


    
dev.off()
```

```{r}

myR <- unique(wholeMaanasa$region)
i <-1
for(r in myR){
  name <- paste("~/Projects/Botocudos/Plots/MDS/mini/", r, ".png", sep = "")
  png(name, width = 4, height = 4, res = 300, units = "in", bg = NA)
  par(mar = rep(2,4))
  plot(bty = "n", x = wholeMaanasa[,i]/2e5,
       y = wholeMaanasa[,i+1]/2e5, col = alpha(wholeMaanasa$color,0.7),
       pch = 16,  xlab = i, ylab = i+1, lwd = 2, cex = 2,
       main = "MDS, all populations", type = "n")
  
  points(x = wholeMaanasa[wholeMaanasa$population == r,i]/2e5,
         y = wholeMaanasa[wholeMaanasa$population == r,i+1]/2e5,
         col = alpha(wholeMaanasa$color[wholeMaanasa$population == r],0.7),
         cex = 2, pch = 16)
  dev.off()
}


  points(x = wholeMaanasa[wholeMaanasa$region == "Botocudos",i]/2e5,
     y = wholeMaanasa[wholeMaanasa$region == "Botocudos",i+1]/2e5,
     col = alpha(wholeMaanasa$color[wholeMaanasa$region == "Botocudos"],0.7),
     cex = 2, pch = 16)
points(x = wholeMaanasa[wholeMaanasa$population == "Botocudos2014",1]/2e5,
     y = wholeMaanasa[wholeMaanasa$population == "Botocudos2014",2]/2e5,
     bg = alpha(wholeMaanasa$color[wholeMaanasa$population == "Botocudos2014"],0.7),
     cex = 2, pch = 23, col ="black")
```


# Plot a Map

```{r}
#pdf("~/Projects/Botocudos/Plots/Map/MDS_map.pdf", width = 6, height = 8)
png("~/Projects/Botocudos/Plots/Map/MDS_map.png", width = 6, height = 8,
    res = 300, units = "in")

par(fg = NULL)
map('world', fill = T, col = "gray90",
    xlim=c(-168,-10), ylim = c(-60, 100),
    border="white", lforce = "e")

points(x = pointsWColor$long, y = pointsWColor$lat, pch = pointsWColor$shape,
       col = pointsWColor$color, bg = pointsWColor$bg, cex = 1.2)

dev.off()
```




## Make plot individual by individual

```{r}
ind <- read.table("~/Projects/Botocudos/Files/MDS/2019_05_15/Maanasa_mask1_flip.B.P.ASM.VMAnc.haploid.dist.id")
Dist <- as.matrix(read.table("~/Projects/Botocudos/Files/MDS/2019_05_15/Maanasa_mask1_flip.B.P.ASM.VMAnc.haploid.dist.gz"))

colnames(ind) <- c("indID", "x")
myPanel <- join(ind, panel, by = "indID")
boto <- ind[grep("MN", ind$indID),]

for(i in 1:nrow(boto)){
  myInd <- ind
  toKeep <- boto[i,1]
  print(toKeep)
  idToRemove <- boto$indID[which(!(boto$indID %in% toKeep))]
  myIndex <- which(!(ind$indID %in% idToRemove) & 
                     !(is.na(Dist[, which(ind$indID == toKeep)])))
  myDist <- Dist[myIndex, myIndex]
  myInd <- myInd[myIndex,]
  myIndex <- sapply(1:ncol(myDist), function(x) sum(is.na(Dist[,x])))
  while(sum(myIndex)){
    idToRemove <- which(myIndex == max(myIndex))[1]
    print(paste("removing", myInd[idToRemove,1], myIndex[idToRemove], max(myIndex)))
    myInd <- myInd[-idToRemove,]
    myDist <- myDist[-idToRemove, -idToRemove]
    myIndex <- sapply(1:ncol(myDist), function(x) sum(is.na(myDist[,x])))

  }
  #res <- cmdscale(dist(myDist), k = dim(myDist)[1]-1, eig = T)
  f <- paste("~/Projects/Botocudos/Files/MDS/2019_05_15/mds_maansa_",
             toKeep,".ind", sep = "")
  print(paste("Saving to ", f))
  #save(myInd, file=f)
  write.table(myInd, file = f, col.names = F, row.names = F, quote = F, sep  ="\t")
}

#)

#load("~/Projects/Botocudos/Files/MDS/2019_05_15/mds_maansa_ALL.Rda")
#points <- as.data.frame(res$points)
#points <- cbind(points, myPanel)
```

```{r}
pdf("~/Projects/Botocudos/Plots/MDS/24Boto.pdf",
   width = 16, height = 19)

nRow <- 9
nCol <- 6
layout(matrix(seq(1, nRow*nCol), byrow = T, nrow = nRow, ncol = nCol),
       widths = c(rep(2, nCol-1 ), 2.5), heights = rep(2, nRow))

for(j in 1:nrow(boto)){
  if(!(j-1)%%3){
    name <- paste("~/Projects/Botocudos/Plots/MDS/2019_05_15/", j, "_", j+2,
                  ".pdf", sep = "")
    pdf(name, width = 8, height = 12)
    layout(matrix(seq(1, 6), byrow = T, nrow = 3, ncol = 2),
       widths = c(rep(2, 2 )), heights = rep(2, 3))
    
  }
  name <- paste("~/Projects/Botocudos/Files/MDS/2019_05_15/mds_maansa_", boto[j, 1],
  ".Rda", sep = "")
  load(name)
  points <- res$points ; dim(points)
  name <- sub("Rda", "ind", name)
  ind <- read.table(name)
  colnames(ind) <- c("indID", "x")
  
  # Add "language" label
  myPanel <- join(ind, panel, by = "indID")
  points <- cbind(points, myPanel)
  
  myColors <- c(blue,  yellowOrange, benzolGreen,
                seashellPink, cobaltGreen, rainetteGreen, salviaBlue,
                grayishLavender,
                helvetiaBlue, aconiteViolet)
  
  points$color <- "black"
  
  regions <- c("Africa", "Europe", "Caucasus", "NearEast", 
               "EastAsia", "SouthAsia", "SoutheastAsia",
               "Siberia", "Oceania", "Americas")
  
  for(i in 1:length(regions)){
    points$color[points$region == regions[i]] <- myColors[i]
  }
  
  myColors <- c(myColors, "black")
  regions <- c(regions, "Botocudos")
  
  wholeMaanasa <- points
  for(i in 1:2){
    par(xpd = NA)
    plot(bty = "n", x = wholeMaanasa[,i]/2e5,
         y = wholeMaanasa[,i+1]/2e5, col = alpha(wholeMaanasa$color,0.7),
         pch = 16,  xlab = i, ylab = i+1, lwd = 2, cex = 2,
         main =boto[j,1])
    points(x = wholeMaanasa[wholeMaanasa$region == "Botocudos",i]/2e5,
           y = wholeMaanasa[wholeMaanasa$region == "Botocudos",i+1]/2e5,
           col = alpha(wholeMaanasa$color[wholeMaanasa$region == "Botocudos"],0.7),
           cex = 2, pch = 16)
    points(x = wholeMaanasa[wholeMaanasa$population == "Botocudos2014",1]/2e5,
           y = wholeMaanasa[wholeMaanasa$population == "Botocudos2014",2]/2e5,
           bg = alpha(wholeMaanasa$color[wholeMaanasa$population == "Botocudos2014"],0.7),
           cex = 2, pch = 23, col ="black")
    #put.fig.letter(LETTERS[i], location = "topleft", font = 2)
  }
  if(!j%%3){
    dev.off()
  }
}

par(xpd = NA)
plot(1:10, 1:10, type = "n", bty = "n", axes = F, xlab=NA, ylab=NA, xpd = T)
legend(legend = c(regions[1:9], "Botocudos 2014", regions[10:11]), 
       col = c(myColors[1:9], "black", myColors[10:11]),
       x = 0, y = 10, bty = "n", pch = c(rep(16,9), 23, 16),
       pt.bg = c(rep(NA, 9), myColors[9], NA, NA),
       cex = 1.2, ncol = 2)

dev.off()

```




# Admixture

```{r}
"2_26,21,81,61,41,1,81,61,41,41,1,81,72,81"
"for(k in paste(seq(2, 15 ), c( 26,21,81,61,41,1,81,61,41,41,1,81,72,81 ), sep = '_'))" 

```

```{r}
preffix <- "~/Projects/Botocudos/Files/ADMIX/2019_05_19/subset_maanasa_k"
suffix <- ".qopt"

ind <- read.table("~/Projects/Botocudos/Files/ADMIX/2019_05_19/B.P.ASM.VMAnc")
ind$V1 <- sub(".*/", "", ind$V1)
ind$V1 <- sub(".bam", "", ind$V1)

p <- read.table("~/Projects/Botocudos/Files/ADMIX/2019_05_19/Maanasa_mask1_flip.fam", stringsAsFactors = F)


x <- read.csv("~/Projects/Botocudos/Files/ADMIX/2019_05_19/columns2cut_maanasa.txt", header = F, stringsAsFactors = F)
x[1,] <- sub("-.*", "", x[1,])

x <- x[1,-1]
x <- as.integer(x[1,])
x <- (x -1)/3


myPanel <-  data.frame(indID = c(ind$V1, p$V1)[x])
myPanel <- join(myPanel, panel, by = "indID")

languages <- read.table("~/Projects/Botocudos/Files/Panels/Maanasa/Maanasa.americas.Moreno_names.txt", header = F)
colnames(languages) <- c("population", "language")

myPanel <- join(myPanel, languages, by = "population")
myPanel$language <- as.character(myPanel$language)
myPanel <- join(myPanel, coordinates, by = "population")

# If not "language" label given, take the population name
index <- is.na(myPanel$language)
myPanel[index,]$language <- as.character(myPanel[index,]$label)
print(unique(myPanel$language[myPanel$region %in% c("Americas", "Botocudos") & 
                                myPanel$type == "Ancient"]))

```


```{r}

```


```{r}
# Aesthetics
# Widths
myPanel$width <- .2
myPanel$width[myPanel$region != "Americas"] <- 1.5
myPanel$width[myPanel$language %in% c("Yorubas", "Papuans","French",
                                                        "Botocudos2014",
                                                        "Dai", "Sindhi", "Koryaks", "USR1",
                                                        "Saqqaq", 
                                                        "TrailCreek", 
                                                        "Anzick1", "Clovis", "SpiritCave", "Lovelock",
                                                        "SouthWesternOntario", "Taino",
                                                        "Belize", 
                                                        "EarlyAndes", "LateCentralAndes", "HistoricalAndean",
                                                        "Ayayema", "AncientKaweskar", "AncientYamana",
                                                        "Aconcagua", "PuntaSantaAna",
                                                        "Southern_Cone",  "Brazil",
                                                        "LagoaSanta", "Botocudos")] <- 1.5
myPanel$width[myPanel$population == "Botocudos"] <- 4
myPanel$width[myPanel$population == "Botocudos2014"] <- 4

# Order things
myPanel$region <- factor(myPanel$region, levels = c("Africa", "Europe", "Oceania",
                                                    "EastAsia", "SouthAsia",
                                                    "SoutheastAsia", "Siberia",
                                                    "Americas", 
                                                    "Botocudos"), 
                         ordered  = T)

myPanel$language <- factor(myPanel$language, levels = c("Yorubas", "Papuans","French",
                                                        "Botocudos2014",
                                                        "Dai", "Sindhi", 
                                                        "Koryaks", "EskimoAleut", 
                                                        "Saqqaq", "USR1",
                                                        "TrailCreek", 
                                                        "Anzick1", "Clovis", "SpiritCave", "Lovelock",
                                                        "SouthWesternOntario","NaDene",
                                                        "NorthernAmerind", "Mexican",
                                                        "CentralAmerind",
                                                        "ChibchanPaezan", "Taino",
                                                        "Belize", 
                                                        "EarlyAndes", "LateCentralAndes", "HistoricalAndean",
                                                        "Andean",
                                                        "Ayayema", "AncientKaweskar", "AncientYamana",
                                                        "Aconcagua", "PuntaSantaAna",
                                                        "Southern_Cone", "EcuatorialTucanoan",
                                                        "Ge-PanoCaribean", "Brazil",
                                                        "LagoaSanta", "Botocudos"),
                           ordered = T)

myColor <- data.frame( k = seq(1,15), colors = c(#rgb(0.1, 0.1, 0.1), rgb(0.3, 0.3, 0.3),
                                                 #rgb(0.5, 0.5, 0.5), rgb(0.7, 0.7, 0.7),#"gray10", "gray30", "gray50", "gray70",
  apricotYellow,  paleRawUmber, nileBlue,ivoryBuff,oldRose,                                               
  rawSienna, ochraceousSalmon, benzolGreen,
  slateColor, spectrumRed, rainetteGreen, sulphurYellow,
  oliveYellow, lightBrownDrab, dullVioletBlack))

myColor$colors <- as.character(myColor$colors)

```

```{r}
write.table(myPanel, "~/Projects/Botocudos/Files/ADMIX/2019_05_19/panel.txt",
            col.names = T, row.names = F, sep = "\t", quote = F)

write.table((myColor$colors), "~/Projects/Botocudos/Files/ADMIX/2019_05_19/colors.txt",
            col.names = F, row.names = F, sep = "\t", quote = F
            )

write.table(levels(myPanel$language), "~/Projects/Botocudos/Files/ADMIX/2019_05_19/order_language.txt",
            col.names = F, row.names = F, sep = "\t", quote = F)

write.table(unique(myPanel[order(myPanel$language), c("population", "population")]), "~/Projects/Botocudos/Files/ADMIX/2019_05_19/pop_pop.txt",
            col.names = F, row.names = F, sep = "\t", quote = F)
```


```{r}
# Functions to prepare admix plot

orderMeanComp <- function(k, nCol, admix){
  # Take the ordered levels from the column of interest
  myLevels <- levels(admix[, nCol])
  col <- which(colnames(admix) == colnames(admix)[nCol])
  myOrder <- c()
  K <- seq(1, k)
  while(length(K)){
    for(l in myLevels){
      ind <- which(admix[,col] == l)
      meanComp <- colMeans(as.matrix(admix[ind,K]))
      meanComp <- meanComp[!(names(meanComp) %in% myOrder) & meanComp > 0.0002]
      #print(l)
      if(length(K) == 1){
        myOrder <- c(myOrder, K)
        K <- c()
        break
      }else if(length(meanComp)){
        highestComponent <- as.integer(names(meanComp)[which(meanComp == max(meanComp))])
        myOrder <- c(myOrder, highestComponent)
        #print(highestComponent)
        K <- K[-which(K ==highestComponent)]
      }
    }
  }
  return(admix[,c(myOrder, (length(myOrder)+1):ncol(admix))])
  
}

admixPrepare <- function(k){
  name <- paste(preffix, k, ".qopt", sep = "")
  admix <- read.table(name)
  nInd <- dim(admix)[1]
  ncomp <-  dim(admix)[2]
  colnames(admix) <- seq(1,ncomp)
  # Add the metadate defined in myPanel
  admix <- cbind(admix, myPanel)
  # Order according to the previously established levels
  admix <- admix[order( admix$region, 
                        admix$language, admix$population,
                        decreasing = F),]
  # admix[, 1:ncomp] <- admix[, order_comp(admix, c("Botocudo", "Africa", "Oceania", "EastAsia", "Europe"))]
  
}

get_popline <- function(nCol, admix){
  pop_line <- sapply(unique(admix[,nCol]), 
                     function(X) which(
                       admix$indID == 
                         tail(admix$indID[admix[,nCol] == X], 1)))
  names(pop_line) <- unique(admix[,nCol])
  pop_line <- pop_line[order(pop_line)]
  w <- sapply(names(pop_line), function(x) sum(admix$width[admix[,nCol] == x]))
  pop_line <- cumsum(pop_line*w)
}
  
```

```{r}
#png("~/Desktop/admixture.png", width = 20, height = 20, res = 300, units = "in")
# feature from which we will order
featureOrder <- "language"
par(mfrow = c(15,1))
par(mar = c(2, .5, 1, 2))

for(k in paste(seq(2, 15 ), c( 26,21,81,61,41,1,81,61,41,41,1,81,72,81 ), sep = '_')){
    name <- paste(preffix, k, ".qopt", sep = "")
  admix <- read.table(name)
  nInd <- dim(admix)[1]
  ncomp <-  dim(admix)[2]
  colnames(admix) <- seq(1,ncomp)
   admix <- cbind(admix, myPanel)
  nCol <- which(colnames(admix) == featureOrder)
  admix <- admix[order(
                       admix$language, admix$population,
                       decreasing = F),]
  
  pop_line <- get_popline(nCol = nCol, admix = admix)
  admix <- orderMeanComp(k = as.integer(sub("_.*", "",k)), 
                         nCol = nCol, 
                         admix = admix)
  admix <- admix[which(admix$population != "Bajo"),]
  barplot(as.matrix(t(admix[1:nInd,1:ncomp])), space=0,border=NA, 
          xlab=NULL,
          ylab = paste("k =", sub("_.*", "",k)), horiz = F, cex.main = 3,
          cex.lab = 2, axes = F, axisnames= F, width = admix$width, col = as.character(myColor$colors))
  abline(v = cumsum(w),
         col = "white")
}

rownames(admix) <- 1:dim(admix)[1]
pop_names <- sapply(unique(admix$language), 
                    function(X) mean(as.numeric(rownames(admix[admix$language == X,]))))
names(pop_names) <- unique(admix$language)
text(cumsum(w)-0.5*w, -0., names(pop_names), xpd=NA, cex = 1, srt = 90)

#dev.off()
```



# F3

## Maanasa panel

```{r}
f3 <- read.table("~/Projects/Botocudos/Files/F3/2019_05_17/boto_maanasa.txt",
                 header = T)
colnames(coordinates) <- c("Source2", "lat", "long", "type")
f3$Source2 <- sub("Conchal.*", "ConchaliSantiagoRM", f3$Source2)
f3 <- join(f3, coordinates, by = "Source2")
f3 <- f3[f3$SNPs >10000,]

f3$population <- f3$Source2
myPanel <- unique(panel[,c("population", "region")])
f3 <- join(f3, myPanel, by = "population")
 f3 <- join(f3, languages, by = "population")
```

```{r}
myPalette <- choose_palette()
```

### Color by F3 value

```{r}
f3 <- f3[f3$region == "Americas",]
f3 <- f3[order(f3$f_3, decreasing = T),]
myColor <- myPalette(nrow(f3))[rank(f3$f_3, "max")]
myColor <- myPalette((max(f3$f_3)-min(f3$f_3))*1e6)[(f3$f_3-min(f3$f_3))*1e6+1]

myBg <- myColor
myColor[f3$type=="Ancient"] <- "black"
f3$shape <- 16
f3$shape[f3$type == "Ancient"] <- 21

topHits <- f3[1:20,c("population", "type", "lat", "long")]
rownames(topHits) <- 1:20
 pdf("~/Projects/Botocudos/Plots/F3/Maanasa_F3_americas.pdf",
      width = 8, height = 8)
layout(matrix(seq(1, 3), byrow = F, nrow = 1, ncol = 3),
       widths = c(1,2,6), heights = c(6))


barplot(cbind(seq(min(f3$f_3), max(f3$f_3), length.out = 100)),
        border = NA,
        col = myPalette(100), yaxt="n", horiz = F)
axis(2, at = seq(0, 25, length.out = 10), 
     las = 2,
     labels = round(seq(min(f3$f_3), max(f3$f_3), length.out = 10), 3))

nBreaks <- 50
x <- hist(x=f3$f_3, breaks = nBreaks, border = NA, 
          col = myPalette(20), plot = F)

barplot(x$counts, horiz = T, col = myPalette(length(x$breaks)), beside = T, axes = F,
        border = NA)
par(fg = NULL)
map('world', fill = T, col = "gray90",
    xlim=c(-168,-10), ylim = c(-60, 100),
    border="white", lforce = "e")


topHits$offx <- 0
topHits$offx[topHits$long > -55] <- 10
topHits$offx[topHits$long < -70] <- -10

topHits$offy <- 0
topHits$offy[topHits$lat > 0] <- 10
topHits$offy[topHits$lat < -25] <- -10

text(x = topHits$offx + topHits$long, y = topHits$lat + topHits$offy, 
     labels = rownames(topHits),
     adj = 1)
segments(x0 = topHits$long + topHits$offx, 
         x1 = topHits$long,
          y0 = topHits$lat + topHits$offy,
         y1 = topHits$lat)

points(x = f3$long, y = f3$lat, 
       col = alpha(myColor, 0.8), 
       bg = alpha(myBg, 0.8),
       pch=f3$shape, cex = 2, lwd = 1
       )
text(x = -150, y = -6*as.integer(rownames(topHits))+50, 
     labels = paste(rownames(topHits),
                    topHits$population),
     adj = 0)

offset <- 10


dev.off()

```

```{r}
f3$Source2 <- factor(f3$Source2, levels = f3$Source2, ordered = T)

ggplot(f3, aes(x = Source2, y = f_3, ymin = f_3 - std.err, ymax = f_3 + std.err)) +
  geom_point() +
  geom_errorbar() +
  coord_flip()
```


### Jabuticabeira
```{r}
f3 <- read.table("~/Projects/Botocudos/Files/F3/2019_05_17/jabuticabeira_maanasa.txt",
                 header = T)
colnames(coordinates) <- c("Source2", "lat", "long", "type")
f3$Source2 <- sub("Conchal.*", "ConchaliSantiagoRM", f3$Source2)
f3 <- join(f3, coordinates, by = "Source2")
f3 <- f3[f3$SNPs >10000,]

f3$population <- f3$Source2
myPanel <- unique(panel[,c("population", "region")])
f3 <- join(f3, myPanel, by = "population")
```

```{r}
myPalette <- choose_palette()
```



```{r}
f3 <- f3[f3$region %in% c("Americas", "Botocudos") & f3$Source1 == "Jabuticabeira2",]
f3 <- f3[order(f3$f_3, decreasing = T),]
myColor <- myPalette(nrow(f3))[rank(f3$f_3, "max")]
myColor <- myPalette((max(f3$f_3)-min(f3$f_3))*1e6)[(f3$f_3-min(f3$f_3))*1e6+1]

myBg <- myColor
myColor[f3$type=="Ancient"] <- "black"
f3$shape <- 16
f3$shape[f3$type == "Ancient"] <- 21

topHits <- f3[1:20,c("population", "type", "lat", "long")]
rownames(topHits) <- 1:20
 # pdf("~/Projects/Botocudos/Plots/F3/Maanasa_F3_jabuticabeira.pdf",
 #      width = 8, height = 8)
layout(matrix(seq(1, 3), byrow = F, nrow = 1, ncol = 3),
       widths = c(1,2,6), heights = c(6))


barplot(cbind(seq(min(f3$f_3), max(f3$f_3), length.out = 100)),
        border = NA,
        col = myPalette(100), yaxt="n", horiz = F)
axis(2, at = seq(0, 25, length.out = 10), 
     las = 2,
     labels = round(seq(min(f3$f_3), max(f3$f_3), length.out = 10), 3))

nBreaks <- 50
x <- hist(x=f3$f_3, breaks = nBreaks, border = NA, 
          col = myPalette(20), plot = F)

barplot(x$counts, horiz = T, col = myPalette(length(x$breaks)), beside = T, axes = F,
        border = NA)
par(fg = NULL)
map('world', fill = T, col = "gray90",
    xlim=c(-168,-10), ylim = c(-60, 100),
    border="white", lforce = "e")


topHits$offx <- 0
topHits$offx[topHits$long > -55] <- 10
topHits$offx[topHits$long < -70] <- -10

topHits$offy <- 0
topHits$offy[topHits$lat > 0] <- 10
topHits$offy[topHits$lat < -25] <- -10

text(x = topHits$offx + topHits$long, y = topHits$lat + topHits$offy, 
     labels = rownames(topHits),
     adj = 1)
segments(x0 = topHits$long + topHits$offx, 
         x1 = topHits$long,
          y0 = topHits$lat + topHits$offy,
         y1 = topHits$lat)

points(x = f3$long, y = f3$lat, 
       col = alpha(myColor, 0.8), 
       bg = alpha(myBg, 0.8),
       pch=f3$shape, cex = 1.5, lwd = 1
       )
text(x = -150, y = -6*as.integer(rownames(topHits))+50, 
     labels = paste(rownames(topHits),
                    topHits$population),
     adj = 0)

offset <- 10


# dev.off()

```

```{r}
f3$Source2 <- factor(f3$Source2, levels = f3$Source2, ordered = T)

ggplot(f3, aes(x = Source2, y = f_3, ymin = f_3 - std.err, ymax = f_3 + std.err)) +
  geom_point() +
  geom_errorbar() +
  coord_flip()
```


# D-stat

## Stratify things by type of data

```{r read in files}
d <- read.table("~/Projects/Botocudos/Files/Dstat/2019_05_19/h_corrected.txt",
                stringsAsFactors = F)
colnames(d) <- c("result","Pop1",  "Pop2", "Pop3",  
                 "Pop4",  "Dstat","Z", "BABA",
                 "ABBA",	"SNPs")
d <- d[d$SNPs >10000,]
panel <- read.table("~/Projects/Botocudos/Files/Panels/Magic.panel", header = T)

d$Pop1 <- sub("Conchal.*", "ConchaliSantiagoRM", d$Pop1)
d$Pop2 <- sub("Conchal.*", "ConchaliSantiagoRM", d$Pop2)

# 161 populations
# unique(c(d$Pop1, d$Pop2, d$Pop3))

#------------------------------------------------------------------------------#
# Set of samples that were captured (Posth)
posth <- read.table("~/Projects/Botocudos/Files/Panels/Posth/Posth.panel.txt",
                    header = T)
#------------------------------------------------------------------------------#
# Set of samples for which we sampled a random read
sampledReads <- read.table("~/Projects/Botocudos/Files/Dstat/2019_05_19/B.P.ASM.VMAnc")
colnames(sampledReads) <- "indID"
sampledReads$indID <- sub(".*/", "", sampledReads$indID)
sampledReads$indID <- sub(".bam", "", sampledReads$indID)

sampledReads <- join(sampledReads, panel, by = "indID")
sampledReads <- sampledReads[-which(sampledReads$indID %in% posth$indID),]
sampledReads <- sampledReads[!(sampledReads$indID %in% c("Taino", "USR1", "Ayayema")),]
#------------------------------------------------------------------------------#
# Set of samples for which we called genotypes 
calledGenos <- read.table("~/Projects/Botocudos/Files/Panels/h/h.panel", 
                          header = T)
calledGenos <- calledGenos[-which(
  calledGenos$population %in% c("Taino", #"USR1",
                                "Ayayema", #"Lovelock2", 
                                #"Lovelock3",
                                "Saqqaq", "Inuit",
                                #"Lovelock3", "Anzick1",
                                "Tlingit"#, #"SpCave", #"LagoaSta",
                                #"Chane", "Piapoco"
                                )),]
```

```{r separate by sampled or called}

keepOrRemove <- function(group, action, dataset){
  if(action == "keep"){
    return(which(dataset %in% group))
  }else{
    return(which(!(dataset %in% group)))
  }
}
stratifyMe <- function(set, action, dataset){
  if(set == "captured"){
    index <- keepOrRemove(posth$population, action, dataset)
  }else if(set == "sampled"){
    index <- keepOrRemove(sampledReads$population, action, dataset)
  }else if(set == "called"){
    index <- keepOrRemove(calledGenos$population, action, dataset)
  }else if(set == "all"){
    index <- 1:length(dataset)
  }
  return(index)
}

```

```{r}
# Function to switch the desired population to Pop1
switchPops <- function(myD, pop1){
  index <- myD$Pop2 == pop1
  pop2 <- myD$Pop1[index]
  myD$Pop2[index] <- pop2
  myD$Pop1[index] <- pop1
  myD$Z[index] <- -myD$Z[index]
  myD$Dstat[index] <- -myD$Dstat[index]
  
  return(myD)
}
```

```{r organizeDstat}
# Function to organize the Dstat dataset, according to what we want to
# display for Pop1, Pop2, Pop3

organizeDstat <- function(set1 = "all", set2 = "all", set3 = "all", d){
  
  if(set1 == "called"){
    popsToSwitch <- unique(calledGenos$population)
  }else if(set1 == "captured"){
    popsToSwitch <- unique(posth$population)
  }else if(set1 == "sampled"){
    popsToSwitch <- unique(sampledReads$population)
  }else if(set1 == "all"){
    popsToSwitch <- unique(c(d$Pop1, d$Pop2))
  }
  
  # Put all the desired populations in Pop1
  # plus subset the tests with set3 in Pop3
  # aaand subset to tests with set2 in Pop2
  myD <- data.frame()
  for(pop in popsToSwitch){ 
    tmp <- switchPops(d, pop)
    tmp <- tmp[tmp$Pop1 == pop,]
    index <- stratifyMe(set = set2, action = "keep", dataset = tmp$Pop2)
    tmp <- tmp[index,]
    index <- stratifyMe(set = set3, action = "keep", dataset = tmp$Pop3)
    tmp <- tmp[index,]
    if(nrow(tmp[complete.cases(tmp),])){
      myD <- rbind(myD, tmp)
    }else{
      myD <- tmp
    }
  }
  return(myD)
  
}
```

```{r}
set1 = "called"
set2 = "captured"
set3 = "all" 

x <- organizeDstat(set1, set2, set3, d)

```

```{r orderNean}
orderNean <- function(pop, myD, POP){
  myD <- switchPops(d, pop)
  m <- mean(myD$Z[myD$Pop3 == "AltaiNean"])
  m <- median(unlist(sapply(POP, 
                            function(x) rank(myD$Z[myD$Pop2 == x])[which(myD$Pop3[myD$Pop2 == x] == "AltaiNean")])))
  return(m)
}
```

## h panel
## Archaic trend
## Australasian trend
## Within the Americas

```{r}
# d <- read.table("~/Projects/Botocudos/Files/Dstat/2019_05_19/natam.txt", 
#                 sep = "\t", stringsAsFactors = F)
# colnames(d) <- c("result","Pop1",  "Pop2", "Pop3",  
#                  "Pop4",  "Dstat","Z", "BABA",
#                  "ABBA",	"SNPs")
# d <- d[d$SNPs >10000,]
# panel <- read.table("~/Projects/Botocudos/Files/Panels/Magic.panel", header = T)
# 
# d$Pop1 <- sub("Conchal.*", "ConchaliSantiagoRM", d$Pop1)
# d$Pop2 <- sub("Conchal.*", "ConchaliSantiagoRM", d$Pop2)
# 
# d <- switchPops(d, "Botocudos")


```


```{r getmyDplot}
getMyDplot <- function(i , nCol, popTitle, pop1, set3 = "all", lwd = 1.2, cex = 1, archaic = T){

  myD <- organizeDstat(set1 = "all", set2 = "all", set3 = set3, d = d)
  myD <- myD[which(myD$Pop2 == popTitle), ]
  index <- keepOrRemove(group = pop1$Pop1, action = "keep", dataset = myD$Pop1)
  myD <- myD[index,]
  
  myD <- join(myD, pop1, by = "Pop1")
  myD$population <- myD$Pop3
  miniPanel <- unique(panel[,c("population", "region")])
  myD <- join(myD, miniPanel, by = "population")
  
  myD$color <- "gray65"
  if(archaic){
      myD$color[myD$Pop3 == "Denisova"] <- "brown1"
      myD$color[myD$Pop3 == "AltaiNean"] <- "blueviolet"
  }

  myD$color[myD$region == "Oceania"] <- "deepskyblue"

  myD <- myD[order(myD$color, decreasing = T),]
  
  if(i%%nCol == 1){
    par(mar = c(1,14,2,1)+0.1)
    plot(x = myD$Z, y = myD$index, bty = "n", col = myD$color, axes = F,
         xlab = NA, ylab = NA, xlim = c(-6,6), main = popTitle, type = "n")
    abline(h = pop1$index, lty = "dotted", col = "gray")
    points(x = -myD$Z, y = myD$index, col = myD$color, pch = 8, lwd = lwd, cex = cex)
    abline(v = c(-3.3, 3.3), lty = "dashed")
    axis(2, at = seq(1, nrow(pop1)), 
         las = 2,
         labels = pop1$Pop1)
  }else{
    par(mar = c(1,1,2,1)+0.1)
    plot(x = myD$Z, y = myD$index, bty = "n", col = myD$color, axes = F,
         xlab = NA, ylab = NA, xlim = c(-6,6), main = popTitle, type = "n")
    abline(h = pop1$index, lty = "dotted", col = "gray")
    points(x = -myD$Z, y = myD$index, col = myD$color,  pch = 8, lwd = lwd, cex = cex)
    abline(v = c(-3.3, 3.3), lty = "dashed")
  }
}
```



```{r}

set1 <- "called"
set2 <- "sampled"
set3 <- "all"

# index <- stratifyMe(set = set1, "keep", d$Pop1)
# myD <- d[index,]
# POP1 should contain poopulations in Y-axis (pop1)
index <- stratifyMe(set = set1, "keep", d$Pop1)
myD <- d[index,]

# POPT contains populations that will be displayed in the title
index <- stratifyMe(set = set2, "keep", d$Pop1)
myD <- d[index,]
POPT <- unique(myD$Pop1)

allPops <- unique(c(myD$Pop1, myD$Pop2))
names(allPops) <- 1:length(allPops)

POP1 <- data.frame(Pop1 = allPops, index = names(allPops))
POP1$meanNean <- sapply(POP1$Pop1, orderNean, myD = myD, POP = POPT)
POP1 <- POP1[order(POP1$meanNean),]
POP1 <- POP1[which(POP1$Pop1 != "Tlingit"),]
POP1 <- POP1[which(POP1$Pop1 != "Saqqaq"),]
POP1$index <- 1:nrow(POP1)


```

```{r}
s2 <- "called"
set3 <- "all"
for(s2 in c("called","sampled", "captured")){
    # POPT contains populations that will be displayed in the title (pop2)
    index1 <- stratifyMe(set = s2, "keep", d$Pop2)
    index2 <- stratifyMe(set = s2, "keep", d$Pop1)
    myD <- d[c(index1, index2),]
    
    allPops <- unique(c(myD$Pop2, myD$Pop1))
    allPops <- allPops[stratifyMe(s2, "keep", allPops)]
    names(allPops) <- 1:length(allPops)
    
    POPT <- unique(unique(myD$Pop1), unique(myD$Pop2))
    POPT <- POPT[stratifyMe(s2, "keep", POPT)]
    print(s2)
    print(POPT)
    nCells <- ceiling(sqrt(length(POPT)))
    for(s1 in c("called")){#, "captured", "sampled")){
      # name <- paste("~/Projects/Botocudos/Plots/Dstat/MidThesis/mini/title_",
      #               s1, "_yAxis_", s2, ".png", sep = "")
      # print(name)
      # png(name, width = 3+2*(nCells-1), 
      #     height = 2.5*nCells, res = 250, units = "in", bg = NA)
      # layout(matrix(seq(1, nCells**2), byrow = T, 
      #               nrow = nCells, ncol = nCells),
      #        widths = c(2,rep(1, nCells-1)), heights = rep(2, nCells))
      
      
      #nCells <- 1
      # POP1 should contain poopulations in Y-axis (pop1)
      index1 <- stratifyMe(set = s1, "keep", d$Pop1)
      index2 <- stratifyMe(set = s1, "keep", d$Pop2)
      myD <- d[c(index1, index2),]
      allPops <- unique(c(d$Pop1[index1],
                                       d$Pop2[index2]))
      POP1 <- data.frame(Pop1 = allPops, index = 1:length(allPops))
      
      #POP1$meanNean <- sapply(POP1$Pop1, orderNean, myD = myD, POP = POPT)
      #POP1 <- POP1[order(POP1$meanNean),]
      POP1 <- POP1[which(POP1$Pop1 != "Tlingit"),]
      POP1 <- POP1[which(POP1$Pop1 != "Saqqaq"),]
      POP1$Pop1 <- factor(POP1$Pop1, levels = c("USR1", "Anzick1", "SpCave",
                                                "Lovelock2", "Lovelock3",
                                                "Ayayema",
                                                "Quechua", "Aymara",
                                                "Mixe", "Huichol", "Piapoco", "Pima",
                                                "Athabascan", "Maya",
                                                "Taino","Chane", "Yukpa",
                                                "Karitiana", "Surui",
                                                "LagoaSta"))
      POP1 <- POP1[order(POP1$Pop1, decreasing = T),]
      
      POP1$index <- 1:nrow(POP1)
      for(i in 1:length(POPT)){
        name <- paste("~/Projects/Botocudos/Plots/Dstat/MidThesis/mini/arc_title_",
                      s2, POPT[i], "_yAxis_", s1, ".png", sep = "")
        print(name)
        png(name, width = 5, height = 4, res = 250, units = "in", bg = NA)
        # layout(matrix(seq(1, nCells**2), byrow = T, nrow = nCells, ncol = nCells),
        #        widths = c(2,rep(1, nCells-1)), heights = rep(2, nCells))
        print(POPT[i])
        getMyDplot(i = 1,#i, 
                   nCol = nCells, popTitle = POPT[i], 
                   pop1 = POP1, set3 = set3, 
                   lwd = 3, cex = 1.5, archaic = T)
         dev.off()
      }
      #dev.off()
    }
    
}

```


```{r rank tests for a pop}

Drank <- function(myD, pop1, pop3, set2){
  dsw <- switchPops(myD, pop1)
  index <- keepOrRemove(pop1, "keep", dsw$Pop1)
  D <- dsw[index,]
  index <- stratifyMe(set2, "keep", D$Pop2)
  D <- D[index,]
  pop2 <- unique(D$Pop2)
  
  getRank <- function(x, D){
    z <- D[D$Pop2 == x,]
    return(which(z[order(rank(z$Z)),]$Pop3 == pop3))
  }
  myRank <- sapply(pop2, function(x) getRank(x, D))
  myRank <- data.frame(pop2 = names(myRank), rank = myRank)
  return(myRank)
}

```

```{r Plot the different ranks for the Z-scores per population}

myAm <- unique(calledGenos$population[calledGenos$region == "Americas"])
myAm <- factor(myAm, levels = c("Anzick1", "USR1", "SpCave", "LagoaSta",
                                "Ayayema", "Saqqaq", "Lovelock2", "Lovelock3",
                                "Taino", "Aconcagua", "Quechua", "Aymara", "Athabascan",
                                "Tlingit", 
                                "Inuit", "Pima", "Mixe", "Huichol", "Maya", "Piapoco",
                                "Chane", "Yukpa","Karitiana", "Surui"),
               ordered = T)

popTitle <- "LagoaSta"
myRank <- Drank(myD, pop1 = popTitle, "AltaiNean", "called")

myRank$pop2 <- factor(myRank$pop2, levels = levels(myAm), ordered = T)

myRank <- myRank[order(myRank$pop2),]

    plot(x = myRank$rank, y = 1:nrow(myRank), bty = "n", axes = F,
         xlab = NA, ylab = NA, xlim = c(0, 110), main = popTitle, type = "n")
    #abline(h = , lty = "dotted", col = "gray")
    points(x = myRank$rank, y = 1:nrow(myRank),pch = 8, lwd = 1.2)
    
    abline(v = 55, lty = "dashed")
    axis(2, at = seq(1, nrow(myRank)), 
         las = 2,
         labels = myRank$pop2)
    
```



## Correlations
```{r}


myD <- switchPops(d, "Mixe")
myD <- myD[myD$Pop1 == "Mixe",] ;unique(myD$Pop1)
index <- stratifyMe("called", "keep", myD$Pop2)
mixe <- myD[index,]

myD <- switchPops(d, "LagoaSta")
myD <- myD[myD$Pop1 == "LagoaSta",] ;unique(myD$Pop1)
index <- stratifyMe("called", "keep", myD$Pop2)
lagoa <- myD[index,]

nean <- which(mixe$Pop3 == "AltaiNean")
mixe <- mixe[nean,]
nean <- which(lagoa$Pop3 == "AltaiNean")
lagoa <- lagoa[nean,]

mixe$ABBA <- mixe$ABBA/mixe$SNPs
mixe$BABA <- mixe$BABA/mixe$SNPs
lagoa$ABBA <- lagoa$ABBA/lagoa$SNPs
lagoa$BABA <- lagoa$BABA/lagoa$SNPs
j <- join(mixe[,c("Pop2","Pop3" ,"ABBA", "BABA")], lagoa[,c("Pop2", "ABBA", "Pop3","BABA")],
          by = "Pop2")

j$D1 <- j[,3]-j[,4]
j$D2 <- j[,5]-j[,7]
plot(j$D1[j[,2] == "UstIshim"], j$D2[j[,6] == "UstIshim"])
for(i in 2:4){
  for(k in 2:4){
    plot(j[,i], j[,k])
  }

}


j

plot(j$D1, j$D2)
cor(j[,2]-j[,3], j[,4]-j[,5])


```

# Microbiome

```{r}

```

```{r}
library(plyr)

# FAM file with the population name and individual ID
originalFam <- read.table(path_to_original)
colnames(originalFam) <- c("pop", "ind", seq(3, ncol(originalFam)))

# FAM file with apparent 0's in the population entry
myFam <- read.table(path_to_new_fam)
colnames(myFam) <- c("pop", "ind", seq(3, ncol(myFam)))

# Find index
index <- sapply(myFam$pop, function(x) which(originalFam$pop == x))

# Population names
newPop <- originalFam$pop[index]

```

