---
title: "Dstat_genotype_calling_comparison"
author: "Cruz-Dávalos, Diana I."
date: "10/14/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
```

```{r}
library(ggplot2)
library(plyr)
source("~/Projects/Botocudos/Scripts/Plotting/plot_F3.R")
source("~/Projects/Botocudos/Scripts/Plotting/plot_Dstats.R")
library(colorspace)
library(cowplot)
panel <- read.table("~/Projects/Botocudos/Files/Panels/Magic.panel", 
                    header = T, stringsAsFactors = F)
magic <- read.table("~/Projects/Botocudos/Files/Panels/Magic.panel", 
                    header = T, stringsAsFactors = F)
languages <- read.table("~/Projects/Botocudos/Files/Panels/Maanasa/Maanasa.americas.Moreno_names.txt", header = F, stringsAsFactors = F)
colnames(languages) <- c("population", "language")
```

## Z-scores from qpDstat

```{r}
# switch_h1_h2 <- function(Dstat, pop_in_h1){
#   idx_to_switch <- which(Dstat$h2 == pop_in_h1)
#   h2 <- Dstat$h1[idx_to_switch]
#   Dstat$h1[idx_to_switch] <- pop_in_h1
#   Dstat$h2[idx_to_switch] <- h2
#   Dstat$Z[idx_to_switch] <- -Dstat$Z[idx_to_switch]
#   return(Dstat)
# }
```

```{r}

Dstat <- read.table("~/Projects/Botocudos/Files/Dstat/2019_10_14/geno_calls/Moreno2019/h_merged_qpDstat.results",
                    stringsAsFactors = F)
colnames(Dstat) <- c("H1", "H2", "H3", "H4",
                     "D", "Z", "BABA", "ABBA", "nSites")
Dstat$population <- Dstat$H3

Dstat <- join(Dstat, magic[,c("population", "region")], by = "population")
```

```{r}

subset_dstat <- function(Dstat, pop){
  new_dstat <- Dstat[Dstat$H1 == pop | Dstat$H2 == pop,]
  new_dstat <- switch_h1_h2(Dstat = new_dstat, pop_in_h1 = pop)
  new_dstat$h2_h3 <- paste(new_dstat$H2, new_dstat$H3, sep = "_")
  new_dstat <- unique(new_dstat)
  return(new_dstat)
}

victor <- subset_dstat(Dstat, "LagoaSta")
diana <- subset_dstat(Dstat, "LagoaSanta_called")
```

```{r}
myComparison <- diana[,c("h2_h3", "Z", "population", "region")]
myComparison$diana <- myComparison$Z
myComparison$Z <- NULL

myComparison <- join(myComparison, victor[,c("h2_h3", "Z")], by = "h2_h3")
myComparison$victor <- myComparison$Z
myComparison <- myComparison[complete.cases(myComparison),]
```

```{r}

make_plot_comparison <- function(myComparison, region){
  if(sum(!(region %in% unique(myComparison$region))) > 0){
    myRegion <- "all other regions"
  }else{
    myRegion <- region
  }
  myComparison <- myComparison[myComparison$region %in% region,]
  myCor <- cor(myComparison$victor, myComparison$diana)
  
  main_txt <- paste("Z-scores for D(LagoaSta, NatAm,",
                    myRegion, ", Yoruba)", sep = "")
  
  plot(myComparison$victor, myComparison$diana,
       xlab = "H1 = LagoaSta, calls by Víctor", 
       ylab = "H1 = LagoaSta, calls by Diana", 
       main = main_txt,
       bty = "n", col = as.factor(myComparison$region), xlim = c(-12,6), ylim = c(-12,6))
  grid()
  abline(a = 0, b = 1, col = "red")
  text(paste("Correlation:", round(myCor,4)), 
       x = min(myComparison$victor), y = max(myComparison$diana))
}


```

```{r}
layout(matrix(seq(1,12), ncol = 4))
for(r in c("Siberia", "EastAsia", "Caucasus", "CentralAsia", "WestAsia", "Europe", "SouthAsia", "Oceania", "SEAsia", "Africa", "Archaic")){#{unique(myComparison$region)){
  make_plot_comparison(myComparison, r)
}
```

### Australasians vs others
```{r}
Dstat <- read.table("~/Projects/Botocudos/Files/Dstat/2019_10_14/h_merged_qpDstat_Oct22.results",
                    stringsAsFactors = F)
colnames(Dstat) <- c("H1", "H2", "H3", "H4",
                     "D", "Z", "BABA", "ABBA", "nSites")
Dstat$population <- Dstat$H3

Dstat <- join(Dstat, magic[,c("population", "region")], by = "population")

```

```{r, fig.width=12, fig.height=12}
column_to_highlight <- "region"
to_highlight <- c("Oceania", "SEAsia")
color_to_highlight <- c("purple", "pink")

pop_in_h3 <- unique(Dstat$H3)
pop_order <- unique(c(Dstat$H1, Dstat$H2))

myPops <- c("MN0008_L3U", "MN0008_3lanes",
            "Surui", "Karitiana", "LagoaSanta_called",
            "LagoaSta",
            "Mixe", "Huichol", "Quechua", "Ayayema")
layout(matrix(seq(1,10), ncol = 5, byrow = T))

for(pop_in_h2 in myPops){
  mySubset <- prepare_set(abba = Dstat, pop_in_h2 = pop_in_h2,
                        pop_in_h3 = pop_in_h3, 
                        pop_order = pop_order)
moreno_4a(abba = mySubset, pop = pop_in_h2,
          column_to_highlight = column_to_highlight, 
          to_highlight = to_highlight, 
          color_to_highlight = color_to_highlight)
}
pop_in_h2 <- "LagoaSanta_called"





```

## Relationship between Botocudos and other Native Americans

```{r, old results}

coordinates <- read.table("~/Projects/Botocudos/Files/Maps/Magic.coordinates", header = F)
colnames(coordinates) <- c("population", "lat", "long", "type")

f3 <- read.table("~/Projects/Botocudos/Files/F3/2019_05_17/boto_maanasa.txt",
                 header = T)
colnames(coordinates) <- c("Source2", "lat", "long", "type")
f3$Source2 <- sub("Conchal.*", "ConchaliSantiagoRM", f3$Source2)
f3 <- join(f3, coordinates, by = "Source2")
f3 <- f3[f3$SNPs >10000,]

f3$population <- f3$Source2

myPanel <- unique(panel[,c("population", "region")])
f3 <- join(f3, myPanel, by = "population")
 f3 <- join(f3, languages, by = "population")
f3 <- f3[f3$region == "Americas",]

```

```{r, fig.height=20, eval = F}
ciempies_plot(f3, main = NULL)
```

# F3
# Arara, Surui, LagoaSanta, Karitiana

```{r}
f3 <- read.table("~/Projects/Botocudos/Files/F3/2019_10_17/f3.results",
                 header = T, stringsAsFactors = F)

f3$Source2 <- sub("Conchal.*", "ConchaliSantiagoRM", f3$Source2)
f3 <- join(f3, coordinates, by = "Source2")
f3 <- f3[f3$SNPs >10000,]

f3$population <- f3$Source2

```

```{r, fig.width=12, fig.height=14}
myPops <- c("Botocudos", "L3U",
            "Arara", "LagoaSanta", "Surui",
             "Karitiana")
# pdf("~/Projects/Botocudos/Plots/F3/2019_10_17/f3_bidir.pdf",
    # width = 12, height = 14)
layout(matrix(seq(1,6), byrow = T, ncol = 6))
for(p in myPops){
  ciempies_plot(f3[f3$Source1 == p,], main = p, 
                doLayout = F, doBar = F, cex.axis = 0.7)
}
# dev.off()
```


# ANGSD, error corrected
  # do not trim 5bp, 
  # MN0008
  # All botocudos
  # no need to rmTrans
  
  # Fig. 3C
questions is:
Compared to other Native Americans, how close are botocudos to Mixe or to Lagoa Santa?


```{r}
nat_americans <- list("northern" = c("Lovelock2", "Lovelock3",
                                     "Pima", "Huichol", "Maya", "Taino"),
                      "east_andes" = c("Yukpa", "Aconcagua",
                                       "Aymara", "Quechua", 
                                       "Chane", "Surui", "Karitiana",
                                       "Piapoco"),
                      "west_andes" = c("Ayayema", "PtaStaAna",
                                       "AncKaweskar2", "AncKaweskar3",
                                       "AncYamana", "AncYamana2"),
                      "botocudos" = c("Botocudos"))
```


```{r}

```


### rmTrans
```{r, fig.width=12, fig.height=6}
abba <- read.table("~/Projects/Botocudos/Files/Dstat/2019_10_14/Corrected/23Boto_inc.ErrorCorr.TransRem.txt",
                   header = T, stringsAsFactors = F)

# first plot is D(NatAm, Mixe; LagoaSanta_15x, Yoruba)
myAlpha <- 0.001
pop_order <- rev(c("Botocudos", "MN0008_L3U",
               "Surui", "Karitiana", "Ayayema", "Aymara","Huichol"))



layout(matrix(seq(1,2), byrow = T, ncol = 2))
first_set <- prepare_set(abba, pop_in_h2 = "Mixe", 
                         pop_in_h3 = "LagoaSanta_15x",
                         pop_order = pop_order)


cex <- 1.2
lwd <- 2
par(mar = c(5,8,4,2))
moreno_3c(first_set, 
          title =  "D(NatAm, Mixe; LagoaSanta_15x, Yoruba)\n(rmTrans)",
          lwd = lwd, cex = cex)

# second plot is D(NatAm, LagoaSanta_15X; Mixe, Yoruba)
second_set <- prepare_set(abba, pop_in_h3 = "Mixe", 
                         pop_in_h2 = "LagoaSanta_15x",
                         pop_order = pop_order)

par(mar = c(5,2,4,8))
moreno_3c(second_set, 
          title =  "D(NatAm, LagoaSanta_15x; Mixe, Yoruba)\n(rmTrans)",
          yaxis = F, lwd = lwd, cex = cex)


```

### all
```{r,  fig.width=12, fig.height=6}
abba <- read.table("~/Projects/Botocudos/Files/Dstat/2019_10_14/Corrected/23Boto_inc.ErrorCorr.txt",
                   header = T, stringsAsFactors = F)

# first plot is D(NatAm, Mixe; LagoaSanta_15x, Yoruba)
myAlpha <- 0.001
pop_order <- rev(c("Botocudos", "MN0008_L3U",
               "Surui", "Karitiana", "Ayayema", "Aymara","Huichol"))

cex <- 1.2
lwd <- 2
color <- "#0057ba"
layout(matrix(seq(1,2), byrow = T, ncol = 2))
first_set <- prepare_set(abba, pop_in_h2 = "Mixe", 
                         pop_in_h3 = "LagoaSanta_15x",
                         pop_order = pop_order)

par(mar = c(5,8,4,2))

moreno_3c(first_set, 
          title =  "D(NatAm, Mixe; LagoaSanta_15x, Yoruba)\n(all)",
          xlim = c(-0.7, 0.05), cex = cex, lwd = lwd,
          color = color)

# second plot is D(NatAm, LagoaSanta_15X; Mixe, Yoruba)
second_set <- prepare_set(abba, pop_in_h3 = "Mixe", 
                         pop_in_h2 = "LagoaSanta_15x",
                         pop_order = pop_order)
par(mar = c(5,2,4,8))

moreno_3c(second_set, 
          title =  "D(NatAm, LagoaSanta_15x; Mixe, Yoruba)\n(all)",
           cex = cex, lwd = lwd,
          color = color, yaxis = F)

```


## Figure S35A

### Error-corrected D-stats, rmTrans

```{r, fig.width=12, fig.height=12}
abba <- read.table("~/Projects/Botocudos/Files/Dstat/2019_10_14/Corrected/S35A_Bot8.ErrorCorr.TransRem.txt",
                   header = T, stringsAsFactors = F)

pops_in_title <- c("Botocudos", 
                   "MN0008_L3U", "LagoaSanta_15x",
                   "Surui", "Karitiana",
                   "Mixe", "Huichol", 
                   "Aymara", "Ayayema")

pops_in_y <- pops_in_title
pops_in_h3 <- c("French", "Han",
                          "Andaman", "WCDesert")
pop_colors <- c("deepskyblue1", "darkgoldenrod1", 
                "pink", "blueviolet", 
                "black")

plot_list <- moreno_S35A(abba = abba, 
            pops_in_title = pops_in_title,
            pops_in_y = pops_in_y,
            pops_in_h3 = pops_in_h3,
            pop_colors = pop_colors,
            rm_legend = T)

myPlot <- moreno_S35A(abba = abba,
                      pops_in_title = c("Botocudos"),
                      pops_in_y = pops_in_y,
                      pops_in_h3 = pops_in_h3,
                      pop_colors = pop_colors, rm_legend = F)

legend <- get_legend(myPlot[[1]])
plot_list[[length(plot_list)+1]] <- legend
plot_grid(plotlist = plot_list, ncol = 3)

```

#### Affinities of Native Americans to non-NA (vs Mixe) 
```{r, fig.width=12, fig.height=12}

pops_to_plot <- c("WCDesert", "Andaman","Han", "French")
colors <- c("blueviolet", "pink", 
             "darkgoldenrod1", "deepskyblue1", 
                "black")

layout(matrix(seq(1,4), byrow = T, ncol = 2))
for(i in 1:4){
  pop <- pops_to_plot[i]
  color <- colors[i]
  myAbba <- prepare_set(abba, pop_in_h2 = "Mixe", pop_in_h3 = pop, pop_order = c("Botocudos", 
                                                                                        "MN0008_L3U", "LagoaSanta_15x",
                                                                                        "Surui", "Karitiana",
                                                                                        "Mixe", "Huichol", 
                                                                                        "Aymara", "Ayayema"))
  
  myAbba <- myAbba[complete.cases(myAbba),]
  
  moreno_3c(myAbba, 
            title = paste("D(NatAm, Mixe; ", pop, ", Yoruba)"),
            color = color)
}


```

#### Affinities of Native Americans to non-NA (vs Lagoa Santa)

```{r, fig.width=12, fig.height=12}
pops_to_plot <- c("WCDesert", "Andaman","Han", "French")
colors <- c("blueviolet", "pink", 
             "darkgoldenrod1", "deepskyblue1", 
                "black")

layout(matrix(seq(1,4), byrow = T, ncol = 2))
for(i in 1:4){
  pop <- pops_to_plot[i]
  color <- colors[i]
  myAbba <- prepare_set(abba, pop_in_h2 = "LagoaSanta_15x", pop_in_h3 = pop, pop_order = c("Botocudos", 
                                                                                        "MN0008_L3U", "LagoaSanta_15x",
                                                                                        "Surui", "Karitiana",
                                                                                        "Mixe", "Huichol", 
                                                                                        "Aymara", "Ayayema"))
  
  myAbba <- myAbba[complete.cases(myAbba),]
  
  moreno_3c(myAbba, 
            title = paste("D(NatAm, LagoaSta; ", pop, ", Yoruba)"),
            color = color)
}


```

## Error-corrected D-stats, all type of polymorphisms
```{r, fig.width=12, fig.height=12}
abba <- read.table("~/Projects/Botocudos/Files/Dstat/2019_10_14/Corrected/S35A_Bot8.ErrorCorr.txt",
                   header = T, stringsAsFactors = F)

plot_list <- moreno_S35A(abba = abba, 
            pops_in_title = pops_in_title,
            pops_in_y = pops_in_title,
            pops_in_h3 = pops_in_h3,
            pop_colors = pop_colors,
            rm_legend = T)
myPlot <- moreno_S35A(abba = abba,
                      pops_in_title = c("Botocudos"),
                      pops_in_y = pops_in_y,
                      pops_in_h3 = pops_in_h3,
                      pop_colors = pop_colors, rm_legend = F)

legend <- get_legend(myPlot[[1]])
plot_list[[length(plot_list)+1]] <- legend
plot_grid(plotlist = plot_list, ncol = 3)

```


# D-stats, called genotypes
```{r}

```


# qpGraph
reporta modelo más sencillo, sin admixtures
de todos modos intenta todas las combinaciones

fit score es usado como log-likelihood

fija los admixtures proportions para todas las ramas, excepto para las que estás preguntando

# Admixture graph
# 2012 Reich
# India - qpWave or qpGraph
# Heterozygosity

```{r}
het_dir <- "~/Projects/Botocudos/Files/Heterozygosity/2019_10_30/"

files <- list.files(het_dir, pattern = "Nigeria") ; files

get_samples <- function(pattern){
  x <- sapply(strsplit(files[grep(pattern, files)], "_"), 
              function(x) c(x[[1]], x[[2]]))
  
  samples <- unique(c(x[1,], x[2,]))
  return(samples)
}


boto_sm <- get_samples(pattern = "MN")
wcd_sm <- get_samples("WCD")
surui_sm <- get_samples("Surui")
lovelock_sm <- get_samples("Lovelock")
mixe_sm <- get_samples("Mixe")
maya_sm <-get_samples("Maya")
pima_sm <- get_samples("Pima")
usr_sm <- get_samples("USR")

mbuti_sm <- c("HGDP00456", "SS6004471")
dai_sm <- c("SS6004467", "HGDP01307")
french_sm <- c("SS6004468", "HGDP00521")
han_sm <- c("SS6004469", "HGDP00778")
mandenka_sm <- c("SS6004470", "HGDP01284")
papuan_sm <- c("SS6004472", "HGDP00542")
san_sm <- c("SS6004473", "HGDP01029")
sardinian_sm <- c("SS6004474", "HGDP00665")
yoruba_sm <- c("SS6004475", "HGDP00927")
karitiana_sm <- c("SS6004476", "HGDP00998")
australian_sm <- c("SS6004477", "SS6004478")
lagoa_sm <- c("LagoaSta", "LS4", "LS6", "LS7", "LS8")

myGroups <- list(
                 WCD = wcd_sm,
                 Surui = surui_sm,
                 Lovelock = lovelock_sm,
                 Mixe = mixe_sm,
                 Maya = maya_sm,
                 Pima = pima_sm,
                 USR = usr_sm,
                 Mbuti = mbuti_sm,
                 Dai = dai_sm,
                 French = french_sm,
                 Han = han_sm,
                 Mandenka = mandenka_sm,
                 Papuan = papuan_sm,
                 San = san_sm,
                 Sardinian = sardinian_sm,
                 Yoruba = yoruba_sm,
                 Karitiana = karitiana_sm,
                 Australian = australian_sm,
                 LagoaSta = lagoa_sm,
                 Botocudos = boto_sm
                 )
myRegions <- matrix(data = c("WCD", "Australian_WCD", "Oceania",
                             "Surui", "Surui", "Americas",
                             "Lovelock", "Lovelock", "Americans_anc",
                             "Mixe", "Mixe", "Americas",
                             "Maya", "Maya", "Americas",
                             "Pima", "Pima", "Americas",
                             "USR", "USR", "Americas_anc",
                             "Mbuti", "Mbuti", "Africa",
                             "Dai", "Dai", "SotuhAsia",
                             "French", "French", "Europe",
                             "Han", "Han", "EastAsia",
                             "Mandenka", "Mandenka", "Africa",
                             "Papuan", "Papuan", "Oceania",
                             "San", "San", "Africa",
                             "Sardinian", "Sardinian", "Europe",
                             "Yoruba", "Yoruba", "Africa",
                             "Karitiana", "Karitiana", "Americas",
                             "Australian", "Australian", "Oceania",
                             "LagoaSta", "LagoaSanta", "Americas_anc",
                             "Botocudos", "Botocudos", "Botocudos"
                             ), ncol = 3, byrow = T)

myRegions <- data.frame(myRegions)
colnames(myRegions) <- c("group", "population", "region")
```

```{r}
match_files <- function(group, files){
  x <- sapply(strsplit(files, "_"), 
              function(x) x[[1]])

  index <- which(x %in% group)
  return(index)
}

lagoa_file <- match_files(lagoa_sm, files = files)

```

```{r}
num_files <- function(files){
  myTable <- table(sapply(files, function(x) strsplit(x, "_")[[1]][1]))
  return(myTable)
}
```

```{r}
blank_plot <- function(){
  plot(1,1, type = "n", axes = F, xlab = NA, ylab = NA)
}
```

```{r}
plot_pairwise_het <- function(samples){
  index <- match_files(samples, files = files)
  
  sub_files <- files[index]
  nFiles <- num_files(sub_files)
  nFiles <- nFiles[order(nFiles, decreasing = T)]
  individuals <- names(nFiles)[order(nFiles, decreasing = T)]
  individuals2 <- c(individuals, samples[!samples %in% individuals])
  nInd <- length(individuals)+1
  
  layout(matrix(1:((nInd-1)*(nInd-1)), byrow = T, ncol = nInd-1))
  
  for(ind1 in individuals){
    individuals2 <- individuals2[!( individuals2 == ind1) ]
    for(ind2 in individuals2){
      to_read <- paste(het_dir, ind1, 
                       "_", ind2, "_Nigeria_B_Yoruba-3.pi.stats.txt",
                       sep = "")
      if(file.exists(to_read)){
        print(to_read)
        myPi <- read.table(to_read, header = F)
        colnames(myPi) <- c("estimate",	"bias",	"stderror",	"confint1",	"confint2")
        
        plot(myPi$estimate, ylim = c(0, 0.5), axes = F,
             xlab = ind1, ylab = ind2)
        segments(x0 = 1, y0 = myPi$confint1, y1 = myPi$confint2)
        axis(2)
        grid()
      }
      else{
        blank_plot()
      }
    }
    if(ind1 != tail(individuals,1)){
      to_plot <- nInd - length(individuals2) -1
      while(to_plot > 0){
        to_plot <- to_plot -1
        blank_plot()
      }
    }
    
  }
}
```

```{r}
plot_pairwise_het(karitiana_sm)

for(group in names(myGroups)){
  nInd <- length(myGroups[[group]])
  png(paste("~/Projects/Botocudos/Plots/Het/2019_11_01/", group, ".pi.pairwise.png", sep = ""),
      width = 2*nInd, height = 2*nInd, re = 250, units = "in")
  plot_pairwise_het(myGroups[[group]])
  dev.off()
}

```

```{r}
myFile <- file[1]
myPi <- read.table(paste(het_dir, myFile, sep = ""), header = F)

ind1 <- strsplit(myFile, "_")[[1]][1]
ind2 <- strsplit(myFile, "_")[[1]][2]


```

```{r}

match_groups <- function(file, groups, regions){
  ind <- strsplit(file, "_")[[1]][1]
  for(group in names(groups)){
    if(ind %in% groups[[group]]){
      return(regions[regions$group == group,])
    }
  }
}

x <- match_groups(files[1], myGroups, myRegions)

```

```{r}
myPi <- data.frame()
for(f in files){
  x <- read.table(paste(het_dir, f, sep = ""))
  regions <- match_groups(f, myGroups, myRegions)
  colnames(x) <- c("estimate",	"bias",	"stderror",	"confint1",	"confint2")
  x <- cbind(x, regions)
  myPi <- rbind(myPi, x)
}


```

```{r}
ggplot(myPi, aes(x = population, y = estimate, col = region)) +
  #geom_point() +
  geom_jitter() +
  geom_violin()

```

