---
title: "DataQuality"
author: "DI Cruz Dávalos"
date: "7/26/2019"
output:
  pdf_document: default
  html_document: default
---

# Results

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message = F, warning = F, fig.width = 16)
```

```{r}
library(plyr)
library(ggplot2)
library(scales)
library(cowplot)
library(ggridges)
source("~/Projects/Botocudos/Scripts/Plotting/length_distribution_plot.R")
source("~/Projects/Botocudos/Scripts/Plotting/mapDamage_plot.R")
source("~/Projects/Botocudos/Scripts/Complexity/preseq_functions.R")
# Order in which DNA libraries will be displayed
lib_order <- c("All", "mtCapture", "L1", "L2", "L3U")
colors_per_library <- c("gray40", "#c4bf33", "#b08699","#e0b81f","#0d75ff")
names(colors_per_library) <- lib_order
```

```{r}
# Plot for work in progress
workInProgress <- function(gg = FALSE){
  if(gg){
    X <- data.frame(x = 1, y = 1, text = "Work in progress")
    ggplot(X, aes(x = x, y = y)) + geom_text(label = X$text)
  
    }else{
    plot(1,1, type = "n")
    text(1,1, "Work in progress")
  }
}
```

```{r}
plot_letters <- function(letter = "A", at = -1, line = 1){
  mtext(letter, side = 3, line = line, at = at, adj = 0, cex = 1.3, font = 2)
}
```

```{r}
paste_path <- function(end, path = "~/Projects/Botocudos/Plots/"){
  return(paste(path, end, sep = ""))
}
  
```

```{r, read in summary files}
myDir <- "~/Projects/Botocudos/Files/Summaries/2019_07_26/"

files <- list.files(path = myDir, pattern = ".new.summary$")

boto <- data.frame()

for(file in files){
  tmp_summary <- read.table(paste(myDir, file, sep =""), header = T, sep = "\t")
  boto <- rbind(boto, tmp_summary)
}

boto$library <- sub("\\*", "All", boto$library)
boto$library <- sub("S1", "L1", boto$library)

boto$library <- factor(boto$library, levels = lib_order, ordered = T)

```

```{r, merge isotopic data}
backbone <- read.csv("~/Projects/Botocudos/Files/Tables/Backbone_Table1.csv")

colnames(backbone) <- c("sample", "CalibratedDate", "Area", "State") 

boto <- join(boto, backbone, by = "sample")

boto <- boto[order(boto$hits_unique_frac_endogenous, boto$library),]
boto$sample <- factor(boto$sample, levels = unique(boto$sample), ordered = T)
```

# Sequenced reads

```{r}
# ggplot(boto[boto$library != "All",], aes(y = seq_reads_se, x = sample, fill = library)) +
#   geom_bar(stat = "identity") +
#   theme_void()


data <- reshape(boto[,c("sample", "seq_reads_se", "library")],
               timevar = "library", direction = "wide", 
               idvar = c("sample"))

myData <- t(as.matrix(data[,2:ncol(data)]))
colnames(myData) <- data$sample
myData[is.na(myData)] <- 0
rownames(myData) <- sub(".*\\.", "", rownames(myData))

myData <- myData[match(lib_order, rownames(myData)),]


```

## Total number of reads

```{r}
myRows <- which(rownames(myData) != "All")
#colors <-  c("#b08699","#e0b81f","#0d75ff")

sequenced <- myData[myRows,]
myOrder <- match(rownames(sequenced), names(colors_per_library))
# png(filename = paste_path("Sequencing/total_reads.png"),
#     width = 7, height = 4, res = 250, units = "in", bg = NA)
pdf(paste_path("Sequencing/total_reads.pdf"),
    width = 9, height = 5)
barplot(sequenced, col = NA, border = NA, 
        main = "Total sequenced reads (per library)", las = 2, axisnames =F)
  grid(nx = 0, ny = NULL)
  barplot(sequenced, col = colors_per_library[myOrder], border = NA, 
        main = "Total sequenced reads (per library)", las = 2, add = T)
legend(x = "topleft", legend = rownames(sequenced),
       fill = colors_per_library[myOrder], border = NA, bty = "n", title = "Library")

dev.off()
```

## Reads retained after trimming

```{r}
data <- reshape(boto[,c("sample", "seq_reads_se", "seq_retained_reads",
                        "library")],
               timevar = "library", direction = "wide", 
               idvar = c("sample"))

myData <- t(as.matrix(data[,2:ncol(data)]))
colnames(myData) <- data$sample
myData[is.na(myData)] <- 0

rownames(myData) <- sub(".*\\.", "", rownames(myData))

myData <- myData[rep(match(lib_order, rownames(myData)), each = 2) + rep(c(0,1), 5),]
```

```{r}
#colors <-  c("#b08699","#e0b81f","#0d75ff")
myRows <- seq(2, nrow(myData), 2)
myRows <- myRows[!(rownames(myData)[myRows] %in% "All")]

retained <- myData[myRows,]
myOrder <- match(rownames(retained), names(colors_per_library))
barplot(myData[myRows,], col = colors_per_library[myOrder], border = NA, 
        main = "Retained reads after trimming", las = 2)
legend(x = "topleft", legend = rownames(myData)[myRows],
       fill = colors_per_library[myOrder], border = NA, bty = "n", title = "Library")

```

```{r}
myData[myRows-1,][myData[myRows-1,] == 0] <- 1


retained_frac <- myData[myRows,]/myData[myRows-1,]
myOrder <- match(rownames(retained), names(colors_per_library))

barplot(myData[myRows,]/myData[myRows-1,], col = colors_per_library[myOrder], border = NA, 
        main = "Retained reads after trimming (fraction)", las = 2, beside = T,
        ylim = c(0, 1), yaxt = "n")

axis(2, at = seq(0,1, 0.1), labels = percent(seq(0, 1, 0.1)), las =1)

legend(x = "topright", legend = rownames(myData)[myRows],
       fill = colors_per_library[myOrder], border = NA, bty = "n", title = "Library")

```

```{r, eval = F}
  # Three plots in one figure

# png(paste_path("Sequencing/total_reads.png"),
#     width = 6, height = 7.5, res = 250, units = "in", bg = NA)
pdf(paste_path("Sequencing/total_reads.pdf"),
    width = 6, height = 7.5)
  layout(matrix(seq(1, 3), byrow = T, nrow = 3, 
                  ncol = 1), widths = c(3), heights = c(1,1,1))
  
  # Sequenced
  barplot(sequenced, col = NA, border = NA, 
          main = "Total sequenced reads", las = 2, beside = F,
           yaxt = "n", axisnames = F)
  grid(nx = 0, ny = NULL)
  barplot(sequenced, col = colors_per_library[myOrder], border = NA, 
          main = "Total sequenced reads", las = 2, beside = F,
           yaxt = "n", add = T)
  axis(2, las =1)
  legend(x = "topleft", legend = rownames(myData)[myRows],
         fill = colors_per_library[myOrder], border = NA, bty = "n", title = "Library")
  plot_letters(letter = "A", at = -3, line = 1.5)
  
  # Retained
  barplot(retained, col = NA, border = NA, 
          main = "Retained reads after trimming", las = 2, beside = F,
          yaxt = "n", axisnames = F)
  grid(nx = 0, ny = NULL)
  barplot(retained, col = colors_per_library[myOrder], border = NA, 
          main = "Retained reads after trimming", las = 2, beside = F,
          yaxt = "n", add = T)
  axis(2, las =1)
  plot_letters(letter = "B", at = -3, line = 1.5)
  
  # Retained, fraction
  barplot(retained_frac, col = NA, border = NA, 
          main = "Retained reads after trimming (percentage)", las = 2, beside = T,
          ylim = c(0, 1), yaxt = "n", axisnames = F)
  grid(nx = 0, ny = NULL)
  barplot(retained_frac, col = colors_per_library[myOrder], border = NA, 
          main = "Retained reads after trimming (percentage)", las = 2, beside = T,
          ylim = c(0, 1), yaxt = "n", add = T)
  axis(2, at = seq(0,1, 0.1), labels = percent(seq(0, 1, 0.1)), las =1)
    
  plot_letters(letter = "C", at = -12, line = 1.5)

dev.off()
```


## Endogenous content

```{r}
data <- reshape(boto[,c("sample", "hits_unique_frac_endogenous",
                        "library")],
               timevar = "library", direction = "wide", 
               idvar = c("sample"))

myData <- t(as.matrix(data[,2:ncol(data)]))
colnames(myData) <- data$sample

rownames(myData) <- sub(".*\\.", "", rownames(myData))

myData <- myData[rev(match(rownames(myData), lib_order)),]

```

```{r}
myRows <- which(rownames(myData) == "All")


maxEndo <- 0.6
barplot(myData[myRows,], border = NA, 
        main = "Endogenous content per sample", las = 2,
        ylim = c(0, maxEndo), yaxt = "n")
axis(2, at = seq(0, 0.5, 0.1), labels = percent(seq(0, 0.5, 0.1)))

```

```{r}
myRows <- seq(1, nrow(myData))

myData <- myData[(match(lib_order, rownames(myData))),]
myOrder <- match(rownames(myData), names(colors_per_library))
# png(paste_path("Sequencing/endogenous.png"), width = 7, height = 4,
#     res = 250, units = "in", bg = NA)
pdf(paste_path("Sequencing/endogenous.pdf"), width = 7, height = 4)
  barplot(myData, border = NA, 
          main = "Endogenous content per library", col = NA,  axes = FALSE, las = 2,
          xlab = NA, beside = T, axisnames = F, ylim = c(0, maxEndo))
  grid(nx = 0, ny = NULL)
  barplot(myData, border = NA, 
          las = 2, add =T,
          ylim = c(0, maxEndo), yaxt = "n", beside = T, col = colors_per_library[myOrder])
  axis(2, at = seq(0, 0.5, 0.1), labels = percent(seq(0, 0.5, 0.1)))
  legend(x = "topleft", legend = rownames(myData),
           fill = colors_per_library[myOrder], border = NA, bty = "n", title = "Library")
dev.off()
```

## Read length

```{r}
# ggplot(l, aes(x = color, y = Length, weight = Freq)) + 
#   geom_violin(trim = F) +
#   geom_jitter()

data  <- reshape(boto[,c("sample", "seq_retained_length",
                                "hits_length_endogenous",
                        "library")],
               timevar = "library", direction = "wide", 
               idvar = c("sample"))

myData <- t(as.matrix(data[,2:ncol(data)]))
colnames(myData) <- data$sample

old_names <- rownames(myData)
rownames(myData) <- sub(".*\\.", "", rownames(myData))

# I want to get:
# 3,4,5,6,1,2,7,8
myData <- myData[rep(match(lib_order, rownames(myData)), each = 2) + rep(c(0,1)),]

#myData[is.na(myData)] <- 0
rownames(myData)[seq(1, nrow(myData), 2)] <- sub("^", "Retained, ", rownames(myData)[seq(1, nrow(myData), 2)])
rownames(myData)[seq(2, nrow(myData), 2)] <- sub("^", "Mapped, ", rownames(myData)[seq(2, nrow(myData), 2)])


# myData <- myData[rev(seq(1, nrow(myData))),]
```

```{r}
colors <- rep(colors_per_library, each = 2)
colors[seq(1,length(colors),2)] <- alpha(colors[seq(1,length(colors),2)], 0.3)
myRows <- which(rownames(myData) != "All")

density <- rep(40, 3)
angle1 <- rep(c(0, 135), 3)

# png(paste_path("Sequencing/length_ret_map.png"), width = 9,
#     height = 5, res = 250, units = "in", bg = NA)
 pdf(paste_path("Sequencing/length_ret_map.pdf"), width = 12,
     height = 6)
  barplot(myData, border = "black", 
          main = "Average read length (retained and mapped)", las = 2,
           beside = T, col = NA, ylab = "Length (bp)", space = c(0, 1.5),
          ylim = c(0,90), axisnames = F
          #density = density, angle = angle1
          )
  grid(nx = 0, ny = NULL)
  barplot(myData, border = "black", 
          main = "Average read length (retained and mapped)", las = 2,
           beside = T, col = colors, ylab = "Length (bp)", space = c(0, 1.5),
          ylim = c(0,90), add = T
          #density = density, angle = angle1
          )
  legend(x = "top", legend = rownames(myData),
         fill = colors, bty = "n", title = "", horiz = F, ncol = 5)
dev.off()
#axis(2, at = seq(0, 0.5, 0.1), labels = percent(seq(0, 0.5, 0.1)))

```

```{r, fig.width=20, fig.height=4}
individuals <- unique(boto$sample)
nInd <- length(individuals)
nCol <- max(rowSums(table(boto[,c("sample", "library")]))) - 1

# png(paste_path("Sequencing/length_dist.png"), width = 12, height = 14,
    # res = 150, units = "in")
pdf(paste_path("Sequencing/length_dist.pdf"), width = 10, height = 12)
layout(matrix(seq(1, 56), byrow = T, ncol = 7),
       widths = rep(1, 7), heights = rep(1, 1))
i <- 0
for(ind in individuals){
  par(mar = c(4,4,2,2))
  plot(1,1, bty = "n", axes = F, type = "n", xlab = NA, ylab = NA)
  text(1,1, ind)
  i <- i+1
  plotted <- 6

  path <- paste("~/Projects/Botocudos/Files/Length/2019_07_30/", 
                ind,sep = "")
  files <- unique(boto$library[boto$sample == ind])
  files <- files[files != "All"]
  nCol <- length(files)
  files <- files[order(files)]

  for(lib in files){
    # print(plotted)
    for(type in c("nuc", "mito")){
      plotted <- plotted - 1
      gglength(ind = ind, type = type, d = path,
               lib = lib, gg = F, cex = 0.5, 
               plot.title = paste(lib, type, sep = ", "))
    }
    # Blank plot
  }
      while(plotted>0 ){
        plotted <- plotted - 1 
        plot(1,1, bty = "n", axes = F, type = "n", xlab = NA, ylab = NA)
    }
}

 dev.off()
```

## Molecular damage

```{r, fig.width=21, fig.height=3}

# png(paste_path("Sequencing/damage.png"), width = 16, height = 20,
#     res = 150, units = "in")
pdf(paste_path("Sequencing/damage.pdf"), width = 12, height = 16)

layout(matrix(seq(1, 70), byrow = T, 
              ncol = 7), widths = rep(3, 7), heights = rep(1, 10))
par(mar = c(1,1,1,1))
asis <- T
individuals <- unique(boto$sample)
for(ind in individuals){
  plot(1,1,type = "n", bty = "n", axes= F, xlab = NA, ylab = NA)
  text(1,1,ind)
  
  plotted <- 3
  # layout(matrix(seq(1, 6), byrow = T, nrow = 10, 
  #             ncol = 6), widths = rep(3, 8), heights = rep(1, 10))
  libs <- unique(boto$library[boto$sample == ind & boto$library != "All" & boto$library != "*"])
  type <- NULL
  d <- "~/Projects/Botocudos/Files/mapDamage/2019_08_08/"
  libs <- libs[order(libs)]
  
  # layout(matrix(seq(1, 6), byrow = T, nrow = 3, 
  #               ncol = 2), widths = rep(3,2), heights = rep(1, 3))
  for(lib in libs){
    plotted <- plotted -1
    sufix <- paste("/", lib, "/", lib, "_results_mapDamage", sep = "")
    damage_plot(ind = ind, type = type, lib, d = d,
                asis = asis, sufix = sufix, plot.title = lib)
  }
  while(plotted >0){
    plotted <- plotted -1
    plot(1,1, type = "n", xlab = NA, ylab = NA, bty = "n", axes = F)
    plot(1,1, type = "n", xlab = NA, ylab = NA, bty = "n", axes = F)
  }
}
dev.off()
```

## Genome coverage

```{r}
data <- data <- reshape(boto[,c("sample", 
                                "hits_coverage_nuclear",
                        "library")],
               timevar = "library", direction = "wide", 
               idvar = c("sample"))

myData <- t(as.matrix(data[,2:ncol(data)]))
colnames(myData) <- data$sample

#myData[is.na(myData)] <- 0
rownames(myData) <- sub(".*\\.", "", rownames(myData))
```

```{r}
myData <- myData[match(lib_order, rownames(myData)),]
myOrder <- match(rownames(myData), names(colors_per_library))
barplot(myData, border = NA, 
        main = "Depth of coverage (nuclear)", las = 2,
         beside = T, col = colors_per_library[myOrder], ylab = "Depth of coverage", 
        #density = density, angle = angle1
  )

legend(x = "topright", legend = rownames(myData),
       fill = colors_per_library[myOrder], bty = "n", title = "", border = NA)

genomeCov <- myData
```

```{r}

#myRows <- match(c("L1", "L2", "L3U", "All"), rownames(myData))
barplot(myData, border = NA, 
        main = "Depth of coverage (nuclear, log-scale)", las = 2,
         beside = T, col = colors_per_library[myOrder], ylab = "Depth of coverage", 
        #density = density, angle = angle1
        log = "y")

legend(x = "topright", legend = rownames(myData),
       fill = colors_per_library[myOrder], bty = "n", title = "", border = NA)

```


```{r}
data <- data <- reshape(boto[,c("sample", 
                                "hits_coverage_mitochondrial",
                        "library")],
               timevar = "library", direction = "wide", 
               idvar = c("sample"))

myData <- t(as.matrix(data[,2:ncol(data)]))
colnames(myData) <- data$sample

#myData[is.na(myData)] <- 0
rownames(myData) <- sub(".*\\.", "", rownames(myData))
```

```{r, eval = F}
myData <- myData[match(lib_order, rownames(myData)),]
myOrder <- match(rownames(myData), names(colors_per_library))
barplot(myData, border = NA, 
        main = "Depth of coverage (mitochondrial)", las = 2,
         beside = T, col = colors_per_library[myOrder], ylab = "Depth of coverage", log = "y"
        #density = density, angle = angle1
  )

legend(x = "topright", legend = rownames(myData),
       fill = colors_per_library[myOrder], bty = "n", title = "", border = NA)

mitoCov <- myData
```


```{r, eval = F}

# png(paste_path("Sequencing/coverage.png"), width = 7, height = 7,
#     res = 200, units = "in", bg = NA)  
pdf(paste_path("Sequencing/coverage.pdf"), width = 7, height = 7)
myOrder <- match(rownames(genomeCov), names(colors_per_library))

par(mfrow = c(2,1))
  barplot(genomeCov, border = NA, col = NA,  axes = FALSE, las = 2,
          xlab = NA, beside = T, axisnames = F, 
          main = "Depth of coverage (nuclear)")
  grid(nx = 0, ny = NULL)
  barplot(genomeCov, border = NA, las = 2,
          beside = T, col = colors_per_library[myOrder], 
          ylab = "Depth of coverage", add = T
  )
  plot_letters("A", at = -12, line = 1.5)
  legend(x = "topleft", legend = rownames(genomeCov),
         fill = colors_per_library[myOrder], bty = "n", title = "", border = NA)
  
  barplot(mitoCov, border = NA, col = NA,  axes = FALSE, las = 2,
          xlab = NA, beside = T, 
          main = "Depth of coverage (mitochondrial)", axisnames = F)
  grid(nx = 0, ny = NULL)
    barplot(mitoCov, border = NA, las = 2,
          beside = T, col = colors_per_library[myOrder], ylab = "Depth of coverage",add = T)
  plot_letters("B", at = -12, line = 1.5)
dev.off()
```

```{r, eval = F}

# png(paste_path("Sequencing/coverage_log.png"), width = 7, height = 7,
#     res = 200, units = "in", bg = NA)  
pdf(paste_path("Sequencing/coverage_log.pdf"), width = 7, height = 9)
myOrder <- match(rownames(genomeCov), names(colors_per_library))

par(mfrow = c(2,1))
  barplot(genomeCov, border = NA, col = NA,  axes = FALSE, las = 2,
          xlab = NA, beside = T, axisnames = F, log = "y",
          main = "Depth of coverage (nuclear, log-scale)", ylim = c(0.0001,10))
  grid(nx = 0, ny = NULL)
  barplot(genomeCov, border = NA, las = 2,
          beside = T, col = colors_per_library[myOrder], 
          ylab = "Depth of coverage", add = T, log = "y"
  )
  legend(x = "topleft", legend = rownames(genomeCov),
         fill = colors_per_library[myOrder], bty = "n", title = "", border = NA)
  
  barplot(mitoCov, border = NA, col = NA,  axes = FALSE, las = 2,
          xlab = NA, beside = T, log = "y",
          main = "Depth of coverage (mitochondrial, log-scale)", axisnames = F, ylim = c(0.01, 1000))
  grid(nx = 0, ny = NULL)
    barplot(mitoCov, border = NA, las = 2, log = "y",
          beside = T, col = colors_per_library[myOrder], ylab = "Depth of coverage",add = T)

dev.off()
```

## Clonality

```{r}

boto$clon_perc <- boto$hits_clonality_endogenous/boto$hits_raw_endogenous
data <- data <- reshape(boto[,c("sample", 
                                "clon_perc",
                        "library")],
               timevar = "library", direction = "wide", 
               idvar = c("sample"))

myData <- t(as.matrix(data[,2:ncol(data)]))
colnames(myData) <- data$sample

#myData[is.na(myData)] <- 0
rownames(myData) <- sub(".*\\.", "", rownames(myData))
myData <- myData[match(lib_order, rownames(myData)),]
```

```{r}
myOrder <- match(rownames(myData), names(colors_per_library))

# png(paste_path("Sequencing/clonality.png"), width = 7, height = 5, 
#     res = 200, units = "in", bg = NA)
pdf(paste_path("Sequencing/clonality.pdf"), width = 7, height = 5)
barplot(myData, border = NA, 
        main = "Clonality",col = NA,  axes = FALSE, las = 2,
          ylab = "% of duplicates",xlab = NA, beside = T, axisnames = F
        #density = density, angle = angle1
  )
grid(nx = 0, ny = NULL)
barplot(myData, border = NA,  las = 2,
         beside = T, col = colors_per_library[myOrder],  axes = F, add = T
        #density = density, angle = angle1
  )
axis(2, at = seq(0,1, 0.1), labels = percent(seq(0, 1, 0.1)), las =1)

legend(x = "topleft", legend = rownames(myData),
       fill = colors_per_library[myOrder], bty = "n", title = "", border = NA)
dev.off()
```

## Sex determination

```{r}
boto <- boto[order(boto$sample, boto$library),]

data <- reshape(boto[,c("sample", "Ry", "Ry_confint",
                        "library")],
               timevar = "library", direction = "wide", 
               idvar = c("sample"))

myData <- t(as.matrix(data[,2:ncol(data)]))
colnames(myData) <- data$sample

rownames(myData) <- sub(".*\\.", "", rownames(myData))

index <- match(lib_order, rownames(myData))
myData <- myData[c(index, index+1),]


```

```{r, fig.width=7, fig.height=12}
myOrder <- match(rownames(myData), names(colors_per_library))
nSamples <- length(unique(boto$sample))
maxY <- ((nrow(myData)/2)*ncol(myData)) ; print(maxY)

pdf(paste_path("Sequencing/sex_determination.pdf"), width = 7, height = 10)

par(mar = c(4,6,6,2))
plot(myData[1:5,], 1:maxY+1, type = "n", bty = "n", axes = F,
      xlab = expression(R[y]), ylab = NA,
     xlim = c(0,max(myData[1:5,]+myData[6:10,], na.rm = T)),
     main = "Sex determination")

rect(xleft = 0, xright = 0.016, col = "snow2", border = NA, 
     ybottom  = 0, ytop = maxY+1)
rect(xleft = 0.075, xright = 0.11, col = "snow2", border = NA, 
     ybottom = 0, ytop = maxY+1)
points(y = 1:maxY,x = myData[1:5,], pch = 19, col = colors_per_library[myOrder])
abline(col = "gray", v = 0.042, lty = "dashed")

segments(y0 = 1:maxY, x0 = myData[1:5,] - myData[6:10,], 
         x1 = myData[1:5,]+myData[6:10,])
abline(h = seq(5, nSamples*5, 5) + 0.5, lty = "solid", col = "gray")
axis(side = 2, at =  seq(3, nSamples*5, 5) , labels = colnames(myData), las = 2)
axis(side = 1)

text(x = 0.01, y = maxY+2.5, label = "Female") 
text( x = 0.08, y = maxY+2.5, label = "Male")

legend(x = -0.001, y = maxY+8, col = colors_per_library[myOrder], pch = 19, 
       legend = lib_order, horiz = T,
       bg = NA, bty = "n")
dev.off()
```



## Contamination estimates

Work in progress

### Mitochondrial


```{r}
prefix <- "~/Projects/Botocudos/Files/Contamination/2019_08_06/"
samples <-unique(boto$sample)
rmTrans <- c("all", "rmTrans")


# subsample_contammmix <- function(path, sample, rmTrans, size = 100, lib = "All",
#                                  burnIn = 0.1){
#   if(lib == "All"){
#       myRda <- paste(path, sample, "/",sample, "_", rmTrans, ".Rdata", sep = "")
#   }else{
#     myRda <- paste(path, sample, "/", lib, "_", rmTrans, ".Rdata", sep = "")
#   }
#   print(myRda)
#   load(myRda)
#   notBurnIn <- as.integer(burnIn*size)
#   tiny <- # sample(
#     #remove burnin
#     c(res$chains[[1]][size - ((size- notBurnIn):0),2])#,
#                  #res$chains[[2]][,2], res$chains[[3]][,2])#, size)
#   result <- data.frame(estimate = tiny, damage = rmTrans, lib = lib,
#                        sample = sample)
#   return(result)
# }
# 
# contammix <- data.frame()
# for(s in samples){
#   libs <- unique(boto$library[boto$sample == s])
#   for(l in libs){
#     for(r in rmTrans){
#       tmp_contam <- subsample_contammmix(prefix, s, r, 100000, l)
#       contammix <- rbind(contammix, tmp_contam)
#     }
#   }
# }

```

```{r}
# contammix$lib <- factor(contammix$lib,
#                         levels = c("L1", "L2", "L3U", "mtCapture","All"), ordered = T)
# contammix$coverage <- 0
# for(s in unique(contammix$sample)){
#   for(l in unique(contammix$lib[contammix$sample == s])){
#     contammix$coverage <- boto$hits_coverage_mitochondrial[boto$sample == s & boto$library == l]
#   }
# }

# save(contammix, file="~/Projects/Botocudos/Files/Contamination/2019_08_06/contamination.Rdata")
load("~/Projects/Botocudos/Files/Contamination/2019_08_06/contamination.Rdata")
```

```{r}
path <- "~/Projects/Botocudos/Files/Contamination/2019_08_06/"
samples <- unique(boto$sample)
rmTrans <- c("all", "rmTrans")
annot_mito <- data.frame()

for(s in samples){
  libs <- unique(boto$library[boto$sample == s])
  for(lib in libs){
    for(r in rmTrans){
      if(lib == "All"){
        myRda <- paste(path, s, "/",s, "_", r, ".Rdata", sep = "")
      }else{
        myRda <- paste(path, s, "/", lib, "_", r, ".Rdata", sep = "")
      }
      #print(myRda)
      load(myRda)
      nReads <- dim(res$mnMatrix)[1]
      coverage <- boto$hits_coverage_mitochondrial[boto$sample == s & boto$library == lib]
      annot_tmp <- data.frame(nReads = nReads, coverage = coverage, damage = r, lib = lib,
                           sample = s)
      
      annot_mito <- rbind(annot_mito, annot_tmp)
    }
  }
}



MAP <- read.csv("~/Projects/Botocudos/Files/Contamination/2019_08_06/estimates.txt", header = F)

colnames(MAP) <- c("sample", "lib", "damage","map", "low", "high")
MAP$lib <- sub("S1", "L1", MAP$lib)
annot_mito <- join(annot_mito, MAP, by = c("sample", "lib", "damage"))

```

```{r, fig.width=14}
# ggplot(contammix, aes(x = sample, y = 1-estimate, color = damage, fill = lib)) +
#   geom_violin(draw_quantiles = c( 0.025, 0.5, 0.975), trim = F
#              )

# ggplot(contammix, aes(x = sample, y = 1-estimate, fill = lib, linetype = damage) )+
#   geom_violin(draw_quantiles = c( 0.025, 0.5, 0.975), trim = F) +
#   geom_abline(intercept = 0.05, slope = 0) + 
#   theme(axis.text.x = element_text(angle = 90))
```


Text on plots:

Mitochondrial coverage, 

number of reads (all polymorphisms),  number of reads (removing transitions)

```{r ggrides, fig.width=16, fig.height=16}
colors <- colors_per_library#c("#b08699","#e0b81f","#0d75ff", "gray")

# png(paste_path("Sequencing/contamination.png"), width = 12, height = 9,
#     res = 200, units = "in"#, bg = NA
#     )
pdf(paste_path("Sequencing/contamination.pdf"), width = 12, height = 10 )
ggplot(contammix, aes(y = lib, x = 1-estimate, fill = lib, linetype = damage) )+
  geom_density_ridges(scale = 0.98)+
  labs(title = "Contamination estimates (mitochondrial data)",
       x = NULL, y = "Estimate") +
  facet_wrap(sample~., ncol = 5)+
  geom_vline(xintercept = 0.05, color = "red") + 
  theme(axis.text.x = element_text(angle = 90, size = 10)) +
  scale_linetype_manual(name = "Polymorphisms considered", 
                        values = c("solid", "dotted")) +
  scale_fill_manual(values = colors_per_library, name = "Libraries") +
  geom_text(data = annot_mito, 
            nudge_y = 0.3,
            aes(y = lib, x = 0.2 + ifelse(damage == "all",0,0.3),
            label = prettyNum(nReads, big.mark = ",")),
            size = 2.4) +
  geom_text(data = annot_mito[annot_mito$damage == "all",], 
            nudge_y = 0.8, 
            aes(y = lib, x = 0.2,
            label = paste(prettyNum(round(coverage,1)), "x", sep = "")),
            size = 2.4) +
  coord_cartesian(xlim = c(0, 1)) +
  scale_x_continuous(labels = percent) 

dev.off()
```



```{shell, eval = F}
for mn in ${MN[@]} 
do  
  for type in all rmTrans 
  do 
    map=$(cat ${mn}/${mn}_${type}.txt |grep MAP |sed 's/.*: //' |cut -f1 -d ' ') 
    low=$(head -n9 ${mn}/${mn}_${type}.txt |tail -n1 |cut -f1 -d ' ') 
    high=$(head -n9 ${mn}/${mn}_${type}.txt |tail -n1 |cut -f2 -d ' ')
    echo "${mn},All,$type,$map,$low,$high" 
    for lib in L1 S1 L2 L3U  
    do 
      if [ -e ${mn}/${lib}_${type}.txt ] 
      then 
        map=$(cat ${mn}/${lib}_${type}.txt |grep MAP |sed 's/.*: //' | cut -f1 -d ' ') 
        low=$(head -n9 ${mn}/${lib}_${type}.txt |tail -n1 |cut -f1 -d ' ') 
        high=$(head -n9 ${mn}/${lib}_${type}.txt |tail -n1 |cut -f2 -d ' ') 
        echo "${mn},$lib,$type,$map,$low,$high"
      fi 
    done  
  done 
done >estimates.txt

```


### X

Work in progress

```{r}
minDepth <- seq(1,3)
path <- "~/Projects/Botocudos/Files/Contamination/2019_08_06/X/"
mySamples <- unique(boto$sample)
libs <- unique(boto$library)

s <- mySamples[1]
l <- libs[2]
md <- minDepth[3]

empty_df <- function(sample, lib){
  myDF <- data.frame(method = c("One-cns", "Two-cns"),
                     estimate = c(NA, NA), lb = c(NA, NA), ub = c(NA, NA),
                     err = c(NA, NA), nSites = c(NA, NA))
  return(myDF)
}

x_cont <- data.frame()
for(s in mySamples){
  for(l in libs){
    for(md in minDepth){
      if(l == "All"){
        tmp_res <- tryCatch(tmp_res <- read.table(paste(path, s, "/", s, 
                                                        "_md", md, ".result", sep = "")),
                            error = function(e) empty_df(s, l),
                            finally = function() return(tmp_res)
                            )
      }else{
        tmp_res <- tryCatch(tmp_res <- read.table(paste(path, s, "/", l, 
                                                        "/", l, "_md", md, ".result", sep = "")),
                            error =  function(e) empty_df(s, l),
                            finally = function() return(tmp_res))
      }
      colnames(tmp_res) <- c("method", "estimate", "lb", "ub", "err", "nSites")
      tmp_res$sample <- s
      tmp_res$library <- l
      tmp_res$depth <- md
      x_cont <- rbind(x_cont, tmp_res)
    }
  }
}


```

```{r}
x_cont <- join(x_cont, boto[, c("sample", "library", "sex")], by = c("sample", "library"))
x_cont <- x_cont[!is.na(x_cont$sex),]

males <- x_cont[x_cont$sex %in% c("XY", "consistent with XY but not XX"),]

# ggplot(males[males$method == "Two-cns" & males$depth >1,], aes(x = sample,
#                                                 y = estimate,
#                                                 ymin = lb, ymax = ub,
#                                                 color = library, shape = factor(depth))) +
#   geom_point(position = position_dodge(width = .2)) +
#   geom_errorbar(position = position_dodge(width = .2))

```

```{r}
males <- males[males$method == "Two-cns",]
males <- males[order(males$sample, males$library),]

# write.table(males, "~/Projects/Botocudos/Files/Contamination/2019_08_06/X/estimates_X_males.txt", 
            # sep = "\t", quote = F, col.names = T, row.names = F)
```

```{r}
males_fmt <- males

males_fmt$estimate <- percent(males_fmt$estimate)
males_fmt$lb <- percent(males_fmt$lb)
males_fmt$ub <- percent(males_fmt$ub)

#males_fmt
```

```{r}
males_dp3 <- males[males$depth == 3,]
colnames(annot_mito) <- c("mito_num_Reads", "mito_coverage", "mito_polymorphic_sites",
                          "library", "sample", "mito_estimate", "mito_low_est",
                          "mito_upper_est")
colnames(males_dp3) <- c("X_method", "X_estimate", "X_low_est", "X_upper_est",
                         "X_error", "X_num_sites", "sample", "library", "X_min_depth",
                         "sex")
estimates <- join(males_dp3, annot_mito[annot_mito$mito_polymorphic_sites == "all",],
                  by = c("sample", "library"))

estimates <- estimates[,c("sample", "library",
                          "mito_estimate", "mito_low_est", "mito_upper_est",
                          "X_estimate", "X_low_est", "X_upper_est",
                          "mito_num_Reads", "mito_coverage",
                          "X_num_sites", "X_min_depth",
                          "sex")]

estimates$mito_estimate <- 1 - estimates$mito_estimate
estimates$mito_low_est <- 1 - estimates$mito_low_est
estimates$mito_upper_est <- 1 - estimates$mito_upper_est
tmp_cont <- estimates$mito_low_est
estimates$mito_low_est <- estimates$mito_upper_est
estimates$mito_upper_est <- tmp_cont

for(myColumn in c("mito_estimate", "mito_low_est", "mito_upper_est",
                  "X_estimate", "X_low_est", "X_upper_est",
                  "mito_coverage") ){
  estimates[, myColumn] <- signif(estimates[,myColumn], 3)
}

for(myColumn in c("mito_estimate", "mito_low_est", "mito_upper_est",
                  "X_estimate", "X_low_est", "X_upper_est"
                  ) ){
  estimates[, myColumn] <- percent(estimates[,myColumn], 2)
}
```

```{r, eval = F}
knitr::kable(estimates, format = "markdown")
```

### Supplementary table 1

```{r}
mito <- MAP
colnames(mito) <- c("sample", "library","damage", "cont_MT", "cont_MT_upper", "cont_MT_lower")
mito[,c("cont_MT", "cont_MT_upper", "cont_MT_lower")] <- 1 - mito[,c("cont_MT", "cont_MT_upper", "cont_MT_lower")]
mito <- mito[mito$damage == "all",]
mito$damage <- NULL

X <- males_dp3
colnames(X) <- c("X_method", "cont_X","cont_X_lower", "cont_X_upper", "X_err","X_sites_dp3", "sample", "library", "depth", "sex")
X <- X[,c("sample", "library","cont_X", "cont_X_lower", "cont_X_upper")]


mito[,c("cont_MT", "cont_MT_upper", "cont_MT_lower")] <- round( mito[,c("cont_MT", "cont_MT_upper", "cont_MT_lower")], 2)
X[, c("cont_X", "cont_X_lower", "cont_X_upper")] <- round(X[, c("cont_X", "cont_X_lower", "cont_X_upper")], 2)

```

```{r}
boto_complete <- join(boto, mito, by = c("sample", "library"))
boto_complete <- join(boto_complete, X, by = c("sample", "library"))



boto_complete$cont_MT <- paste(percent(boto_complete$cont_MT), 
                               " (", percent(boto_complete$cont_MT_lower), 
                               ", ", percent(boto_complete$cont_MT_upper), ")",
                               sep = "")

boto_complete$cont_X <- paste(percent(boto_complete$cont_X), 
                               " (", percent(boto_complete$cont_X_lower), 
                               ", ", percent(boto_complete$cont_X_upper), ")",
                               sep = "")
boto_complete$seq_retained_frac <- 1 - boto_complete$seq_trash_se_frac

to_round <- as.data.frame(matrix(data=c("Ry", 3,
              "Ry_confint", 3,
              "seq_trash_se_frac", 2,
              "seq_retained_length", 1,
              "hits_raw_frac_endogenous", 4,
              "hits_unique_frac_endogenous", 4, 
              "hits_coverage_endogenous", 3,
              "hits_length_endogenous", 1,
              "hits_unique_frac_mitochondrial", 6,
              "hits_coverage_mitochondrial",1,
              "hits_length_mitochondrial", 1,
              "hits_length_nuclear",1,
              "hits_coverage_nuclear", 3),
              ncol = 2, byrow = T), stringsAsFactors = F)
to_round$V2 <- as.integer(to_round$V2)

#to_round_digits <- c(3,3,2,1,4,4,3,1,6,1,1,3)
for(i in 1:nrow(to_round)){
  boto_complete[,to_round[i,1]] <- round(boto_complete[, to_round[i,1]], 
                                       to_round[i,2])

}

# Compute other numbers to report in the table
boto_complete$hits_clonality_endogenous_frac <- percent(boto_complete$hits_clonality_endogenous/boto_complete$hits_raw_endogenous)

boto_complete$hits_raw_frac_endogenous <- percent(boto_complete$hits_raw_frac_endogenous, accuracy = 0.01)
boto_complete$hits_unique_frac_endogenous <- percent(boto_complete$hits_unique_frac_endogenous, accuracy = 0.01)

#boto_complete$hits_raw_frac_nuclear <- percent(boto_complete$hits_raw_frac_nuclear, accuracy = 0.01)
boto_complete$hits_unique_frac_nuclear <- percent(boto_complete$hits_unique_frac_nuclear, accuracy = 0.01)

boto_complete$hits_unique_frac_mitochondrial <- percent(boto_complete$hits_unique_frac_mitochondrial, accuracy = 0.0001)
boto_complete$seq_trash_se_frac <- percent(boto_complete$seq_trash_se_frac)
boto_complete$seq_retained_frac <- percent(boto_complete$seq_retained_frac)

boto_complete$Ry <- paste(boto_complete$Ry, " (", boto_complete$Ry-boto_complete$Ry_confint, ", ",
                 boto_complete$Ry+boto_complete$Ry_confint, ")", sep = "")

# Add thousands separator
need_comma <- which(unlist(sapply(1:ncol(boto_complete), 
                                  function(x) class(boto_complete[,x])) == "integer"))

boto_complete[,need_comma] <- sapply(need_comma, 
                                     function(x) prettyNum(boto_complete[,x], big.mark = ","))
```

```{r}
myColNames <- c("sample",                         "Sample", "National Museum of Brazil (Museu Nacional, MN) identifier",
                # "target",                         "Sample2",
                "Area",                           "Region", "Sample origin (region) according to MN's archives",
                "State",                          "State", "Sample origin (state) according to MN's archives", 
                "CalibratedDate",                 "Calibrated date (AD)", "OxCal estimated mean and standard deviation for calibrated dates",
                
                "library",                        "Library", "DNA library identifier",
                "lib_type",                       "library_type", "Library type",
                
                "seq_reads_se",                   "sequenced_reads", "Number of sequenced reads",
                #"seq_trash_se",                   "trashed_reads_trimming", 
                #"seq_trash_se_frac",              "fraction_trashed_reads_trimming",
                "seq_retained_reads",             "retained_reads_trimming", "Number of reads retained after trimming",
                "seq_retained_frac",              "percent_retained_reads_trimming", "Percentage of sequenced reads retained after trimming",
                #"seq_retained_nts",               "retained_nucleotides_trimming",
                "seq_retained_length",            "length_retained_reads", "Average read length for retained reads after trimming",
                
                "hits_clonality_endogenous",      "duplicated_reads", "Number of reads flagged as PCR duplicates by picardtools MarkDuplicates",
                "hits_clonality_endogenous_frac", "percent_duplicated_reads", "Percentage of reads flagged as PCR duplicates by picardtools MarkDuplicates",
                
                "hits_unique_endogenous",         "unique_reads_endogenous", "Number of unique reads mapped to the human genome (build 37.1)",
                "hits_unique_frac_endogenous",    "percent_unique_reads_endogenous", "Percentage of retained reads that were uniquely mapped to the human genome",
                "hits_length_endogenous",         "length_reads_endogenous", "Average length of uniquely mapped reads",
                "hits_coverage_endogenous",       "DoC_endogenous", "Depth of coverage on human genome (number of unique bases divided by the length of the human genome reference build 37.1)",
                
                #"hits_unique_nuclear",            "unique_reads_nuclear",
                #"hits_coverage_nuclear",          "DoC_nuclear",
                "hits_unique_mitochondrial",      "unique_reads_MT", "Number of unique reads mapped to the mitochondrial genome",
                "hits_coverage_mitochondrial",    "DoC_MT", "Depth of coverage on mitochondrial genome (number of unique bases divided by the length of the mitochondrial genome, build 37.1)",
                #"hits_raw_endogenous",            "reads_endogenous_raw",
                #"hits_raw_frac_endogenous",       "fraction_reads_raw_endogenous",


                #"hits_unique_frac_nuclear",       "percent_unique_reads_nuclear",
                #"hits_length_nuclear",            "length_reads_nuclear",
                
                #"hits_unique_frac_mitochondrial", "fraction_unique_reads_mitochondrial",
                "hits_length_mitochondrial",      "length_reads_mitochondrial", "Average length of uniquely mapped reads on mitochondrial genome",
                
                "cont_MT",                        "contamination_MT", "Maximum a posteriori estimate and 95% credible interval for contamination based on the mitochondrial genome (contamMix)",
                #"cont_MT_upper",                  "contamination_mitochondrial_upper",
                #"cont_MT_lower",                  "contamination_mitochondrial_lower",
                "cont_X",                         "contamination_X", "Maximum likelihood estimate and 95% confidence interval for contamination based on X-chromosome data (contaminationX)",
                #"cont_X_lower",                   "contamination_X_lower",
                #"cont_X_upper",                   "contamination_X_upper",
                
                "Ry",                             "Ry", "Ratio of the reads mapping to the Y-chromosome and the reads mapping to sexual chromosomes (X and Y) and 95% confidence interval",
                #"Ry_confint",                     "Ry_confint",
                "sex",                            "Sex", "Molecular sex determined by the Ry ratio"
                )

change_name <- function(name){
  new_name <- myColNames[which(myColNames == name)[1] + 1]
  ifelse(is.na(new_name), return(name), return(new_name))
}


colnames(boto_complete) <- sapply(colnames(boto_complete), function(name) change_name(name))

boto_complete <- boto_complete[,myColNames[seq(2, length(myColNames), 3)]]

boto_complete$Library <- factor(boto_complete$Library, levels = c("mtCapture",
                                                                  "L1",
                                                                  "L2",
                                                                  "L3U",
                                                                  "All"),
                                ordered = T)

boto_final <- subset(boto_complete, select = myColNames[seq(2, length(myColNames), 3)])
```


```{r, eval = F}
# boto_complete <- boto_complete[order( boto_complete$Sample, boto_complete$Library,boto_complete$percent_unique_reads_endogenous ),]
boto_final <- boto_final[order(boto_final$Sample, boto_final$Library, boto_final$percent_unique_reads_endogenous),]
write.table(boto_final, "~/Projects/Botocudos/Files/Summaries/2019_07_26/Sup_Table_1.txt",
            col.names = T, row.names = F, sep = "\t", quote = F)

```

```{r}
require(xlsx)
wb_name <- "tabla1.xlsx"
sh_name <- "SI_Table1"
# define workbook and sheet
wb <- createWorkbook(type="xlsx")
sheet <- createSheet(wb, sheetName = sh_name)

# Define row style to be used
mySamples <- data.frame(sample = levels(boto_final$Sample), 
                        color = rep(c("blue", "white"), 
                                    round(length(unique(boto_final$Sample))/2)))
white_row <- CellStyle(wb) 
blue_row <- CellStyle(wb) + Fill(foregroundColor = "lightblue")

# Function to add a row with specified cell style
# xlsx.addRow <- function(sheet, rowIndex, data, rowStyle, sample){
#   rows <- createRow(sheet, rowIndex = rowIndex)
#   sheetRow <- createCell(rows, colIndex=1)
#   sapply(data, function(x) setCellValue(sheetRow[[1,1]], x))
#   cells <- getCells(getRows(sheet, rowIndex = rowIndex))
#   
#   if(mySamples$color[mySamples$sample == sample] == "blue"){
#      lapply(names(cells), function(ii) setCellStyle(sheetRow[[ii]], blue_row))
#   }else{
#     #setCellStyle(sheetRow[[1,1]], white_row)
#   }
# }

# for(i in 1:nrow(boto_final)){
#   xlsx.addRow(sheet = sheet, rowIndex = i, data = boto_final[i,],
#               sample = as.character(boto_final$Sample[i]))
# }

# Add table + column name explanation
addDataFrame(boto_final, sheet, startRow = 1, startColumn = 1, row.names = F)

# Add color
rows <- getRows(sheet, rowIndex = 2:(nrow(boto_final)+1))
cells <- getCells(rows, colIndex = 1:ncol(boto_final))
values <- lapply(cells, getCellValue)

highlight_blue <- c()
for(i in names(values)){
  if(values[i] %in% mySamples$sample){
    currentSample <- values[i]
  }
  if(mySamples$color[mySamples$sample == currentSample] == "blue"){
    highlight_blue <- c(highlight_blue, i)
  }
}
lapply(names(cells[highlight_blue]), function(ii) setCellStyle(cells[[ii]], blue_row))
  


myLegend <- as.data.frame(matrix(myColNames, ncol = 3, byrow = T) )[,2:3]; #print(myLegend)
addDataFrame(myLegend, sheet, startRow = nrow(boto_final) + 3,
             startColumn = 1, row.names = F, col.names = F)

autoSizeColumn(sheet, colIndex = 1:ncol(boto_final))
saveWorkbook(wb, wb_name)

```


<!-- ## Complexity curves -->

<!-- For some libraries -->

```{r, eval = F}
complexity <- function(myTable, maxSize = 5, total_sequenced){
  constructor <- get_constructor(myTable, r=1)
  
  exp_sizes <- seq(0, maxSize, 0.1)
  
  df <- data.frame(sequenced = exp_sizes*total_sequenced ,
                   expected = sapply(exp_sizes, 
                                     function(x) constructor$f(x)),
                   lower = sapply(exp_sizes, 
                                  function(x) constructor$lb(x)),
                   upper = sapply(exp_sizes, 
                                  function(x) constructor$ub(x)))
  return(df)
}


L2 <- read.table("~/Projects/Botocudos/Files/PreSeq/2019_08_12/MN0008/L2.duphist")
L3U  <- read.table("~/Projects/Botocudos/Files/PreSeq/2019_08_12/MN0008/L3U.duphist")

l2_complex <- complexity(L2, maxSize = 15, 
                         total_sequenced = boto$seq_reads_se[
                           boto$sample == "MN0008" & boto$library == "L2"])
l3_complex <- complexity(L3U, maxSize = 15, 
                         total_sequenced = boto$seq_reads_se[
                           boto$sample == "MN0008" & boto$library == "L3U"])
```


```{r, eval = F}
oneLane <- 400000000
already_seq <- data.frame(lib = boto$library[boto$sample == "MN0008"],
                          sequenced = boto$seq_reads_se[boto$sample == "MN0008"],
                          expected = boto$hits_unique_endogenous[boto$sample == "MN0008"],
                          lower = boto$hits_unique_endogenous[boto$sample == "MN0008"],
                          upper = boto$hits_unique_endogenous[boto$sample == "MN0008"],
                          coverage = boto$hits_coverage_endogenous[boto$sample == "MN0008"])

myResult <- list(L2 = l2_complex, L3U = l3_complex)
plots <- vector("list", 2)
libs <- c("L2", "L3U")
myCoverage <- round(already_seq$coverage[already_seq$lib == "L2"],1)
myMapped <- already_seq$expected[already_seq$lib == "L2"]/already_seq$sequenced[already_seq$lib == "L2"]
for(i in 1:2){
  lib <- libs[i]
  plots[[i]] <- ggplot(myResult[[lib]], aes(x = sequenced, y = expected,
                                            ymin = lower, ymax = upper)) +
    geom_line() +
    geom_ribbon(alpha = 0.6, col = "chartreuse4") +
    geom_vline( xintercept = seq(oneLane,max(myResult[[lib]]$sequenced),oneLane), 
                linetype = "dashed", color = "gray") +
    geom_hline( yintercept = myMapped*seq(oneLane,max(myResult[[lib]]$sequenced),oneLane), 
                linetype = "dashed", color = "gray") +
    geom_point(data = already_seq[already_seq$lib == lib,], color = "blue") +
    scale_x_continuous(labels = seq(oneLane,max(l2_complex$sequenced),oneLane)/oneLane,
                       breaks = seq(oneLane,max(l2_complex$sequenced),oneLane)) +
    labs(x = "Sequenced lanes (~400M reads)", y = "Approximate genome coverage",
         title = paste(lib,", MN0008", sep = "")) +
    scale_y_continuous(labels = myCoverage*seq(oneLane,
                                                         max(l2_complex$sequenced),oneLane)/oneLane,
                       breaks = myMapped*seq(oneLane,max(l2_complex$sequenced),oneLane)) +
    coord_cartesian(ylim = c(0, 6*oneLane))
  
}

# png("~/Projects/Botocudos/Plots/Sequencing/complexity.png", width = 10,
#     height = 6, res = 200, units = "in")

plot_grid(plotlist = plots, ncol = 2)

# dev.off()

```

```{r, eval = F}
pdf("~/Projects/Botocudos/Plots/Sequencing/complexity_MN0008_L2.pdf", 
    width = 7, height = 7)
print(plots[[1]])
dev.off() 


```

<!-- ## Admixture -->

```{r, eval = F}
path <- "~/Projects/Botocudos/Files/ADMIX/2019_09_01/Wollstein_50ind_MN0008_libs_k"
myColors <- c("#b08699", "#c74300", "#5c8a73", "#008aa1", "#ff4dc9", "#b319ab", "#718600", "#06004f")
names <- read.table("~/Projects/Botocudos/Files/ADMIX/2019_09_01/wollstein_ordered.txt", stringsAsFactors = F)[,5]
names <- c("L2", "L3U", "L1", "MN0008_merged", names)
myLines <- sapply(unique(names), function(x) max(which(names == x)))
myNames <- sapply(unique(names), function(x) mean(which(names == x)))

pdf("~/Projects/Botocudos/Plots/Sequencing/admixture_MN0008.pdf", height = 6)
par(mfrow = c(8,1))
par(mar = c(2, 4 , 1, 2))
for(k in paste(seq(2, 8 ), c( 91,96,26,76,96,71,16 ), sep = '_')){
  file <- read.table(paste(path, k, ".qopt", sep = ""))
  myAdmix <- t(as.matrix(file))
  barplot(myAdmix, col = myColors, border = NA, space = 0,
          xlab = NULL, ylab = sub("_.*", "",k))
  abline(v = myLines, col = "white")
}

text(myNames-0.5, -1., names(myNames), xpd=NA, cex = 1, srt = 90)
dev.off()
```

